### 2020-06-13 paper/paper.tex
Paper setup.

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Conference Paper Title*\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
This document is a model and instructions for \LaTeX.
Please observe the conference page limits.

\section{Ease of Use}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
paper in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
\centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2020-06-13 nim_hw2_report/evoman_report.tex
Started paper.

\documentclass[conference]{IEEEtran}

\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{hyperref}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{url}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{A Comparison of Algorithms Playing EvoMan}

\author{\IEEEauthorblockN{Cojocaru Gabriel-Codrin, Dinu Sergiu-Andrei}
\IEEEauthorblockA{\textit{"Alexandru Ioan Cuza" University (UAIC) Iasi} \\
\textit{Faculty of Computer Science}\\
\textit{Computational Optimization Master}\\
Iasi, Romania \\
\{gabriel.cojocaru, sergiu.dinu\}@info.uaic.ro\\
\{gcodrincojocaru, dinusergiuandrei997\}@gmail.com}
\and
}

\maketitle

\begin{abstract}
This paper describes a comparison between algorithms for
evolving agents for the game Evoman.
The algorithms we have tested are Q-learning with neural networks, sparse reward neuroevolution and
iterative neuroevolution algorithms.
Both neuroevolution algorithms were tested with genetic algorithms and particle swarm optimization.
As of the results, the Q-learning approach lead to the worst results.
There was no significant difference between the performance of the iterative neuroevolution
and the sparse neuroevolution when used with genetic algorithms.
When particle swarm optimization was used, the iterative approach lead to worse results than the sparse one.
Both of the neuroevolutions with particle swarm optimization lead to much better results than both of the
neuroevolutions with genetic algorithms.
We have also shown that there is not a significant change in the performance of sparse neuroevolutive
algorithms with particle swarm optimization even when the cognitive and social weights suffer great changes.
\end{abstract}

\begin{IEEEkeywords}
game-playing agent, artificial intelligence, EvoMan, genetic algorithm, reinforcement learning,
q-learning, neuroevolution, particle swarm optimization
\end{IEEEkeywords}

\section{Introduction}\label{sec:introduction}
EvoMan~\cite{karinemiras,evoman} is a framework for testing competitive game-playing agents.
This framework is inspired by Mega Man II~\cite{capcom}, the game created by Capcom.
In the original game, the player would have to beat $8$ opponents and acquire their weapons
as they are defeated.
The additional difficulty of EvoMan comes from the fact the player has to defeat all
the opponents using only the starting weapon.
The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman_framework}}
and it is currently compatible with Python 3.6 and 3.7.
There is also an extensive documentation
available\footnote{\url{https://github.com/karinemiras/evoman_framework/blob/master/evoman1.0-doc.pdf}}.

The agent will collect information about the environment through 20 sensors (Fig 1):
\begin{itemize}
    \item 16 correspond to horizontal and vertical distances to a maximum of 8 different opponent projectiles.
    \item 2 correspond to the horizontal and vertical distance to the enemy.
    \item 2 describe the directions the player and the enemy is facing
\end{itemize}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/Evoman3.png}
    \caption{Sensors available for the player.}
    \label{fig:sensors}
\end{figure}
The actions which the agent may take are:
\begin{itemize}
    \item walk left
    \item walk right
    \item jump
    \item shoot
    \item release of the jump
\end{itemize}

The lives of the player and the enemy start at 100.
Everytime one of them gets hit, their life deplenishes.
Whoever's life reaches 0 loses the game.

In our work, we have focused on the designing specialized models able of learning how to win a
match against specific enemies of each level.

\section{Method}\label{sec:method}
The formula used for evaluating an agent is:

\boldmath
\begin{gather*}
    0.9 * (100 - enemy\_life) + 0.1 * player\_life\\
    - ln(nr\_of\_game\_timesteps)\\
\end{gather*}
\unboldmath
We have decided to reduce the fitness in time in order to encourage the agents to be active.
We are mainly rewarding the damage done to the opponent, while also considering, but with less importance,
the player's life.
\newpage
The author defines a success criterion for a solution as:
\begin{itemize}
    \item Effectiveness: to zero the life of the enemy.
    \item Efficiency: the amount of life preserved by the player.
\end{itemize}
From our reward function, it is visible that we have biased our models towards effectiveness,
rather than efficiency.
This decision is solely based on intuition.

We have tested five algorithms to control the EvoMan agent.

\subsection{Q-learning with Neural Networks}\label{subsec:q-learning-with-neural-network}
The first algorithm is a classic Q-learning\cite{q_learning} algorithm with neural networks which we have used
as a baseline model.

We used a neural network to predict the reward function for each possible move
(left, right, shoot, jump, release of jump) from a given state.
We then take the action with the highest predicted reward.
The input of this neural network is composed of the current game sensors and the
previous 2 game sensors with the moves taken at that specific point.

The neural network used for experiments has 2 hidden layers with 32 neurons each.
Each layer has l2 regularization applied with a weight decay of 0.01 and sigmoid activation.
After each predicted move, we updated the neural network using backpropagation,
using as input the game sensors and as output the true reward function.
We have trained the agent on 5000 games.

\subsection{Sparse Reward Neuroevolution}\label{subsec:sparse-reward-neuroevolution}
The second algorithm is a sparse reward neuroevolution\cite{neuro}.
We called it "sparse" reward because an agent doesn't find out how well it's doing until
the end of the game, with no feedback during the game.
An individual is represented as the weights of a neural network.

The neural network role is to predict the next move using the current game sensors and
the previous 2 game sensors with the moves taken at that specific point.
We used 2 hidden layers with 32 neurons for each individual.

We start with randomly initialized neural network weights and then represent them as a bitstring.
The weights used for the neural network were values between 2 and -2 with a precision of 6 digits.

The search of individuals that lead to good fitness values can not be done with evolutive algorithms.

For evaluation, the bitstring is transformed into the weights of the neural network and a game is played,
from which we can obtain the individual fitness.

\begin{itemize}
    \item Genetic Algorithms\cite{genetic_algorithm} \\
    Since we are already representing our individuals as bitstrings, the classical operations of a
genetic algorithm on bits can be applied: crossover, mutation and selection.
    The configurations used for the genetic algorithm are:
    \begin{itemize}
        \item population size 50
        \item number of generation 500
        \item crossover rate 0.7
        \item mutation rate 0.008 and 0.1
    \end{itemize}
    The solution of the genetic algorithm is the best individual from the last generation.
    \newpage
    \item Particle Swarm Optimization\cite{pso} \\
    The configurations used for the particle swarm optimization algorithm are:
    \begin{itemize}
        \item population size 30
        \item number of iterations 200
        \item cognitive weight 0.4 - 0.8 - 1.5
        \item social weight 0.8 - 0.4 - 3
    \end{itemize}
\end{itemize}

\subsection{Iterative Neuroevolution}\label{subsec:iterative-neuroevolution}
The last algorithm is an iterative neuroevolution.
We called it "iterative" because the agents are trained on a small number of game steps
first and then the number of game timesteps slowly increases.

After a number generations we increase the number of game timesteps the agents are allowed to train on.
The fitness function was scaled based on the number of game timesteps in a way that if an agent
training on x game timesteps will always have a fitness lower than an agent training on y game
timesteps if x is lower than y.

We used the same genetic algorithm, neural network configuration and
very similar particle swarm optimization configurations as the ones for the sparse reward neuroevolution.
The difference in configuration of the neuroevolution with particle swarm optimization is that we are
no longer testing configurations using only a constant inertia weight.
The new try is with a uniformly decreasing inertia weight with respect to the percent of the iterations passed,
having values from 1 to 0.3.

\section{Experiment}\label{sec:experiment}
For the highest game difficulty, for all algorithms we have run 30 experiments.
For both neuroevolutive algorithms we used the same size for the neural network of an
individual while the reward predictor neural network used in Q-learning had the same size.
We also tried to balance the number of iterations used for training the genetic algorithms
and particle swarm optimization algorithms with the number of games the Q-learning agent trains on. \\

Especially in the case of the neuroevolution with particle swarm optimization, the algorithm configurations were
tested and decided empirically along the way. \\

When decreasing the game difficulty to observe how the best model behaves in increasingly easier environments,
the number of experiments was not fixed and was determined by a fixed time frame equal to 16 hours.
This lead to a number of runs per difficulty level varying from 10 to 24.
This liberty was taken due to the fact that the results of such an experiment are of a lesser importance than the ones
at the highest difficulty.

\newpage
\section{Results}\label{sec:results}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/q_vs_ga_iterative.eps}
    \caption{The fitness comparison between Q-learning and iterative neuroevolution with genetic algorithms.}
    \label{fig:q_vs_ga_iterative}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/ga_mutation_rates.eps}
    \caption{The fitness comparison between sparse neuroevolutions with genetic algorithms with mutation rate of 0.1 and 0.008.}
    \label{fig:ga_mutation_rates}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/overview.eps}
    \caption{The fitness comparison between the iterative neuroevolution and the sparse neuroevolution,
    both with genetic algorithms and particle swarm optimization algorithms. }
    \label{fig:overview}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/pso_sparse_vs_iterative_time.eps}
    \caption{The time comparison between sparse and iterative neuroevolutions with particle swarm optimization.}
    \label{fig:pso_sparse_vs_iterative_time}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/pso_configs.eps}
    \caption{The fitness comparison between sparse neuroevolutions with particle swarm optimization
    with different cognitive and social weights.}
    \label{fig:pso_configs}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/pso_inertia.eps}
    \caption{The fitness comparison between neuroevolutions with particle swarm optimization with constant inertia
    weight and with uniformly decreasing in time inertia weight.}
    \label{fig:pso_inertia}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/pso_levels.eps}
    \caption{The fitness comparison of neuroevolutions at different game difficulty levels.}
    \label{fig:pso_levels}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/pso_levels_time.eps}
    \caption{The time comparison of neuroevolutions at different game difficulty levels.}
    \label{fig:pso_levels_time}
\end{figure}

\newpage

\section{Interpretation of Results}\label{sec:interpretation-of-results}

From Fig.~\ref{fig:q_vs_ga_iterative} we can clearly see that the neuroevolution with the genetic algorithm
lead to much better results than the Q-learning algorithm with neural networks.
Despite running 5000 games, a number of evaluations comparable to the one of the neuroevolution,
the Q-learning algorithm did not manage to successfully train an agent.
The neuroevolution with genetic algorithms lead to a good fitness score, but the resulting agent was not good enough
to win the game.

In Fig.~\ref{fig:ga_mutation_rates} we can see the fitness comparison between neuroevolutions with genetic algorithms
with extremely different mutation rates.
This was tried with the hope that the high mutation rate could lead to a better exploration of the game space.
This problem lead to the usual intuition of a genetic algorithm stating that a high mutation rate leads to worse results.
The comparison between the two is clear, with the low mutation rate leading to better results.

In Fig.~\ref{fig:overview} we can see the fitness comparison between the iterative and the sparse neuroevolution,
both with genetic algorithms and particle swarm optimization.
There was not strong enough evidence to suggest a statistically significant difference between the sparse and iterative
neuroevolutions with genetic algorithms.
This means that the iterative approach leads to results that are as good as the sparse approach, but with much less
processing power and time.

By using Welch's t-test, we have shown that the sparse neuroevolution leads to a better fitness value than an iterative
neuroevolution when both are using particle swarm optimization.
This means that in order to benefit of the reduced processing power and time of the iterative approach,
some of the performance has to be sacrificed.

By using the same statistical test we have observed that both neuroevolutions with particle swarm optimization lead
to better results than the neuroevolutions with genetic algorithms.

In Fig.~\ref{fig:pso_sparse_vs_iterative_time} we can see the time comparison between sparse and iterative neuroevolutions
with particle swarm optimization.
Welch's t-test shows that there is significant proof which suggests that the iterative approach is faster than the
sparse approach.
In order to compare the execution times of neuroevolutions with particle swarm optimization and the execution times
of neuroevolutions with genetic algorithms, some observations need to be made:
\begin{itemize}
    \item The neuroevolution with genetic algorithms ran for 500 generations with a population size of 50.
    \item The neuroevolution with particle swarm optimization ran for 200 generations with a population size of 30.
    \item The experiments were run on different machines and on a different number of cores, deeming a direct comparison as incorrect.
\end{itemize}
A huge difference in time was felt during the execution of experiments of the two approaches,
so some time comparison needs to be mentioned.
The adjustments made in order to realize such a comparison are:
\begin{itemize}
    \item The genetic algorithm was limited at 200 generations.
    \item The genetic algorithm was ran 3 times on the same machine as the particle swarm optimization ran on.
\end{itemize}

The average execution time of the iterative neuroevolution with genetic algorithms was 3921 seconds, which is much
higher than both the sparse and iterative neuroevolutive algorithms with particle swarm optimization.

The average execution time of the sparse neuroevolution with genetic algorithms was 6570 seconds,
which is more than twice the median execution times for both neuroevolutions with particle swarm optimization.

In Fig.~\ref{fig:pso_configs} we can see the fitness scores of sparse neuroevolutions with particle swarm optimization
with different cognitive and social weights.
The first experiment was run with cognitive weight 1.5 and social weight 3.
Due to not knowing the roughness of the problem's landscape, we had to check whether high or low weights lead to better results.

The second experiment was run with cognitive weight 0.4 and social weight 0.8.
There is not enough evidence of a significant difference between the fitness results of the two mentioned experiments.
Due to the fact that a significant number of runs with the lower weights lead to higher fitness values and the fact that the
average result when using lower weights is higher than the average result when using higher weights, the lower
weights were used in the following experiments.

The next experiment used cognitive weight 0.8 and social weight 0.4.
It lead to lower median and mean fitness values than the last two configurations.

Welch's t-test did not find enough proof for the existence of a significant difference between the
results of the experiments with the second and third configurations, but it did find that the first configuration
lead to better results than the third one.

In Fig.~\ref{fig:pso_inertia} we can observe the fitness comparison between neuroevolutions with particle swarm
optimization with constant inertia wight and with uniformly decreasing in time inertia weight.
In the case when the inertia weight is not constant it uniformly decreases from 1 to 0.3 inversely proportional
to the number of passed iterations.
This was tried based on the intuition that in the later stages of evolution less exploration is required.
Despite de distribution of the fitness scores looking very differently, Welch's t-test did not
find enough proof for the existence of a significant difference in performance.

In Fig.~\ref{fig:pso_levels} we can observe the fitness comparison of sparse neuroevolutions with particle swarm
optimization at different game difficulty levels:
\begin{itemize}
    \item At level 5 the player loses with the enemy having approximately between a third and a half of life remaining.
    \item At level 4 the player barely loses, leaving the enemy with very little life remaining.
    \item At level 3 the player barely wins, remaining with very little life at the end of the game.
    \item The authors of the environment tested their specialized agents at the second difficulty.
    We have replicated their best results by winning the game while taking very little damage.
    \item At level 1 the game is very easy to win for the agents, taking almost no damage.
\end{itemize}

In Fig.~\ref{fig:pso_levels_time} we observe the time comparison of sparse neuroevolutions with particle swarm
optimization at different game difficulty levels.
We can see that the time is not proportional to the difficulty.
A possible explanation could be based on the game's outcome at each difficulty:
\begin{itemize}
    \item At level 5 the agent loses fast.
    \item At level 4 the agent loses after a longer game.
    \item At level 3, the level that lasts the most on average, the agent and the enemy have long and balanced games.
    \item At level 2 the agent starts to win rather fast.
    \item At level 1 the agent quickly inflicts a lot of damage and each game ends with a fast win.
\end{itemize}

\section{Conclusion}\label{sec:conclusion}
We have shown that there are significant differences in the results of neuroevolutions utilizing different evolutive algorithms.
The Q-learning algorithm did not succeed in capturing any aspect of the environment landscape.
Solving this reinforcement learning problem with a neuroevolutive algorithm with a genetic algorithm using an
iterative approach saves a lot of processing power and time while leading to results that are not worse, and sometimes
better, than the ones obtained through a sparse approach.  \\
The method that lead to much better results was a neuroevolution with the particle swarm optimization algorithm.
All configurations of neuroevolutions, both sparse or iterative and with any parameters, with particle swarm optimization algorithm
lead to better results than any of the neuroevolutions with genetic algorithms.
Unfortunately, using an iterative approach when using a neuroevolution with particle swarm optimization does not lead
to results as good as in the sparse approach.
Of course, this is not a definite conclusion, as extensive parameter tuning could very well make an iterative approach lead
to better results than a sparse approach.
The reduced execution power and time could make the iterative approach a good enough compromise for more complex reinforcement learning problems. \\

We can definitely say that the neuroevolution with particle swarm optimization has the most potential of high scores in the EvoMan game,
with the low execution time being a huge advantage over other reinforcement learning approaches.

\newpage
%\section*{References}

\begin{thebibliography}{00}
\bibitem{evoman} Fabricio Olivetti de Franca, Denis Fantinato, Karine Miras, A.E. Eiben and Patricia A. Vargas.
"EvoMan: Game-playing Competition" arXiv:1912.10445
\bibitem{karinemiras} de Araújo, Karine da Silva Miras, and Fabrício Olivetti de França.
"An electronic-game framework for evaluating coevolutionary algorithms." arXiv:1604.00644 (2016).
\bibitem{neuro} Floreano, D., Dürr, P. & Mattiussi, C. Neuroevolution: from architectures to learning. Evol. Intel. 1, 47–62 (2008). https://doi.org/10.1007/s12065-007-0002-4
\bibitem{ga}Goldberg, David (1989).
Genetic Algorithms in Search, Optimization and Machine Learning.
Reading, MA: Addison-Wesley Professional. ISBN 978-0201157673.
\bibitem{capcom} M. MEGA, "Produced by capcom, distributed by capcom, 1987," System: NES.
\bibitem{q_learning} Watkins, C.J.C.H., Dayan, P. Q-learning. Machine Learning 8, 279–292 (1992)
\bibitem{genetic_algorithm} Holland J.H., Genetic Algorithms and Adaptation. Adaptive Control of Ill-Defined Systems, 1984, Volume 16 ISBN 978-1-4684-8943-9
\bibitem{pso} Kennedy, J.; Eberhart, R. (1995). "Particle Swarm Optimization". Proceedings of IEEE International Conference on Neural Networks. IV. pp. 1942–1948.

\end{thebibliography}

\end{document}

### 2020-06-13 paper/paper.tex
Added csv files for plots.
Started paper.

@@
 \usepackage{xcolor}
+
+\selectcolormodel{cmyk}
+
+\usepackage{tikz, pgfplots}
+\pgfplotsset{compat=1.17}
+\usetikzlibrary{calc}
+\usetikzlibrary{pgfplots.groupplots}
+\usetikzlibrary{plotmarks}
+
+\usepackage{ifthen}
+\usepackage{xargs}
+
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 
-\title{Conference Paper Title*\\
+    \title{A Comparison of Algorithms Playing EvoMan\\
     {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
             email address or ORCID}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
     }
@@
     \begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
+
     \end{abstract}
@@
     \begin{IEEEkeywords}
-component, formatting, style, styling, insert
+        game-playing agent, artificial intelligence, EvoMan, genetic algorithm, reinforcement learning,
+        q-learning, neuroevolution, particle swarm optimization, proximal policy optimization, ppo
     \end{IEEEkeywords}
 
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
+    \section{Introduction}\label{sec:introduction}
 
-\section{Ease of Use}
+    \section{Problem Description}\label{sec:problem-description}
 
-\subsection{Maintaining the Integrity of the Specifications}
+    \subsection{Environment}\label{subsec:environment}
 
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
-
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
-
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
+    EvoMan~\cite{karinemiras,evoman} is a framework for testing competitive game-playing agents.
+    This framework is inspired by Mega Man II~\cite{capcom}, the game created by Capcom.
+    In the original game, the player would have to beat $8$ opponents and acquire their weapons
+    as they are defeated.
+    The additional difficulty of EvoMan comes from the fact the player has to defeat all
+    the opponents using only the starting weapon.
+    The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman\_framework}}
+    and it is currently compatible with Python 3.6 and 3.7.
+    There is also an extensive documentation
+    available\footnote{\url{https://github.com/karinemiras/evoman\_framework/blob/master/evoman1.0-doc.pdf}}.
+    The agent will collect information about the environment through 20 sensors (Fig 1):
     \begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
+        \item 16 correspond to horizontal and vertical distances to a maximum of 8 different opponent projectiles.
+        \item 2 correspond to the horizontal and vertical distance to the enemy.
+        \item 2 describe the directions the player and the enemy is facing
+    \end{itemize}
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.5\textwidth]{images/Evoman3.png}
+        \caption{Sensors available for the player.}
+        \label{fig:sensors}
+    \end{figure}
+    The actions which the agent may take are:
+    \begin{itemize}
+        \item walk left
+        \item walk right
+        \item jump
+        \item shoot
+        \item release of the jump
     \end{itemize}
 
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
+    The lives of the player and the enemy start at 100.
+    Everytime one of them gets hit, their life deplenishes.
+    Whoever's life reaches 0 loses the game.
 
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
+    \subsection{Problem}\label{subsec:problem}
+    The specific problem we are trying to solve is to train an agent on only four of the eight enemies with the purpose
+    of maximizing the score gain evaluated against all opponents.
+    Since the enemies are very different we should aim for the agent to find generally good strategies like
+    aiming towards the opponent and avoiding projectiles.
+    The gain of an agent after a game is computed as:
+    \begin{gather*}
+    gain = 100.01 + player\_energy - enemy\_energy
+    \end{gather*}
+    The final score metric of an agent is the harmonic mean of the gains against all enemies.
 
-\subsection{\LaTeX-Specific Advice}
+    \subsection{Result Upper Bound}\label{subsec:result-upper-bound}
+    (todo: cite https://arxiv.org/pdf/1912.10445.pdf)
+    The author of the EvoMan framework trained multiple specialized agents against all opponents.
+    The highest final gain, 185.67, was obtained with the NEAT(todo: cite) algorithm.
 
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
+    \section{Approach}\label{sec:approach}
 
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
+    \subsection{Q-Learning}\label{subsec:q-learning}
 
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-paper in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
+    \subsection{Neuroevolution with Genetic Algorithms}\label{subsec:neuroevolution-with-genetic-algorithms}
 
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
+    \subsection{Neuroevolution with Particle Swarm Optimization}\label{subsec:neuroevolution-with-particle-swarm-optimization}
 
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
+    \subsection{Proximal Policy Optimization}\label{subsec:proximal-policy-optimization}
 
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
+    \subsection{Neuroevolution with Particle Swarm Optimization Bootstrapping and Proximal Policy Optimization}\label{subsec:neuroevolution-with-particle-swarm-optimization-bootstrapping-and-proximal-policy-optimization}
 
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
+    \section{Experimental Investigation}\label{sec:experimental-investigation}
+    cautare 2d - algoritm si puncte de start.
+    rezultate partiale si deciziile luate dupa fiecare.
 
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
+    \section{Results}\label{sec:results}
+    informatii despre laptop (pentru ca trebuie mentionat timpul de rulare).
+    obs: pentru 1, 6, 7 am obtinut gain mai mare decat orice algoritm specializat (din upper bound)
 
@@
     \end{figure}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
 

### 2020-06-13 evoman_framework/competition_results.py

@@ -89,7 +89,7 @@ if mode == "test":
 	winners_slain["time"] = 3000 - winners_slain["time"]
 	winners_gain["time"] = 3000 - winners_gain["time"]
 	pd.concat([winners_slain, winners_gain], axis=1).to_csv("winners.csv")
-	# Prepare data for radar chart and make plots of winners and whole class
+	# Prepare data for radar chart and make data_for_plots of winners and whole class
 	# adapted from: https://python-graph-gallery.com/391-radar-chart-with-several-individuals/
 	for winner in [winners_slain["group"], winners_gain["group"], ["whole_class"]]:                         # !!!! [:3]
 		for group in winner:

### 2020-06-14 paper/paper.tex
sync
paper sync

@@
 \IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
 \usepackage{cite}
@@
 
-    \title{A Comparison of Algorithms Playing EvoMan\\
-    {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-    should not be used}
-        \thanks{Identify applicable funding agency here. If none, delete this.}
-    }
+    \title{A Comparison of Algorithms Playing EvoMan}
 
@@
     \begin{abstract}
-
+        Abstract is last.
     \end{abstract}
@@
     \section{Introduction}\label{sec:introduction}
+    Introduction is also last.
 
@@
     This framework is inspired by Mega Man II~\cite{capcom}, the game created by Capcom.
-    In the original game, the player would have to beat $8$ opponents and acquire their weapons
-    as they are defeated.
-    The additional difficulty of EvoMan comes from the fact the player has to defeat all
-    the opponents using only the starting weapon.
-    The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman\_framework}}
-    and it is currently compatible with Python 3.6 and 3.7.
-    There is also an extensive documentation
-    available\footnote{\url{https://github.com/karinemiras/evoman\_framework/blob/master/evoman1.0-doc.pdf}}.
+    EvoMan is a 2D shooting game where the player controls an agent playing against an opponent.
     The agent will collect information about the environment through 20 sensors (Fig 1):
@@
 
+    In the original game the player would have to beat $8$ opponents and acquire their weapons
+    as they are defeated.
+    The additional difficulty of EvoMan comes from the fact that the player has to defeat all
+    the opponents using only the starting weapon.
+    The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman\_framework}}
+    and it is currently compatible with Python 3.6 and 3.7.
+    There is also an extensive documentation
+    available\footnote{\url{https://github.com/karinemiras/evoman\_framework/blob/master/evoman1.0-doc.pdf}}.
+
     \subsection{Problem}\label{subsec:problem}
-    The specific problem we are trying to solve is to train an agent on only four of the eight enemies with the purpose
-    of maximizing the score gain evaluated against all opponents.
-    Since the enemies are very different we should aim for the agent to find generally good strategies like
-    aiming towards the opponent and avoiding projectiles.
-    The gain of an agent after a game is computed as:
+    Beating an opponent is relatively easy when using specialized models against a specific enemy.
+    The authors of the EvoMan framework trained multiple specialized agents against all opponents.
+    After playing a game, the formula for the gain by which an agent is evaluated is:
+    \boldmath
     \begin{gather*}
-    gain = 100.01 + player\_energy - enemy\_energy
+        gain = 100.01 + player\_life - enemy\_life
     \end{gather*}
+    \unboldmath
+    The highest final gain is 185.67\cite{evoman} and it was obtained with the NEAT\cite{neat} algorithm. \\
+
+    The problem we are trying to solve is a generalization of the one above.
+    We are to train an agent against four enemies and then test its performance against all eight.
     The final score metric of an agent is the harmonic mean of the gains against all enemies.
-
-    \subsection{Result Upper Bound}\label{subsec:result-upper-bound}
-    (todo: cite https://arxiv.org/pdf/1912.10445.pdf)
-    The author of the EvoMan framework trained multiple specialized agents against all opponents.
-    The highest final gain, 185.67, was obtained with the NEAT(todo: cite) algorithm.
+    The combination of four enemies we are to train the agent is not fixed.
 
     \section{Approach}\label{sec:approach}
+    Our intention was to develop a cascade ensemble method where the resulting model of an algorithm
+    which has a high exploratory bias is the starting point of an algorithm with high exploitative bias.
+    The exploratory algorithms we have considered are Q-Learning\cite{q_learning},
+    genetic algorithms\cite{genetic_algorithm} and particle swarm optimization\cite{pso}.
+    The exploitative algorithm we used was Proximal Policy Optimization\cite{ppo}.
+
+    To find the best exploratory algorithm we have made the following changes:
+    \begin{itemize}
+        \item we have increased the difficulty from the default 2 to 5.
+        \item the evaluation is made only on the second opponent due to the varied environment.
+        \item the gain function was modified as in Karine Miras' analysis\cite{evoman_blog} to:
+        \boldmath
+        \begin{gather*}
+            0.9 \cdot (100 - enemy\_life) + 0.1 \cdot player\_life\\
+            -\hspace{0.2cm} ln(nr\_of\_game\_timesteps)\\
+        \end{gather*}
+        \unboldmath
+    \end{itemize}
+
+    The best exploratory algorithm would be trained on four enemies and
+    used in cascade with proximal policy optimization without the changes mentioned in this section.
 
     \subsection{Q-Learning}\label{subsec:q-learning}
+    The first algorithm is a classic Q-learning\cite{q_learning} algorithm with neural networks.
 
-    \subsection{Neuroevolution with Genetic Algorithms}\label{subsec:neuroevolution-with-genetic-algorithms}
+    We used a neural network to predict the reward function for each possible move
+    (left, right, shoot, jump, release of jump) from a given state.
+    We then take the action with the highest predicted reward.
+    The input of this neural network is composed of the current game sensors and the
+    previous 2 game sensors with the moves taken at that specific point.
 
-    \subsection{Neuroevolution with Particle Swarm Optimization}\label{subsec:neuroevolution-with-particle-swarm-optimization}
+    The neural network used for experiments has 2 hidden layers with 32 neurons each.
+    Each layer has l2 regularization applied with a weight decay of 0.01 and sigmoid activation.
+    After each predicted move, we updated the neural network using backpropagation,
+    using as input the game sensors and as output the true reward function.
+    A game ends when either the agent or the enemy lose all life.
+    We have trained the agent on 5000 games.
+    The average number of frames per game for the best model is (todo: count in game steps).
+
+    \subsection{Evolving Neural Network Weights with Genetic Algorithms}\label{subsec:evolving-neural-network-weights-with-genetic-algorithms}
+    We used genetic algorithms to evolve the weights of neural networks with fixed structures.
+
+    \subsubsection{Sparse Reward Genetic Algorithm}
+    The second algorithm is a sparse reward neuroevolution\cite{neuro}.
+    The reward is defined as "sparse" because an agent doesn't find out how well it's doing until
+    the end of the game, with no feedback during the game.
+    An individual is represented as the weights of a neural network.
+
+    The neural network role is to predict the next move using the current game sensors and
+    the previous 2 game sensors with the moves taken at that specific point.
+    We used 2 hidden layers with 32 neurons for each individual.
+
+    We start with randomly initialized neural network weights and then represent them as a bitstring.
+    The weights used for the neural network were values between 2 and -2 with a precision of 6 digits.
+
+    For evaluation, the bitstring is transformed into the weights of the neural network and a game is played,
+    from which we can obtain the individual fitness.
+
+    Since we are representing the individuals as bitstrings we were able to apply
+    a simple genetic algorithm\cite{genetic_algorithm}.
+
+    The configurations we have used for the genetic algorithm is:
+    \begin{itemize}
+        \item population size: 50
+        \item number of generation: 500
+        \item crossover rate: 0.7
+        \item mutation rate: \{0.008, 0.1\}
+        \item elitism: 1
+    \end{itemize}
+    We have tried two experiments with different mutation rates in order to observe whether a very high
+    mutation rate can lead to good results for the problem. \\
+
+    The solution of the genetic algorithm is the best individual from the last generation.
+    Since we have used an elitism of 1, this means that the solution of the genetic algorithm is
+    the best individual ever evaluated.
+
+    \subsubsection{Iterative Genetic Algorithm}
+    The next algorithm is an iterative neuroevolution.
+    It is "iterative" because the agents are trained on a small number of game steps
+    first and then the number of game steps slowly increases.
+
+    After a number of generations we increase the number of game timesteps the agents are allowed to train on.
+    The fitness function was scaled based on the number of game timesteps in a way that if an agent
+    training on x game timesteps will always have a fitness lower than an agent training on y game
+    timesteps if x is lower than y.
+
+    \subsection{Searching Neural Network Weights with Particle Swarm Optimization}\label{subsec:searching-neural-network-weights-with-particle-swarm-optimization}
+    We have used particle swarm optimization\cite{pso} to search for the weights of a neural network of a
+    fixed structure which maximize the fitness defined in this section.
+    We used the same neural network configuration as in the case of genetic algorithms. \\
+
+    The configurations used for the particle swarm optimization\cite{pso} algorithm are:
+    \begin{itemize}
+        \item population size: 30
+        \item number of iterations: 200
+        \item cognitive weight: \{0.4, 0.8, 1.5\}
+        \item social weight: \{0.8, 0.4, 3\}
+        \item inertia weight: \{constant 1, uniformly decreasing from 1 to 0.3 (decreasing 0.0035 every iteration)\}
+    \end{itemize}
+
+    The same separation between an sparse and an iterative search was made in the case
+    of particle swarm optimization, as in the case of the genetic algorithms.
+
+    We also tried to search with pso using an uniformly decreasing inertia weight with respect
+    to the percent of the iterations passed, having values from 1 to 0.3.
+    This was tried to make the PSO focus even more on exploitation, rather than exploration.
 
     \subsection{Proximal Policy Optimization}\label{subsec:proximal-policy-optimization}
+    PPO\cite{ppo} is the algorithm we have used for exploitation.
+    For this stage we have used the default difficulty of 2, the training was done on 4 opponents
+    and the evaluation was done as described in the Problem Description(\ref{sec:problem-description}).
 
-    \subsection{Neuroevolution with Particle Swarm Optimization Bootstrapping and Proximal Policy Optimization}\label{subsec:neuroevolution-with-particle-swarm-optimization-bootstrapping-and-proximal-policy-optimization}
+    The neural network is randomly initialized. \\
+    The configuration we have used for the PPO is:
+    \begin{itemize}
+        \item hidden layers: (64, 64)
+        \item steps per epoch: 10000
+        \item epochs: 3000
+        \item gamma: 0.99
+        \item clip\_ratio: 0.2
+        \item pi\_lr: 3e-4
+        \item vf\_lr: 1e-3
+        \item train\_pi\_iterations: 80
+        \item train\_v\_iterations: 80
+        \item lambda: 0.97
+        \item target\_KL: 0.01
+    \end{itemize}
+
+    \subsection{Particle Swarm Optimization Cascaded With Proximal Policy Optimization}\label{subsec:particle-swarm-optimization-cascaded-with-proximal-policy-optimization}
+    The output weights of a neural network resulted from a PSO search is the starting agent for PPO\@.
+    In order to solve the generalized problem we have made the following changes from the PSO from the exploratory stage:
+    \begin{itemize}
+        \item The sizes of the hidden layers were increased from (32, 32) to (64, 64).
+        \item It is trained on four opponents, not only one.
+        \item The fitness function is the one from the Problem Description (\ref{sec:problem-description}).
+    \end{itemize}
+    The same configuration was used for PPO as in the case of the PPO with random
+    weights initialization.
 
     \section{Experimental Investigation}\label{sec:experimental-investigation}
-    cautare 2d - algoritm si puncte de start.
-    rezultate partiale si deciziile luate dupa fiecare.
+
+    \subsection{Q-Learning}\label{subsec:q-learning2}
+    The Q-learning algorithm was the starting point of our comparison.
+    Even when trained and evaluated against the same opponent, it would lose every game
+    while inflicting almost no damage.
+
+    \subsection{Genetic Algorithms}\label{subsec:genetic-algorithms}
+    The iterative genetic algorithm leads to much better results than Q-learning (Fig.~\ref{fig:q_vs_ga_iterative}).
+    \begin{figure}[!h]
+        \centering
+        \includegraphics[width=0.5\textwidth]{old_images/q_vs_ga_iterative.eps}
+        \caption{The fitness comparison between Q-learning and iterative genetic algorithms.}
+        \label{fig:q_vs_ga_iterative}
+    \end{figure}
+
+    We have shown that a mutation rate of 0.008 leads to better results than a mutation rate of 0.1 (Fig.~\ref{fig:ga_mutation_rates}).
+    \begin{figure}[!h]
+        \centering
+        \includegraphics[width=0.5\textwidth]{old_images/ga_mutation_rates.eps}
+        \caption{The fitness comparison between sparse genetic algorithms with mutation rate of 0.1 and 0.008.}
+        \label{fig:ga_mutation_rates}
+    \end{figure}
+
+    The results of the sparse genetic algorithm and the iterative genetic algorithm
+    are not significantly different (Fig.~\ref{fig:overview}).
+
+    \subsection{Particle Swarm Optimization}\label{subsec:sparse-particle-swarm-optimization}
+    Both particle swarm optimization algorithms lead to much better results than either of the
+    sparse and iterative genetic algorithms.
+    The iterative PSO lead to worse results than the sparse PSO\@.
+    In Fig.~\ref{fig:overview} we can see that the best exploratory algorithm is the sparse PSO\@.
+    \begin{figure}[!h]
+        \centering
+        \includegraphics[width=0.5\textwidth]{old_images/overview.eps}
+        \caption{The fitness comparison between the iterative and the sparse approached,
+        both for the genetic algorithms and the PSO. }
+        \label{fig:overview}
+    \end{figure}
+
+    The time advantage of iterative PSO vs sparse PSO is not significant (Fig.~\ref{fig:pso_sparse_vs_iterative_time}).
+    \begin{figure}[!h]
+        \centering
+        \includegraphics[width=0.5\textwidth]{old_images/pso_sparse_vs_iterative_time.eps}
+        \caption{The time comparison between sparse and iterative PSO.}
+        \label{fig:pso_sparse_vs_iterative_time}
+    \end{figure}
+
+    We have also searched for good PSO weights for our problem.
+    In the next stages we have used a cognitive weight of 0.4 and a social weight of 0.8 due to
+    its higher variance in results (Fig.~\ref{fig:pso_configs}).
+
+    \begin{figure}[!h]
+        \centering
+        \includegraphics[width=0.5\textwidth]{old_images/pso_configs.eps}
+        \caption{The fitness comparison between sparse PSO with different cognitive and social weights.}
+        \label{fig:pso_configs}
+    \end{figure}
+
+    We have also tried two PSO inertia weight updates (Fig.~\ref{fig:pso_inertia}).
+    We continued with the constant inertia weight approach.
+
+    \begin{figure}[!h]
+        \centering
+        \includegraphics[width=0.5\textwidth]{old_images/pso_inertia.eps}
+        \caption{The fitness comparison between PSO with constant inertia
+        weight and with uniformly decreasing in time inertia weight.}
+        \label{fig:pso_inertia}
+    \end{figure}
+
+    After deciding what the starting model should be, we have tested its performance (Fig.~\ref{fig:pso_levels})
+    and run time (Fig.~\ref{fig:pso_levels_time}) on multiple difficulties.
+
+    \begin{figure}[!h]
+        \centering
+        \includegraphics[width=0.5\textwidth]{old_images/pso_levels.eps}
+        \caption{The fitness comparison of PSO against different game difficulty levels.}
+        \label{fig:pso_levels}
+    \end{figure}
+
+    \begin{figure}[!h]
+        \centering
+        \includegraphics[width=0.5\textwidth]{old_images/pso_levels_time.eps}
+        \caption{The time comparison of PSO at different game difficulty levels.}
+        \label{fig:pso_levels_time}
+    \end{figure}
 
     \section{Results}\label{sec:results}
-    informatii despre laptop (pentru ca trebuie mentionat timpul de rulare).
-    obs: pentru 1, 6, 7 am obtinut gain mai mare decat orice algoritm specializat (din upper bound)
+    For evaluation of an agent, we ran 30 games against each opponent, leading to 8 averages
+    (one per opponent), of which we computed the harmonic mean which is the final result of the agent.
 
+    \subsection{Best Combination of Opponents for Training}\label{subsec:best-combination-of-opponents-for-training}
+    We searched for the opponents that lead to the best results.
+    The first combination of train opponents we considered was \{1, 2, 6, 7\} because (todo: explain).
+    We observed its final gain as 125.37.
+    We realized a set of exploratory experiments where we would search for other combinations of train opponents,
+    but starting with the resulting model of PPO trained for 1000 epochs on the combination of opponents \{1, 2, 6, 7\}.
+    If by starting with a pre-trained model another combination of train enemies does not lead to better results,
+    then we can believe that the respective combination does not lead to better results than \{1, 2, 6, 7\}.
     \begin{table}[htbp]
-        \caption{Table Type Styles}
+        \caption{Various train opponents for pre-trained PPO (2000 epochs)}
         \begin{center}
-            \begin{tabular}{|c|c|c|c|}
+            \begin{tabular}{|c|c|}
                 \hline
-                \textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
+                \textbf{Train opponents}&{\textbf{gain (harmonic mean)}} \\
                 \cline{2-4}
-                \textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
                 \hline
-                copy& More table copy$^{\mathrm{a}}$& &  \\
+                \{3, 4, 5, 8\} & 44.16 \\
+                \{3, 4, 6, 7\} & 75.3 \\
+                \{1, 3, 6, 7\} & 110.19 \\
+                \{2, 3, 6, 7\} & 136.1 \\
+                \{2, 4, 6, 7\} & 85.22 \\
                 \hline
-                \multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
             \end{tabular}
-            \label{tab1}
+            \label{table:variations_of_opponents}
+        \end{center}
+    \end{table}
+    \\
+    The only combination that lead to a better score than 125.37 was \{2, 3, 6, 7\}.
+    We ran an experiment with PPO with random initialization with the combination of train enemies \{2, 3, 6, 7\} and
+    we obtained a final gain of 99.97.
+    This means that we did not find a better combination of train opponents than \{1, 2, 6, 7\}, so it is the one
+    we used for the following experiments.
+
+    \subsection{Random Initialization PPO vs PSO Cascading PPO}\label{subsec:random-initialization-ppo-vs-pso-cascading-ppo}
+    The best PSO configuration was used in cascade before the PPO in 4 runs.
+    PPO with random initialization was ran 3 times.
+    The worst result of the PPO with random initialization is higher than the best result
+    of PSO cascaded with PPO\@ (Fig.~\ref{fig:random_vs_pso_initialization}).
+    \begin{figure}[htbp]
+        \centering
+        \includegraphics[width=0.5\textwidth]{images/random_vs_pso_initialization.eps}
+        \caption{Gain comparison between PPO with random initialization and with PSO cascading.}
+        \label{fig:random_vs_pso_initialization}
+    \end{figure}
+
+    \subsection{PPO 2000 epochs vs PPO 3000 epochs}\label{subsec:ppo-2000-epochs-vs-ppo-3000-epochs}
+    The final gain is greater after 2000 epochs than after 3000 epochs (Fig.~\ref{fig:ppo_2000_vs_3000_epochs}).
+    This means that there is overfitting during the training.
+    We can conclude that by stopping the PPO algorithm after 2000 epochs instead of 3000 epochs
+    we can end up with more generalized agents that perform better on average against all opponents.
+    \begin{figure}[htbp]
+        \centering
+        \includegraphics[width=0.5\textwidth]{images/ppo_2000_vs_3000_epochs.eps}
+        \caption{Gain comparison between PPO after 2000 epochs and after 3000 epochs.}
+        \label{fig:ppo_2000_vs_3000_epochs}
+    \end{figure}
+
+    \begin{figure}[htbp]
+        \centering
+        \includegraphics[width=0.5\textwidth]{images/ppo_gain_evolution_3_runs.eps}
+        \caption{Gain Evolution for PPO in 3 runs.}
+        \label{fig:ppo_gain_evolution_3_runs}
+    \end{figure}
+
+    \subsection{Best Tested Agent}\label{subsec:best-tested-agent}
+    The highest gain for a tested agent is 137.18.
+    It was obtained after 1750 epochs of one of the PPO runs with random initialization.
+    Only against opponents 3 and 4 this agent sometimes loses.
+    \begin{table}[htbp]
+        \caption{Best Agent Against All Opponents}
+        \begin{center}
+            \begin{tabular}{|c|c|}
+                \hline
+                \textbf{Opponent}&{\textbf{Average gain (30 games)}} \\
+                \hline
+                1 & 199.61 \\
+                2 & 189.143 \\
+                3 & 85.01 \\
+                4 & 80.17 \\
+                5 & 158.83 \\
+                6 & 194.37 \\
+                7 & 191.17 \\
+                8 & 140.61 \\
+                \hline
+            \end{tabular}
         \end{center}
@@
 
-    \begin{figure}[htbp]
-        \centerline{\includegraphics{fig1.png}}
-        \caption{Example of a figure caption.}
-        \label{fig}
-    \end{figure}
+    \subsection{Comparison with the Upper Bound}\label{subsec:comparison-with-the-upper-bound}
+    With the PPO algorithm trained on the opponents \{1, 2, 6, 7\} we have
+    beat the best specialized models from the original paper which were given as upper bounds\cite{evoman}.
+    NEAT specialized
+   \begin{table}[htbp]
+        \caption{Specialized NEAT vs Generalized PPO on train opponents}
+        \begin{center}
+            \begin{tabular}{|c|c|c|}
+                \hline
+                \textbf{Opponent}&{\textbf{Specialized NEAT}}&{\textbf{Generalized PPO}} \\
+                \cline{2-4}
+                \hline
+                1 & 190.01 & 199.81 \\
+                2 & 194.01 & 190.34 \\
+                6 & 173.01 & 195.85 \\
+                7 & 177.01 & 193.27 \\
+                \hline
+                \multicolumn{3}{l}{The generalized PPO was trained on the opponents \{1, 2, 6, 7\}}
+            \end{tabular}
+        \end{center}
+    \end{table}
+
+    \begin{table}[htbp]
+        \caption{Specialized NEAT vs Generalized PPO}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|c|c|}
+                \hline
+                \textbf{Opponent}&{\textbf{NEAT}}&{\textbf{PPO Run 1}}&{\textbf{PPO Run 2}}&{\textbf{PPO Run 3}}&{\textbf{PPO Best}} \\
+                \cline{2-4}
+                \hline
+                1 & 190.01 & 198.41 & 199.07 & 199.81 & 199.61 \\
+                2 & 194.01 & 199.74 & 199.27 & 190.34 & 189.14 \\
+                3 & 180.01 & 58.94  &  46.67 &  70.27 & 85.01 \\
+                4 & 194.01 & 58.34  &   66.5 &  64.11 & 80.17 \\
+                5 & 194.01 & 172.61 & 165.65 & 174.61 & 158.83 \\
+                6 & 173.01 & 195.73 & 196.95 & 195.85 & 194.37 \\
+                7 & 177.01 & 199.79 & 195.05 & 193.27 & 191.17 \\
+                8 & 186.01 & 122.17 & 145.49 & 141.89 & 140.61 \\
+                \hline
+            \end{tabular}
+        \end{center}
+    \end{table}
+
+    \subsection{Time Analysis}\label{subsec:time-analysis}
+    Since the match time is relevant, the machine on which the games were run is also relevant.
+    The tests were run on an i7 4750HQ processor.
+    The training was done on 3 threads.
+    All computations were done on the CPU\@.
+    (todo: detalii despre timpul de rulare (din fisierele model\_name.csv))
 
     \begin{thebibliography}{00}
-        \bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-        \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-        \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-        \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-        \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-        \bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-        \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
+        \bibitem{evoman} Fabricio Olivetti de Franca, Denis Fantinato, Karine Miras, A.E. Eiben and Patricia A. Vargas.
+        "EvoMan: Game-playing Competition" arXiv:1912.10445
+        \bibitem{karinemiras} de Araújo, Karine da Silva Miras, and Fabrício Olivetti de França.
+        "An electronic-game framework for evaluating coevolutionary algorithms." arXiv:1604.00644 (2016).
+        \bibitem{neuro} Floreano, D., Dürr, P. & Mattiussi, C. Neuroevolution: from architectures to learning. Evol. Intel. 1, 47–62 (2008). https://doi.org/10.1007/s12065-007-0002-4
+        \bibitem{capcom} M. MEGA, "Produced by capcom, distributed by capcom, 1987," System: NES.
+        \bibitem{q_learning} Watkins, C.J.C.H., Dayan, P. Q-learning. Machine Learning 8, 279–292 (1992)
+        \bibitem{genetic_algorithm} Holland J.H., Genetic Algorithms and Adaptation. Adaptive Control of Ill-Defined Systems, 1984, Volume 16 ISBN 978-1-4684-8943-9
+        \bibitem{pso} Kennedy, J.; Eberhart, R. (1995). "Particle Swarm Optimization". Proceedings of IEEE International Conference on Neural Networks. IV. pp. 1942–1948.
+        \bibitem{neat} Kenneth O.
+Stanly; Rist Miikkulainen (2002).
+"Evolving Neural Networks through Augmenting Topologies".
+Evolutionary Computation, Volume 10, Issue 2, Summer 2002, p.99-127, https://doi.org/10.1162/106365602320169811
+        \bibitem{ppo} John Schulman, Filip Wolski, Prafulla Dhariwal, Alex Radford, Oleg Klimov (2017) "Proximal Policy Optimization Algorithms", arXiv:1707.06347v2
+        \bibitem{evoman_blog} Karine Miras, Evoman, https://karinemirasblog.wordpress.com/portfolio/evoman/
     \end{thebibliography}
-    \vspace{12pt}
-    \color{red}
-    IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
 

### 2020-06-14 code/test_models/reinforcement_learning/main.py
fixed average duration for testing model

@@ -70,7 +70,7 @@ for enemy in parameters.ENEMIES_CHOSEN_FOR_TESTING:
 
         average_player_life += player_life
         average_enemy_life += enemy_life
-        average_time += average_time
+        average_time += time
 
     average_player_life /= parameters.NR_EXPERIMENTS_FOR_EACH_ENEMY
     average_enemy_life /= parameters.NR_EXPERIMENTS_FOR_EACH_ENEMY

### 2020-06-14 code/test_models/reinforcement_learning/test_multiple.py

@@ -72,7 +72,7 @@ def test_model(model_name, ENEMIES_CHOSEN_FOR_TESTING=range(1, 9), NR_EXPERIMENT
 
             average_player_life += player_life
             average_enemy_life += enemy_life
-            average_time += average_time
+            average_time += time
 
         average_player_life /= NR_EXPERIMENTS_FOR_EACH_ENEMY
         average_enemy_life /= NR_EXPERIMENTS_FOR_EACH_ENEMY

### 2020-06-14 paper/paper.tex
Included url tag, citation to image and added info about time.

@@
 \usepackage{xcolor}
+\usepackage{hyperref}
 
@@
         \includegraphics[width=0.5\textwidth]{images/Evoman3.png}
-        \caption{Sensors available for the player.}
+        \caption{Sensors available for the player\cite{evoman}.}
         \label{fig:sensors}
@@
     the opponents using only the starting weapon.
-    The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman\_framework}}
+    The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman_framework}}
     and it is currently compatible with Python 3.6 and 3.7.
     There is also an extensive documentation
-    available\footnote{\url{https://github.com/karinemiras/evoman\_framework/blob/master/evoman1.0-doc.pdf}}.
+    available\footnote{\url{https://github.com/karinemiras/evoman_framework/blob/master/evoman1.0-doc.pdf}}.
 
@@
     All computations were done on the CPU\@.
-    (todo: detalii despre timpul de rulare (din fisierele model\_name.csv))
+    The median game time during training is 28 seconds.
+    (todo: include median nr of epochs at testing)
 
@@
         "An electronic-game framework for evaluating coevolutionary algorithms." arXiv:1604.00644 (2016).
-        \bibitem{neuro} Floreano, D., Dürr, P. & Mattiussi, C. Neuroevolution: from architectures to learning. Evol. Intel. 1, 47–62 (2008). https://doi.org/10.1007/s12065-007-0002-4
+        \bibitem{neuro} Floreano, D., Dürr, P. & Mattiussi, C. Neuroevolution: from architectures to learning. Evol. Intel. 1, 47–62 (2008). \url{https://doi.org/10.1007/s12065-007-0002-4}
         \bibitem{capcom} M. MEGA, "Produced by capcom, distributed by capcom, 1987," System: NES.
@@
 "Evolving Neural Networks through Augmenting Topologies".
-Evolutionary Computation, Volume 10, Issue 2, Summer 2002, p.99-127, https://doi.org/10.1162/106365602320169811
+Evolutionary Computation, Volume 10, Issue 2, Summer 2002, p.99-127, \url{https://doi.org/10.1162/106365602320169811}
         \bibitem{ppo} John Schulman, Filip Wolski, Prafulla Dhariwal, Alex Radford, Oleg Klimov (2017) "Proximal Policy Optimization Algorithms", arXiv:1707.06347v2
-        \bibitem{evoman_blog} Karine Miras, Evoman, https://karinemirasblog.wordpress.com/portfolio/evoman/
+        \bibitem{evoman_blog} Karine Miras, Evoman, \url{https://karinemirasblog.wordpress.com/portfolio/evoman/}
     \end{thebibliography}

### 2020-06-15 paper/paper.tex
paper changes according to discussion

@@
 \usepackage{tikz, pgfplots}
-\pgfplotsset{compat=1.17}
 \usetikzlibrary{calc}
@@
 
-    In the original game the player would have to beat $8$ opponents and acquire their weapons
+    In the original Capcom game the player would have to beat $8$ opponents and acquire their weapons
     as they are defeated.
@@
     the opponents using only the starting weapon.
+    Each opponent can be fought on a specified difficulty level.
+The difficulty level is an integer greater or equal than $1$ which is translated into a factor for the damage taken and damage given by the  player, the higher the difficulty level the lower the damage given and the higher the damage taken.
     The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman_framework}}
@@
     The combination of four enemies we are to train the agent is not fixed.
+    The difficulty level for solving this problem is $2$ for all enemies in training and testing, but we did some exploratory work with greater level of difficulties outside training and testing for obtaining the final solution for the problem.
+
+    The specifications above, like the number of enemies chosen for training or the difficulty level, come directly from a competition organized by the creators of the framework\cite{evoman_competition}.
 
@@
     Our intention was to develop a cascade ensemble method where the resulting model of an algorithm
-    which has a high exploratory bias is the starting point of an algorithm with high exploitative bias.
-    The exploratory algorithms we have considered are Q-Learning\cite{q_learning},
-    genetic algorithms\cite{genetic_algorithm} and particle swarm optimization\cite{pso}.
+    which can achieve a decent gain fast or has a high exploratory bias is the starting point of an algorithm with high exploitative bias.
+    The algorithm we expected to obtain a decent gain fast is a classic Q-learning\cite{q_learning} algorithm.
+The exploratory algorithms we have considered are genetic algorithms\cite{genetic_algorithm} and particle swarm optimization\cite{pso}.
     The exploitative algorithm we used was Proximal Policy Optimization\cite{ppo}.
 
-    To find the best exploratory algorithm we have made the following changes:
+    To find the best bootstraping algorithm we have made the following changes, which had no impact in the training and testing for the final solution to the problem:
     \begin{itemize}
@@
     \subsection{Q-Learning}\label{subsec:q-learning}
-    The first algorithm is a classic Q-learning\cite{q_learning} algorithm with neural networks.
+    The algorithm we expected to obtain a decent gain fast is a classic Q-learning\cite{q_learning} algorithm with neural networks.
 
@@
     We have trained the agent on 5000 games.
-    The average number of frames per game for the best model is (todo: count in game steps).
+    The average number of frames per game for the best model is 287.
 
@@
         \item social weight: \{0.8, 0.4, 3\}
-        \item inertia weight: \{constant 1, uniformly decreasing from 1 to 0.3 (decreasing 0.0035 every iteration)\}
+        \item inertia weight: \{constant 1, uniformly decreasing from 1 to 0.3\}
     \end{itemize}
@@
     to the percent of the iterations passed, having values from 1 to 0.3.
+The inertia weight controls the level of exploration versus exploitation, higher values in the beginning of the algorithm lead to better exploration while lower values towards the end of the algorithm lead to better exploitation.
     This was tried to make the PSO focus even more on exploitation, rather than exploration.
@@
     We searched for the opponents that lead to the best results.
-    The first combination of train opponents we considered was \{1, 2, 6, 7\} because (todo: explain).
+    The first combination of train opponents we considered was \{1, 2, 6, 7\} which was chosen empirically after manually playing against every opponent and chosing the ones which behaved in such a way that covered the most general concepts our agents could learn in our opinion.
     We observed its final gain as 125.37.
@@
     \begin{table}[htbp]
-        \caption{Various train opponents for pre-trained PPO (2000 epochs)}
+        \caption{Results for various opponents when starting with a pre-trained PPO model on enemies \{1, 2, 6, 7\}}
         \begin{center}
@@
                 \hline
-                \{3, 4, 5, 8\} & 44.16 \\
-                \{3, 4, 6, 7\} & 75.3 \\
-                \{1, 3, 6, 7\} & 110.19 \\
-                \{2, 3, 6, 7\} & 136.1 \\
-                \{2, 4, 6, 7\} & 85.22 \\
+                \{1, 2, 6, 7\} \cup \{3, 4, 5, 8\} & 44.16 \\
+                \{1, 2, 6, 7\} \cup \{3, 4, 6, 7\} & 75.3 \\
+                \{1, 2, 6, 7\} \cup \{1, 3, 6, 7\} & 110.19 \\
+                \{1, 2, 6, 7\} \cup \{2, 3, 6, 7\} & 136.1 \\
+                \{1, 2, 6, 7\} \cup \{2, 4, 6, 7\} & 85.22 \\
                 \hline
@@
     \end{figure}
-
-    \subsection{PPO 2000 epochs vs PPO 3000 epochs}\label{subsec:ppo-2000-epochs-vs-ppo-3000-epochs}
-    The final gain is greater after 2000 epochs than after 3000 epochs (Fig.~\ref{fig:ppo_2000_vs_3000_epochs}).
+	Our explanation for getting worse results when using the technique mentioned above is that the function landscape is very big and PSO doesn't explore enough.
+Another possible explanation would be that the landscape is tricky when it comes to searching for generalist versus specialized solutions.
+We could try solving this problem by giving more exploratory power to PSO, but we did not experiment with this due to the lack of time.
+    \subsection{PPO on the best configuration}\label{subsec:ppo-on-the-best-configuration}
+    Considering the results above, we decided that PPO with random initialization and enemies \{1, 2, 6, 7\} chosen for training would be the best configuration.
+We ran 3000 epochs of this configuration, saving all the models along the way after every 250 epochs.
+    After taking a look at the testing results, we noticed that the final gain is greater after 2000 epochs than after 3000 epochs (Fig.~\ref{fig:ppo_2000_vs_3000_epochs}).
     This means that there is overfitting during the training.
@@
         \includegraphics[width=0.5\textwidth]{images/ppo_gain_evolution_3_runs.eps}
-        \caption{Gain Evolution for PPO in 3 runs.}
+        \caption{Gain Evolution for PPO in 3 runs, each model was ran 30 times on each opponent for calculating the gains}
         \label{fig:ppo_gain_evolution_3_runs}
@@
         \bibitem{evoman_blog} Karine Miras, Evoman, \url{https://karinemirasblog.wordpress.com/portfolio/evoman/}
+        \bibitem{evoman_competition} Evoman: Game-playing Competition for WCCI 2020, \url{http://pesquisa.ufabc.edu.br/hal/Evoman.html}
     \end{thebibliography}

### 2020-06-15 code/test_models/reinforcement_learning/main.py
test script percentage of games lost

@@ -3,8 +3,6 @@
 import sys
 import os
 
-
-
 is_running_from_command_line = len(sys.path) <= 7
 if is_running_from_command_line:
     script_path_tokens = sys.path[0].split(os.sep)
@@ -52,12 +50,15 @@ for enemy in parameters.ENEMIES_CHOSEN_FOR_TESTING:
     min_player_life, max_player_life, average_player_life = 100, 0, 0
     min_enemy_life, max_enemy_life, average_enemy_life = 100, 0, 0
     min_time, max_time, average_time = 100000000, 0, 0
+    number_of_games_lost = 0
 
     for experiment in range(parameters.NR_EXPERIMENTS_FOR_EACH_ENEMY):
         evoman_environment = EvomanEnvironmentWrapper('evoman rl test',
                                                       player_controller=TestReinforcementLearningEvomanPlayerController(),
                                                       enemies=[enemy])
         _, player_life, enemy_life, time = evoman_environment.play(pcont=model)
+
+        number_of_games_lost += 1 if player_life == 0 else 0
 
         min_player_life = min(min_player_life, player_life)
         max_player_life = max(max_player_life, player_life)
@@ -75,23 +76,24 @@ for enemy in parameters.ENEMIES_CHOSEN_FOR_TESTING:
     average_player_life /= parameters.NR_EXPERIMENTS_FOR_EACH_ENEMY
     average_enemy_life /= parameters.NR_EXPERIMENTS_FOR_EACH_ENEMY
     average_time /= parameters.NR_EXPERIMENTS_FOR_EACH_ENEMY
+    percentage_of_games_lost = 100 * number_of_games_lost / parameters.NR_EXPERIMENTS_FOR_EACH_ENEMY
 
     gains.append(100.01 + average_player_life - average_enemy_life)
     results.append(
         [enemy, gains[-1], average_player_life, average_enemy_life, average_time, min_player_life, max_player_life,
-         min_enemy_life, max_enemy_life, min_time, max_time])
+         min_enemy_life, max_enemy_life, min_time, max_time, percentage_of_games_lost])
 
 save_csv_results(results)
 
 harmonic_mean_of_gains = len(gains) / np.sum(1.0 / np.array(gains))
 print(f'This model has a score for the competition of {harmonic_mean_of_gains:.2f}/200.01\n')
 
[…]

### 2020-06-15 code/test_models/reinforcement_learning/main.py
Updated results with percent of lost games.

@@ -37,7 +37,7 @@ def save_csv_results(results):
         csvwriter = csv.writer(csvfile, delimiter=',', quotechar='|', quoting=csv.QUOTE_MINIMAL)
         csvwriter.writerow(
             ['enemy', 'gain', 'avg_player_life', 'avg_enemy_life', 'avg_duration', 'min_player_life', 'max_player_life',
-             'min_enemy_life', 'max_enemy_life', 'min_duration', 'max_duration'])
+             'min_enemy_life', 'max_enemy_life', 'min_duration', 'max_duration', 'percentage_games_lost'])
         csvwriter.writerows(results)
 
 

### 2020-06-15 paper/paper.tex
PSO inertia weight updated set
Full table in tex

@@
         \item social weight: \{0.8, 0.4, 3\}
-        \item inertia weight: \{constant 1, uniformly decreasing from 1 to 0.3\}
+        \item inertia weight: \{constant 1, decreasing from 1 by 0.0035 every epoch\}
     \end{itemize}
@@
 We could try solving this problem by giving more exploratory power to PSO, but we did not experiment with this due to the lack of time.
+
     \subsection{PPO on the best configuration}\label{subsec:ppo-on-the-best-configuration}
@@
     beat the best specialized models from the original paper which were given as upper bounds\cite{evoman}.
-    NEAT specialized
     \begin{table}[htbp]
@@
 
+    \begin{table}[htbp]
+        \caption{Results}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|c|c|}
+                \hline
+                \textbf{Opponent}&\multicolumn{4}{|c|}{\textbf{PPO}}&\textbf{Specialized} \\
+                \cline{2-5}
+                & \textbf{\textit{Run 1}}& \textbf{\textit{Run 2}}& \multicolumn{2}{|c|}{\textbf{\textit{Run 3}}} & NEAT \\
+                \cline{4-5}
+                & 3000 ep & 3000 ep & 3000 ep & 1750 ep & \\
+                & (best at training) &  &  & (best tested) & \\
+                \hline
+                1 & 198.41 & 199.07 & 199.81 & 199.61 & 190.01 \\
+                \hline
+                2 & 199.74 & 199.27 & 190.34 & 189.14 & 194.01 \\
+                \hline
+                3 & 58.94 & 46.67 & 70.27 &  85.01 & 180.01 \\
+                \hline
+                4 & 58.34 & 66.5 & 64.11 & 80.17 & 194.01 \\
+                \hline
+                5 & 172.61 & 165.65 & 174.61 & 158.83 & 194.01\\
+                \hline
+                6 & 195.73 & 196.95 & 195.85 & 194.37 & 173.01 \\
+                \hline
+                7 & 199.79 & 195.05 & 193.27 & 191.17 & 177.01 \\
+                \hline
+                8 & 122.17 & 145.49 & 141.89 & 140.61 & 186.01 \\
+                \hline
+                \textbf{\{1, 2, 6, 7\}} &  &  &  & & \\
+                \textbf{harmonic mean} & 198.4 & 197.57 & 194.75 & 193.49 & 183.09 \\
+                \hline
+                \textbf{final} &  &  &  & & \\
+                \textbf{harmonic mean} & 117.22 & 114.91 & 127.01 & 137.18 & 185.67\\
+                \hline
+
+            \end{tabular}
+            \label{tab1}
+        \end{center}
+    \end{table}
+
     \subsection{Time Analysis}\label{subsec:time-analysis}
@@
     The median game time during training is 28 seconds.
-    (todo: include median nr of epochs at testing)
+    The average number of frames per game is 287.
 

### 2020-06-15 paper/paper.tex
sync.

@@
     \begin{abstract}
-        Abstract is last.
+        This paper describes a comparison between algorithms for evolving agents for the game Evoman.
+        We have tried a cascade ensemble method, starting with an algorithm that is either fast either has
+        an explorative bias and refining its solution with an exploitative algorithm.
+        The algorithms we have tested for the first stage are Q-learning, genetic algorithms and particle swarm optimization.
+        All of these algorithms are searching for the weights of a neural network of a fixed structure.
+        Both the genetic algorithms and the particle swarm optimization algorithms were tested with a sparse and
+        an iterative approach.
+        The best explorative algorithm was the sparse particle swarm optimization.
+        The exploitative algorithm we tested is proximal policy optimization.
+        Using PPO with random initialization lead to better results than an ensemble method of PSO cascaded with PPO\@.
+        We also detected overfitting of the agent trained with PPO\@.
     \end{abstract}
@@
     \section{Introduction}\label{sec:introduction}
-    Introduction is also last.
+    This paper presents our solution for the "Evoman: Game-playing Competition for WCCI 2020"\cite{evoman_competition}.
+    We are attempting to train an agent playing the 2D shooting game Evoman\cite{evoman}.
+    We have tried an ensemble cascade method with two stages.
+    For the first stage we tested algorithms that either found acceptable solutions fast either had a high
+    explorative bias.
+    The algorithms we considered for this stage are Q-Learning\cite{q_learning}, genetic algorithms\cite{genetic_algorithm}
+    and particle swarm optimization\cite{pso}.
+    Both the GA and the PSO train the weights of a neural network of a fixed structure.
+    Both the GA and the PSO were tested with a sparse approach and an iterative approach.
+    For the exploitative stage we used Proximal Policy Optimization\cite{ppo}.
 
@@
     Each opponent can be fought on a specified difficulty level.
-The difficulty level is an integer greater or equal than $1$ which is translated into a factor for the damage taken and damage given by the player, the higher the difficulty level the lower the damage given and the higher the damage taken.
+    The difficulty level is an integer greater or equal than $1$ which is translated into a factor for the damage
+    taken and damage given by the player, the higher the difficulty level the lower the damage given and the higher the damage taken.
     The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman_framework}}
@@
     The combination of four enemies we are to train the agent is not fixed.
-    The difficulty level for solving this problem is $2$ for all enemies in training and testing, but we did some exploratory work with greater level of difficulties outside training and testing for obtaining the final solution for the problem.
+    The difficulty level for solving this problem is $2$ for all enemies in training and testing,
+    but we did some exploratory work with greater level of difficulties outside training and
+    testing for obtaining the final solution for the problem.
 
-    The specifications above, like the number of enemies chosen for training or the difficulty level, come directly from a competition organized by the creators of the framework\cite{evoman_competition}.
+    The specifications above, like the number of enemies chosen for training or the difficulty level,
+    come directly from a competition organized by the creators of the framework\cite{evoman_competition}.
 
@@
     PPO with random initialization was ran 3 times.
+    The small number of runs is due to the long run time of the training.
     The worst result of the PPO with random initialization is higher than the best result
@@
 We ran 3000 epochs of this configuration, saving all the models along the way after every 250 epochs.
-    After taking a look at the testing results, we noticed that the final gain is greater after 2000 epochs than after 3000 epochs (Fig.~\ref{fig:ppo_2000_vs_3000_epochs}).
+    After looking at the testing results, we noticed that the final gain is greater after 2000 epochs than after 3000 epochs (Fig.~\ref{fig:ppo_2000_vs_3000_epochs}).
     This means that there is overfitting during the training.
@@
 
-    \subsection{Best Tested Agent}\label{subsec:best-tested-agent}
-    The highest gain for a tested agent is 137.18.
-    It was obtained after 1750 epochs of one of the PPO runs with random initialization.
-    Only against opponents 3 and 4 this agent sometimes loses.
-    \begin{table}[htbp]
-        \caption{Best Agent Against All Opponents}
-        \begin{center}
-            \begin{tabular}{|c|c|}
-                \hline
-                \textbf{Opponent}&{\textbf{Average gain (30 games)}} \\
-                \hline
-                1 & 199.61 \\
-                2 & 189.143 \\
-                3 & 85.01 \\
-                4 & 80.17 \\
-                5 & 158.83 \\
-                6 & 194.37 \\
-                7 & 191.17 \\
-                8 & 140.61 \\
-                \hline
-            \end{tabular}
-        \end{center}
-    \end{table}
+    \subsection{Best Train Agent vs Best Tested Agent}\label{subsec:best-tested-agent}
+    The best train gain was obtained in the first run of the PPO with random initialization.
+    The train gain was computed as the harmonic mean of the averages per opponent of the 30 games played against
+    the enemies \{1, 2, 6, 7\}.
+    The highest train gain is 198.4.
+    When computed against all opponents, this agent has an overall gain of 117.22.
+    Out of the three PPO with random initialization runs, the highest final gain is 127.01.
+    After the experiments were concluded we also looked at the overall gains of the agents before 3000 epochs (Fig.~\ref{fig:ppo_gain_evolution_3_runs}).
+    For the runs 2 and 3 the snapshots of the models were saved every 250 epochs.
+    In the first run these periodic snapshots start after 2000 epochs.
+    When looking at the results of the intermediary agents we observed that there are some high peaks in test gain.
+    The highest test gain was observed during the third run after 1750 epochs and it has a value of 137.18.
 
@@
     beat the best specialized models from the original paper which were given as upper bounds\cite{evoman}.
-    \begin{table}[htbp]
-        \caption{Specialized NEAT vs Generalized PPO on train opponents}
-        \begin{center}
-            \begin{tabular}{|c|c|c|}
-                \hline
-                \textbf{Opponent}&{\textbf{Specialized NEAT}}&{\textbf{Generalized PPO}} \\
-                \cline{2-4}
-                \hline
-                1 & 190.01 & 199.81 \\
-                2 & 194.01 & 190.34 \\
-                6 & 173.01 & 195.85 \\
-                7 & 177.01 & 193.27 \\
-                \hline
-                \multicolumn{3}{l}{The generalized PPO was trained on the opponents \{1, 2, 6, 7\}}
-            \end{tabular}
-        \end{center}
-    \end{table}
 
     \begin{table}[htbp]
-        \caption{Specialized NEAT vs Generalized PPO}
-        \begin{center}
-            \begin{tabular}{|c|c|c|c|c|c|}
-                \hline
-                \textbf{Opponent}&{\textbf{NEAT}}&{\textbf{PPO Run 1}}&{\textbf{PPO Run 2}}&{\textbf{PPO Run 3}}&{\textbf{PPO Best}} \\
-                \cline{2-4}
-                \hline
-                1 & 190.01 & 198.41 & 199.07 & 199.81 & 199.61 \\
-                2 & 194.01 & 199.74 & 199.27 & 190.34 & 189.14 \\
-                3 & 180.01 & 58.94 & 46.67 & 70.27 & 85.01 \\
-                4 & 194.01 & 58.34 & 66.5 & 64.11 & 80.17 \\
-                5 & 194.01 & 172.61 & 165.65 & 174.61 & 158.83 \\
-                6 & 173.01 & 195.73 & 196.95 & 195.85 & 194.37 \\
-                7 & 177.01 & 199.79 & 195.05 & 193.27 & 191.17 \\
-                8 & 186.01 & 122.17 & 145.49 & 141.89 & 140.61 \\
-                \hline
-            \end{tabular}
-        \end{center}
-    \end{table}
-
-    \begin{table}[htbp]
-        \caption{Results}
+        \caption{Specialized NEAT vs Generalized PPO (Gain)}
         \begin{center}
@@
                 1 & 198.41 & 199.07 & 199.81 & 199.61 & 190.01 \\
-                \hline
                 2 & 199.74 & 199.27 & 190.34 & 189.14 & 194.01 \\
-                \hline
                 3 & 58.94 & 46.67 & 70.27 & 85.01 & 180.01 \\
-                \hline
                 4 & 58.34 & 66.5 & 64.11 & 80.17 & 194.01 \\
-                \hline
                 5 & 172.61 & 165.65 & 174.61 & 158.83 & 194.01\\
-                \hline
                 6 & 195.73 & 196.95 & 195.85 & 194.37 & 173.01 \\
-                \hline
                 7 & 199.79 & 195.05 & 193.27 & 191.17 & 177.01 \\
-                \hline
                 8 & 122.17 & 145.49 & 141.89 & 140.61 & 186.01 \\
@@
 
+    \begin{table}[htbp]
+        \caption{Specialized NEAT vs Generalized PPO (Percentage of games won)}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|c|}
+                \hline
+                \textbf{Opponent}&\multicolumn{4}{|c|}{\textbf{PPO}} \\
+                \cline{2-5}
+                & \textbf{\textit{Run 1}}& \textbf{\textit{Run 2}}& \multicolumn{2}{|c|}{\textbf{\textit{Run 3}}} \\
+                \cline{4-5}
+                & 3000 ep & 3000 ep & 3000 ep & 1750 ep\\
+                & (best at training) & & & (best tested) \\
+                \hline
+                1 & 100 & 100 & 100 & 100 \\
+                2 & 100 & 100 & 100 & 100 \\
+                3 & 3.33 & 0 & 3.33 & 46.66 \\
+                4 & 0 & 6.66 & 0 & 6.66 \\
+                5 & 100 & 96.66 & 100 & 100 \\
+                6 & 100 & 100 & 100 & 100 \\
+                7 & 100 & 93.33 & 100 & 100 \\
+                8 & 80 & 100 & 96.66 & 90 \\
+                \hline
+
+            \end{tabular}
+        \end{center}
+    \end{table}
+
     \subsection{Time Analysis}\label{subsec:time-analysis}
@@
 
+    \section{Conclusions}\label{sec:conclusions}
+    After observing that the sparse methods were better than the iterative ones,
+    PPO with random initialization is better than PSO cascaded with PPO and
+    PPO with random initialization trained on four opponents leads to better train results
+    than specialized agents trained with NEAT we can conclude that the problem space is in such a
+    way that the methods with a greedy bias are moving away from the global optimum.
+
     \begin{thebibliography}{00}
+        \bibitem{evoman_competition} Evoman: Game-playing Competition for WCCI 2020, \url{http://pesquisa.ufabc.edu.br/hal/Evoman.html}
         \bibitem{evoman} Fabricio Olivetti de Franca, Denis Fantinato, Karine Miras, A.E. Eiben and Patricia A. Vargas.
@@
         \bibitem{evoman_blog} Karine Miras, Evoman, \url{https://karinemirasblog.wordpress.com/portfolio/evoman/}
-        \bibitem{evoman_competition} Evoman: Game-playing Competition for WCCI 2020, \url{http://pesquisa.ufabc.edu.br/hal/Evoman.html}
     \end{thebibliography}

### 2020-06-15 paper/paper.tex
sync 2.
Added final table.

@@
                 \textbf{Train opponents}&{\textbf{gain (harmonic mean)}} \\
-                \cline{2-4}
                 \hline
-                \{1, 2, 6, 7\} \cup \{3, 4, 5, 8\} & 44.16 \\
-                \{1, 2, 6, 7\} \cup \{3, 4, 6, 7\} & 75.3 \\
-                \{1, 2, 6, 7\} \cup \{1, 3, 6, 7\} & 110.19 \\
-                \{1, 2, 6, 7\} \cup \{2, 3, 6, 7\} & 136.1 \\
-                \{1, 2, 6, 7\} \cup \{2, 4, 6, 7\} & 85.22 \\
+                $\{1, 2, 6, 7\} \cup \{3, 4, 5, 8\}$ & $44.16$ \\
+                $\{1, 2, 6, 7\} \cup \{3, 4, 6, 7\}$ & $75.3$ \\
+                $\{1, 2, 6, 7\} \cup \{1, 3, 6, 7\}$ & $110.19$ \\
+                $\{1, 2, 6, 7\} \cup \{2, 3, 6, 7\}$ & $136.1$ \\
+                $\{1, 2, 6, 7\} \cup \{2, 4, 6, 7\}$ & $85.22$ \\
                 \hline
@@
 
+    \begin{table}[htbp]
+        \caption{Average player and enemy life. Average game duration}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|c|c|c|}
+                \hline
+                \textbf{Opponent}&\multicolumn{3}{|c|}{\textbf{Best train agent}}&\multicolumn{3}{|c|}{\textbf{Best test agent}} \\
+                \cline{2-7}
+                & \textbf{\textit{Avg player life}}& \textbf{\textit{Avg enemy life}}& \textbf{\textit{Avg duration}} & \textbf{\textit{Avg player life}}& \textbf{\textit{Avg enemy life}}& \textbf{\textit{Avg duration}} \\
+                \hline
+                1 & 97.26 &   100 &     0 &     0 & 159.63 & 174.23 \\
+                2 & 98.33 & 89.73 &     0 &     0 &  176.6 & 186.43 \\
+                3 &  0.13 &  7.66 & 43.33 & 21.33 & 457.26 & 410.53 \\
+                4 &     0 &  1.84 & 37.66 & 21.66 &  547.5 & 736.83 \\
+                5 & 71.32 & 60.52 &     0 &     0 & 379.13 & 367.93 \\
+                6 & 96.32 &    95 &     0 &     0 & 165.06 & 203.53 \\
+                7 & 99.74 &  85.9 &     0 &     0 & 162.06 & 135.13 \\
+                8 &  34.3 & 38.76 &     6 &     2 & 520.93 & 443.6 \\
+                \hline
+            \end{tabular}
+        \end{center}
+    \end{table}
+
     \subsection{Time Analysis}\label{subsec:time-analysis}
@@
         "An electronic-game framework for evaluating coevolutionary algorithms." arXiv:1604.00644 (2016).
-        \bibitem{neuro} Floreano, D., Dürr, P. & Mattiussi, C. Neuroevolution: from architectures to learning. Evol. Intel. 1, 47–62 (2008). \url{https://doi.org/10.1007/s12065-007-0002-4}
+        \bibitem{neuro} Floreano, D., Dürr, P. \& Mattiussi, C. Neuroevolution: from architectures to learning. Evol. Intel. 1, 47–62 (2008). \url{https://doi.org/10.1007/s12065-007-0002-4}
         \bibitem{capcom} M. MEGA, "Produced by capcom, distributed by capcom, 1987," System: NES.

### 2020-07-03 code/reinforcement_learning/parameters.py
Added plots with results for general training.

@@ -12,7 +12,7 @@ import os
 # [2, 3, 6, 7] 95.69
 
 ENEMIES_CHOSEN_FOR_TRAINING = range(1, 9)
-ENEMIES_DIFFICULTY = 5
+ENEMIES_DIFFICULTY = 1
 
 MODEL_HIDDEN_LAYERS_SIZES = (64, 64)
 MODEL_ACTIVATION = torch.nn.Tanh

### 2020-07-03 code/test_models/reinforcement_learning/parameters.py

@@ -1,13 +1,13 @@
 import os
 
 ENEMIES_CHOSEN_FOR_TESTING = range(1, 9)
-ENEMIES_DIFFICULTY = 2
+ENEMIES_DIFFICULTY = 5
 NR_EXPERIMENTS_FOR_EACH_ENEMY = 30
 
-# model_name = sorted(os.listdir('../trained_models/reinforcement_learning'))[-1]
+model_name = sorted(os.listdir('../trained_models/reinforcement_learning'))[-1]
 # model_name = '1591964820.216715'  # 124.23
 # model_name = '1591641901.9027252'  # ppo run 1 3000
 # model_name = '1591880695.9546342'  # ppo run 2 3000
 # model_name = '1591972086.6519217'  # ppo run 3 3000
-model_name = '1591935751.1014764'  # ppo run 3 1750
+# model_name = '1591935751.1014764'  # ppo run 3 1750
 MODEL_PATH = f'../trained_models/reinforcement_learning/{model_name}'

### 2020-07-03 code/test_models/reinforcement_learning/test_multiple.py

@@ -1,5 +1,5 @@
 # Fix error which might appear when this script is ran from the command line
-
+import json
 import sys
 import os
 
@@ -37,7 +37,7 @@ def save_csv_results(model_name, results):
         csvwriter = csv.writer(csvfile, delimiter=',', quotechar='|', quoting=csv.QUOTE_MINIMAL)
         csvwriter.writerow(
             ['enemy', 'gain', 'avg_player_life', 'avg_enemy_life', 'avg_duration', 'min_player_life', 'max_player_life',
-             'min_enemy_life', 'max_enemy_life', 'min_duration', 'max_duration'])
+             'min_enemy_life', 'max_enemy_life', 'min_duration', 'max_duration', 'percentage_games_lost'])
         csvwriter.writerows(results)
 
 
@@ -46,6 +46,14 @@ def test_model(model_name, ENEMIES_CHOSEN_FOR_TESTING=range(1, 9), NR_EXPERIMENT
     MODEL_PATH = f'../trained_models/reinforcement_learning/{model_name}'
     model = torch.load(os.path.join(MODEL_PATH, 'pyt_save', 'model.pt'))
 
+    with open(os.path.join(MODEL_PATH, 'config.json')) as data_file:
+        data = json.load(data_file)
+        try:
+            difficulty = data['enemies_difficulty']
+        except Exception as e:
+            difficulty = 2
+
+    print(model_name, difficulty)
     gains = []
     results = []
 
@@ -53,12 +61,16 @@ def test_model(model_name, ENEMIES_CHOSEN_FOR_TESTING=range(1, 9), NR_EXPERIMENT
         min_player_life, max_player_life, average_player_life = 100, 0, 0
         min_enemy_life, max_enemy_life, average_enemy_life = 100, 0, 0
         min_time, max_time, average_time = 100000000, 0, 0
+        number_of_games_lost = 0
 
         for experiment in range(NR_EXPERIMENTS_FOR_EACH_ENEMY):
             evoman_environment = EvomanEnvironmentWrapper('evoman rl test',
                                                           player_controller=TestReinforcementLearningEvomanPlayerController(),
-                                                          enemies=[enemy])
+                                                          enemies=[enemy],
+                                                          level=difficulty)
             _, player_life, enemy_life, time = evoman_environment.play(pcont=model)
+
[…]

### 2020-07-04 code/reinforcement_learning/parameters.py
Running specialized ppo.

@@ -11,14 +11,14 @@ import os
 # corrected
 # [2, 3, 6, 7] 95.69
 
-ENEMIES_CHOSEN_FOR_TRAINING = range(1, 9)
-ENEMIES_DIFFICULTY = 1
+ENEMIES_CHOSEN_FOR_TRAINING = [3]
+ENEMIES_DIFFICULTY = 2
 
 MODEL_HIDDEN_LAYERS_SIZES = (64, 64)
 MODEL_ACTIVATION = torch.nn.Tanh
 
 STEPS_PER_EPOCH = 10000
-EPOCHS = 250  # 2000
+EPOCHS = 1000  # 2000
 
 # trained_models/pso_bootstrap/[1, 2, 6, 7]/47.98_1591201550.222573.pickle 43.2
 # trained_models/pso_bootstrap/[1, 2, 6, 7]/52.13_1591215640.939492.pickle 42.7
@@ -28,10 +28,10 @@ EPOCHS = 250  # 2000
 # STARTING_MODEL_PATH = '../trained_models/pso_bootstrap/[1, 2, 6, 7]/61.38_1591194194.380356.pickle'
 # STARTING_MODEL_PATH = '../trained_models/reinforcement_learning/1591261004.787129'
 
-model_name = sorted(os.listdir('trained_models/reinforcement_learning'))[-1]
-STARTING_MODEL_PATH = f'../trained_models/reinforcement_learning/{model_name}'
+# model_name = sorted(os.listdir('trained_models/reinforcement_learning'))[-1]
+# STARTING_MODEL_PATH = f'../trained_models/reinforcement_learning/{model_name}'
 
-# STARTING_MODEL_PATH = None
+STARTING_MODEL_PATH = None
 IS_STARTING_MODEL_PSO = False
 
 NR_PARALLEL_PROCESSES = 3

### 2020-07-04 code/test_models/reinforcement_learning/parameters.py

@@ -1,13 +1,15 @@
 import os
 
-ENEMIES_CHOSEN_FOR_TESTING = range(1, 9)
-ENEMIES_DIFFICULTY = 5
+ENEMIES_CHOSEN_FOR_TESTING = [2]
+ENEMIES_DIFFICULTY = 2
 NR_EXPERIMENTS_FOR_EACH_ENEMY = 30
 
-model_name = sorted(os.listdir('../trained_models/reinforcement_learning'))[-1]
+# model_name = sorted(os.listdir('../trained_models/reinforcement_learning'))[-1]
 # model_name = '1591964820.216715'  # 124.23
 # model_name = '1591641901.9027252'  # ppo run 1 3000
 # model_name = '1591880695.9546342'  # ppo run 2 3000
 # model_name = '1591972086.6519217'  # ppo run 3 3000
 # model_name = '1591935751.1014764'  # ppo run 3 1750
+
+model_name = '1593785511.0696504'  # ppo opp 1
 MODEL_PATH = f'../trained_models/reinforcement_learning/{model_name}'

### 2020-07-05 code/reinforcement_learning/parameters.py
Updates from laptop.

@@ -11,7 +11,7 @@ import os
 # corrected
 # [2, 3, 6, 7] 95.69
 
-ENEMIES_CHOSEN_FOR_TRAINING = [3]
+ENEMIES_CHOSEN_FOR_TRAINING = [7]
 ENEMIES_DIFFICULTY = 2
 
 MODEL_HIDDEN_LAYERS_SIZES = (64, 64)

### 2020-07-05 code/reinforcement_learning/parameters.py
PC updates

@@ -11,7 +11,16 @@ import os
 # corrected
 # [2, 3, 6, 7] 95.69
 
-ENEMIES_CHOSEN_FOR_TRAINING = [3]
+# 1 - laptop. done
+# 2 - laptop. done
+# 3 - laptop. done
+# 4 - pc. done
+# 5 - laptop. done
+# 6 - pc. done
+# 7 - laptop. in progress
+# 8 - pc. in progress
+
+ENEMIES_CHOSEN_FOR_TRAINING = [8]
 ENEMIES_DIFFICULTY = 2
 
 MODEL_HIDDEN_LAYERS_SIZES = (64, 64)

### 2020-07-05 code/test_models/reinforcement_learning/test_multiple.py
Added final reports for specialized PPO.

@@ -52,8 +52,9 @@ def test_model(model_name, ENEMIES_CHOSEN_FOR_TESTING=range(1, 9), NR_EXPERIMENT
             difficulty = data['enemies_difficulty']
         except Exception as e:
             difficulty = 2
+        ENEMIES_CHOSEN_FOR_TESTING = data['enemies']
 
-    print(model_name, difficulty)
+    print(model_name, difficulty, ENEMIES_CHOSEN_FOR_TESTING)
     gains = []
     results = []
 
@@ -113,4 +114,4 @@ def test_models(models):
 
 
 if __name__ == '__main__':
-    test_models(list(sorted(os.listdir('../trained_models/reinforcement_learning'))[-(12 * 4):]))
+    test_models(list(sorted(os.listdir('../trained_models/reinforcement_learning'))[-8:]))

### 2020-07-05 paper/paper.tex
Corrected table with results.
Added average.
Report updated for specialized PPO.

@@
                 \hline
-                1 & 97.26 &   100 &     0 &     0 & 159.63 & 174.23 \\
-                2 & 98.33 & 89.73 &     0 &     0 &  176.6 & 186.43 \\
-                3 &  0.13 &  7.66 & 43.33 & 21.33 & 457.26 & 410.53 \\
-                4 &     0 &  1.84 & 37.66 & 21.66 &  547.5 & 736.83 \\
-                5 & 71.32 & 60.52 &     0 &     0 & 379.13 & 367.93 \\
-                6 & 96.32 &    95 &     0 &     0 & 165.06 & 203.53 \\
-                7 & 99.74 &  85.9 &     0 &     0 & 162.06 & 135.13 \\
-                8 &  34.3 & 38.76 &     6 &     2 & 520.93 & 443.6 \\
+                1 & 97.26 &     0 &  159.63 &    100 &     0 &  174.23 \\
+                2 & 98.33 &     0 &   176.6 &  89.73 &     0 &  186.43 \\
+                3 &  0.13 & 43.33 &  457.26 &   7.66 & 21.33 &  410.53 \\
+                4 &     0 & 37.66 &   547.5 &   1.84 & 21.66 &  736.83 \\
+                5 & 71.32 &     0 &  379.13 &  60.52 &     0 &  367.93 \\
+                6 & 96.32 &     0 &  165.06 &     95 &     0 &  203.53 \\
+                7 & 99.74 &     0 &  162.06 &   85.9 &     0 &  135.13 \\
+                8 &  34.3 &     6 &  520.93 &  38.76 &     2 &   443.6 \\
+                \hline
+                \textbf{Average} & 62.17 & 10.87 & 321.02 & 59.92 & 5.62 & 332.27 \\
+            \end{tabular}
+        \end{center}
+    \end{table}
+
+    \subsection{Specialized PPO}\label{subsec:specialized-ppo}
+    In order to obtain a score higher than the upper bound we have tried two methods:
+    \begin{itemize}
+        \item A generalized PPO agent trained against all $8$ opponents.
+        \item A specialized PPO agent for each of the $8$ opponents.
+    \end{itemize}
+
+    \subsubsection{PPO trained against all opponents}
+    At the default difficulty of 2, a PPO agent trained against all opponents
+    obtained a final gain of 185.57, being lower with 0.1 than the
+    upper bound found with specialized NEAT\@.
+    For such an agent we have also studied how its gain changes when the difficulty
+    is modified.
+    \begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/general_harmonic_gain.eps}
+    \caption{Gain Evolution for PPO when trained against all 8 opponents.}
+    \label{fig:ppo_train_8}
+    \end{figure}
+
+    \begin{table}[htbp]
+        \caption{PPO trained against all opponents (Gain)}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|c|c|}
+                \hline
+                \textbf{Opponent}&\multicolumn{5}{|c|}{\textbf{Difficulty}} \\
+                \cline{2-6}
+                & \textbf{\textit{1}}& \textbf{\textit{2}}& \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} \\
+                \hline
+                 1 &  200.01 &  199.54 &  170.75 &  155.48 &   41.08\\
+                 2 &  200.01 &  200.01 &  182.81 &  168.81 &  168.51\\
+                 3 &  199.38 &  194.81 &  164.31 &  127.54 &  130.61\\
+                 4 &  186.14 &  180.61 &  154.23 &  181.58 &   60.28\\
+                 5 &  196.56 &  187.77 &  185.34 &  142.41 &  158.31\\
+                 6 &  196.35 &  174.43 &  179.19 &   23.84 &    9.88\\
+                 7 &  199.26 &  183.07 &   95.40 &    8.01 &    8.28\\
+                 8 &  196.57 &  169.31 &  123.00 &   41.34 &   10.01\\
+                \hline
+                \textbf{harmonic mean} & 196.68 & 185.57 & 149.57 & 35.76 & 20.89 \\
                 \hline
             \end{tabular}
+            \label{PPO against all opponents gain}
+        \end{center}
+    \end{table}
+
+    \begin{table}[htbp]
+        \caption{PPO trained against all opponents (percent games lost)}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|c|c|}
+                \hline
+                \textbf{Opponent}&\multicolumn{5}{|c|}{\textbf{Difficulty}} \\
+                \cline{2-6}
+                & \textbf{\textit{1}}& \textbf{\textit{2}}& \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} \\
+                \hline
+                 1 &  0 &  0 &     20 &  23.33 &   100 \\
+                 2 &  0 &  0 &      0 &      0 &     0 \\
+                 3 &  0 &  0 &      0 &  13.33 &  3.33 \\
+                 4 &  0 &  0 &      0 &      0 &    60 \\
+                 5 &  0 &  0 &      0 &      0 &     0 \\
+                 6 &  0 &  0 &      0 &    100 &   100 \\
+                 7 &  0 &  0 &  56.66 &    100 &   100 \\
+                 8 &  0 &  0 &     10 &    100 &   100 \\
+                \hline
+            \end{tabular}
+            \label{PPO against all opponents games lost}
+        \end{center}
+    \end{table}
+
+    \subsubsection{A PPO agent for each opponent}
+    In the second attempt to outscore the upper bound we have trained an agent
+    for each of the $8$ opponents.
+    Due to the high computational time this experiment was done only at
+    difficulty 2 and the number of training epochs for PPO was decreased from 3000 to 1000.
+    It won all its games against all opponents at difficulty 2.
+    \begin{table}[htbp]
+        \caption{PPO generalized vs PPO specialized vs NEAT specialized}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|}
+                \hline
+                \textbf{Opponent}&\textbf{PPO g.}&\textbf{PPO s.}&\textbf{NEAT s.} \\
+                \hline
+                 1 &  199.54 &  199.67 &  190.01 \\
+                 2 &  200.01 &  199.61 &  194.01 \\
+                 3 &  194.81 &  199.94 &  180.01 \\
+                 4 &  180.61 &  195.95 &  194.01 \\
+                 5 &  187.77 &  198.03 &  194.01 \\
+                 6 &  174.43 &  199.67 &  173.01 \\
+                 7 &  183.07 &  197.73 &  177.01 \\
+                 8 &  169.31 &  195.07 &  186.01 \\
+                \hline
+                \textbf{harmonic mean} & 185.57 & 198.19 & 185.67 \\
+                \hline
+            \end{tabular}
+            \label{PPO generalized vs specialized vs NEAT}
         \end{center}

### 2020-07-05 code/reinforcement_learning/main.py
Preparing specialized PPO against difficulty 5.

@@ -3,44 +3,50 @@
 import sys
 import os
 
-is_running_from_command_line = len(sys.path) <= 7
-if is_running_from_command_line:
-    script_path_tokens = sys.path[0].split(os.sep)
-    sys.path.extend([os.path.join('/', *script_path_tokens[:-1]),
-                     os.path.join('/', *script_path_tokens[:-2], 'evoman_framework', 'evoman')])
-
-# Parallelization part, had to be done some of the hacks
-
-from spinup_utils.mpi_tools import mpi_fork
-import reinforcement_learning.parameters as parameters
-
-mpi_fork(parameters.NR_PARALLEL_PROCESSES)
-
-# Fix evoman resources loading not working because framework assumes the current working directory is
-# always `evoman_framework`
-
-if is_running_from_command_line:
-    if not os.getcwd().endswith('evoman_framework'):
-        os.chdir('./evoman_framework')
-else:
-    if not os.getcwd().endswith('evoman_framework'):
-        os.chdir('../../evoman_framework')
-
-# Start of code without hacks
 
 
-from reinforcement_learning.ppo.ppo import ppo
 
-import time
+if __name__ == '__main__':
+    is_running_from_command_line = len(sys.path) <= 7
+    if is_running_from_command_line:
+        script_path_tokens = sys.path[0].split(os.sep)
+        sys.path.extend([os.path.join('/', *script_path_tokens[:-1]),
+                         os.path.join('/', *script_path_tokens[:-2], 'evoman_framework', 'evoman')])
 
-logger_kwargs = dict(output_dir=f'../trained_models/reinforcement_learning/{time.time()}',
-                     exp_name='evoman reinforcement learning')
-seed = int(time.time())
+    # Parallelization part, had to be done some of the hacks
 
-ppo(ac_kwargs={'hidden_sizes': parameters.MODEL_HIDDEN_LAYERS_SIZES, 'activation': parameters.MODEL_ACTIVATION},
-    seed=seed, steps_per_epoch=parameters.STEPS_PER_EPOCH, epochs=parameters.EPOCHS,
-    enemies=parameters.ENEMIES_CHOSEN_FOR_TRAINING, enemies_difficulty=parameters.ENEMIES_DIFFICULTY, gamma=parameters.GAMMA, clip_ratio=parameters.CLIP_RATIO,
[…]

### 2020-07-05 code/reinforcement_learning/parameters.py

@@ -20,14 +20,20 @@ import os
 # 7 - laptop. in progress
 # 8 - pc. in progress
 
-ENEMIES_CHOSEN_FOR_TRAINING = [8]
-ENEMIES_DIFFICULTY = 2
+# with open('last_opp.txt', 'r') as file:
+#     last = int(file.readlines()[-1])
+#
+# with open('last_opp.txt', 'a') as file:
+#     file.write(f'\n{last + 1}')
+#
+# ENEMIES_CHOSEN_FOR_TRAINING = [last + 1]
+ENEMIES_DIFFICULTY = 5
 
 MODEL_HIDDEN_LAYERS_SIZES = (64, 64)
 MODEL_ACTIVATION = torch.nn.Tanh
 
 STEPS_PER_EPOCH = 10000
-EPOCHS = 1000  # 2000
+EPOCHS = 2  # 1000 2000
 
 # trained_models/pso_bootstrap/[1, 2, 6, 7]/47.98_1591201550.222573.pickle 43.2
 # trained_models/pso_bootstrap/[1, 2, 6, 7]/52.13_1591215640.939492.pickle 42.7

### 2020-07-05 scripts/reinforcement_learning.sh

@@ -1,11 +1,11 @@
 source ./venv/bin/activate
 
-NR_OF_EXPERIMENTS_TO_RUN=1
+NR_OF_EXPERIMENTS_TO_RUN=4
 
 printf "Please run this script from the root folder of this repository\n"
 printf "e.g. ~/evoman-game-playing-competition$ scripts/reinforcement_learning.sh"
 
 for ((i = 1; i <= NR_OF_EXPERIMENTS_TO_RUN; i++)); do
   printf "\n\nRunning experiment number %d\n\n" "$i"
-  python "code/reinforcement_learning/main.py"
+  python "code/reinforcement_learning/main.py" $i
 done

### 2020-07-07 code/reinforcement_learning/parameters.py
laptop diff 5 sync.

@@ -33,7 +33,7 @@ MODEL_HIDDEN_LAYERS_SIZES = (64, 64)
 MODEL_ACTIVATION = torch.nn.Tanh
 
 STEPS_PER_EPOCH = 10000
-EPOCHS = 2  # 1000 2000
+EPOCHS = 1000  # 1000 2000
 
 # trained_models/pso_bootstrap/[1, 2, 6, 7]/47.98_1591201550.222573.pickle 43.2
 # trained_models/pso_bootstrap/[1, 2, 6, 7]/52.13_1591215640.939492.pickle 42.7

### 2020-07-07 code/test_models/reinforcement_learning/test_multiple.py
Added results without opponent 4.

@@ -48,13 +48,13 @@ def test_model(model_name, ENEMIES_CHOSEN_FOR_TESTING=range(1, 9), NR_EXPERIMENT
 
     with open(os.path.join(MODEL_PATH, 'config.json')) as data_file:
         data = json.load(data_file)
-        try:
-            difficulty = data['enemies_difficulty']
-        except Exception as e:
-            difficulty = 2
+        # try:
+        #     difficulty = data['enemies_difficulty']
+        # except Exception as e:
+        #     difficulty = 2
         ENEMIES_CHOSEN_FOR_TESTING = data['enemies']
-
-    print(model_name, difficulty, ENEMIES_CHOSEN_FOR_TESTING)
+    difficulty = 5
+    print(model_name, ENEMIES_CHOSEN_FOR_TESTING)
     gains = []
     results = []
 

### 2020-07-07 code/test_models/reinforcement_learning/test_multiple.py
Finished training.

@@ -114,4 +114,4 @@ def test_models(models):
 
 
 if __name__ == '__main__':
-    test_models(list(sorted(os.listdir('../trained_models/reinforcement_learning'))[-8:]))
+    test_models(list(sorted(os.listdir('../trained_models/reinforcement_learning'))[-1:]))

### 2020-07-18 code/reinforcement_learning/parameters.py
Finished specialized PPO training and testing at diff 4.

@@ -27,7 +27,7 @@ import os
 #     file.write(f'\n{last + 1}')
 #
 # ENEMIES_CHOSEN_FOR_TRAINING = [last + 1]
-ENEMIES_DIFFICULTY = 5
+ENEMIES_DIFFICULTY = 4
 
 MODEL_HIDDEN_LAYERS_SIZES = (64, 64)
 MODEL_ACTIVATION = torch.nn.Tanh

### 2020-07-18 code/test_models/reinforcement_learning/test_multiple.py

@@ -53,8 +53,8 @@ def test_model(model_name, ENEMIES_CHOSEN_FOR_TESTING=range(1, 9), NR_EXPERIMENT
         # except Exception as e:
         #     difficulty = 2
         ENEMIES_CHOSEN_FOR_TESTING = data['enemies']
-    difficulty = 5
-    print(model_name, ENEMIES_CHOSEN_FOR_TESTING)
+        difficulty = data['enemies_difficulty']
+    print(model_name, ENEMIES_CHOSEN_FOR_TESTING, 'at difficulty', difficulty)
     gains = []
     results = []
 
@@ -114,4 +114,4 @@ def test_models(models):
 
 
 if __name__ == '__main__':
-    test_models(list(sorted(os.listdir('../trained_models/reinforcement_learning'))[-1:]))
+    test_models(list(sorted(os.listdir('../trained_models/reinforcement_learning'))[-8:]))

### 2020-07-18 scripts/reinforcement_learning.sh

@@ -5,7 +5,7 @@ NR_OF_EXPERIMENTS_TO_RUN=8
 printf "Please run this script from the root folder of this repository\n"
 printf "e.g. ~/evoman-game-playing-competition$ scripts/reinforcement_learning.sh"
 
-for ((i = 5; i <= NR_OF_EXPERIMENTS_TO_RUN; i++)); do
+for ((i = 1; i <= NR_OF_EXPERIMENTS_TO_RUN; i++)); do
   printf "\n\nRunning experiment number %d\n\n" "$i"
   python "code/reinforcement_learning/main.py" $i
 done

### 2020-07-20 code/reinforcement_learning/parameters.py
Dif 3 opp 7, 8.

@@ -27,7 +27,7 @@ import os
 #     file.write(f'\n{last + 1}')
 #
 # ENEMIES_CHOSEN_FOR_TRAINING = [last + 1]
-ENEMIES_DIFFICULTY = 4
+ENEMIES_DIFFICULTY = 3
 
 MODEL_HIDDEN_LAYERS_SIZES = (64, 64)
 MODEL_ACTIVATION = torch.nn.Tanh

### 2020-07-20 code/test_models/reinforcement_learning/test_multiple.py

@@ -114,4 +114,4 @@ def test_models(models):
 
 
 if __name__ == '__main__':
-    test_models(list(sorted(os.listdir('../trained_models/reinforcement_learning'))[-8:]))
+    test_models(list(sorted(os.listdir('../trained_models/reinforcement_learning'))[-2:]))

### 2020-07-20 code/test_models/reinforcement_learning/test_multiple.py
Dif 3 Opp 1-6.

@@ -114,4 +114,4 @@ def test_models(models):
 
 
 if __name__ == '__main__':
-    test_models(list(sorted(os.listdir('../trained_models/reinforcement_learning'))[-8:]))
+    test_models(list(sorted(os.listdir('../trained_models/reinforcement_learning'))[-6:]))

### 2020-07-20 scripts/reinforcement_learning.sh

@@ -5,7 +5,7 @@ NR_OF_EXPERIMENTS_TO_RUN=8
 printf "Please run this script from the root folder of this repository\n"
 printf "e.g. ~/evoman-game-playing-competition$ scripts/reinforcement_learning.sh"
 
-for ((i = 1; i <= NR_OF_EXPERIMENTS_TO_RUN; i++)); do
+for ((i = 7; i <= NR_OF_EXPERIMENTS_TO_RUN; i++)); do
   printf "\n\nRunning experiment number %d\n\n" "$i"
   python "code/reinforcement_learning/main.py" $i
 done

### 2020-07-20 code/reinforcement_learning/parameters.py
Dif 3 Opp 1-8 with results.

@@ -27,7 +27,7 @@ import os
 #     file.write(f'\n{last + 1}')
 #
 # ENEMIES_CHOSEN_FOR_TRAINING = [last + 1]
-ENEMIES_DIFFICULTY = 3
+ENEMIES_DIFFICULTY = 1
 
 MODEL_HIDDEN_LAYERS_SIZES = (64, 64)
 MODEL_ACTIVATION = torch.nn.Tanh

### 2020-07-20 scripts/reinforcement_learning.sh

@@ -1,11 +1,11 @@
 source ./venv/bin/activate
 
-NR_OF_EXPERIMENTS_TO_RUN=8
+NR_OF_EXPERIMENTS_TO_RUN=6
 
 printf "Please run this script from the root folder of this repository\n"
 printf "e.g. ~/evoman-game-playing-competition$ scripts/reinforcement_learning.sh"
 
-for ((i = 7; i <= NR_OF_EXPERIMENTS_TO_RUN; i++)); do
+for ((i = 5; i <= NR_OF_EXPERIMENTS_TO_RUN; i++)); do
   printf "\n\nRunning experiment number %d\n\n" "$i"
   python "code/reinforcement_learning/main.py" $i
 done

### 2020-07-20 paper/paper.tex
Updated report with results of specialized PPO against difficulty 5.
Ready for report.

@@
                 \hline
-                \textbf{Train opponents}&{\textbf{gain (harmonic mean)}} \\
+                \textbf{Train opponents}             & {\textbf{gain (harmonic mean)}} \\
                 \hline
@@
                 \hline
-                \textbf{Opponent}&\multicolumn{4}{|c|}{\textbf{PPO}}&\textbf{Specialized} \\
+                \textbf{Opponent} & \multicolumn{4}{|c|}{\textbf{PPO}} & \textbf{Specialized} \\
                 \cline{2-5}
-                & \textbf{\textit{Run 1}}& \textbf{\textit{Run 2}}& \multicolumn{2}{|c|}{\textbf{\textit{Run 3}}} & NEAT \\
+                & \textbf{\textit{Run 1}} & \textbf{\textit{Run 2}} & \multicolumn{2}{|c|}{\textbf{\textit{Run 3}}} & NEAT \\
                 \cline{4-5}
@@
                 4                       & 58.34              & 66.5    & 64.11   & 80.17         & 194.01 \\
-                5 & 172.61 & 165.65 & 174.61 & 158.83 & 194.01\\
+                5                       & 172.61             & 165.65  & 174.61  & 158.83        & 194.01 \\
                 6                       & 195.73             & 196.95  & 195.85  & 194.37        & 173.01 \\
@@
                 \textbf{final}          &                    &         &         &               &        \\
-                \textbf{harmonic mean} & 117.22 & 114.91 & 127.01 & 137.18 & 185.67\\
+                \textbf{harmonic mean}  & 117.22             & 114.91  & 127.01  & 137.18        & 185.67 \\
                 \hline
@@
                 \hline
-                \textbf{Opponent}&\multicolumn{4}{|c|}{\textbf{PPO}} \\
+                \textbf{Opponent} & \multicolumn{4}{|c|}{\textbf{PPO}} \\
                 \cline{2-5}
-                & \textbf{\textit{Run 1}}& \textbf{\textit{Run 2}}& \multicolumn{2}{|c|}{\textbf{\textit{Run 3}}} \\
+                & \textbf{\textit{Run 1}} & \textbf{\textit{Run 2}} & \multicolumn{2}{|c|}{\textbf{\textit{Run 3}}} \\
                 \cline{4-5}
-                & 3000 ep & 3000 ep & 3000 ep & 1750 ep\\
+                & 3000 ep            & 3000 ep & 3000 ep & 1750 ep       \\
                 & (best at training) &         &         & (best tested) \\
@@
                 \hline
-                \textbf{Opponent}&\multicolumn{3}{|c|}{\textbf{Best train agent}}&\multicolumn{3}{|c|}{\textbf{Best test agent}} \\
+                \textbf{Opponent} & \multicolumn{3}{|c|}{\textbf{Best train agent}} & \multicolumn{3}{|c|}{\textbf{Best test agent}} \\
                 \cline{2-7}
-                & \textbf{\textit{Avg player life}}& \textbf{\textit{Avg enemy life}}& \textbf{\textit{Avg duration}} & \textbf{\textit{Avg player life}}& \textbf{\textit{Avg enemy life}}& \textbf{\textit{Avg duration}} \\
+                & \textbf{\textit{Avg player life}} & \textbf{\textit{Avg enemy life}} & \textbf{\textit{Avg duration}} & \textbf{\textit{Avg player life}} & \textbf{\textit{Avg enemy life}} & \textbf{\textit{Avg duration}} \\
                 \hline
@@
                 \hline
-                \textbf{Opponent}&\multicolumn{5}{|c|}{\textbf{Difficulty}} \\
+                \textbf{Opponent} & \multicolumn{5}{|c|}{\textbf{Difficulty}} \\
                 \cline{2-6}
-                & \textbf{\textit{1}}& \textbf{\textit{2}}& \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} \\
+                & \textbf{\textit{1}} & \textbf{\textit{2}} & \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} \\
                 \hline
-                 1 &  200.01 &  199.54 &  170.75 &  155.48 &   41.08\\
-                 2 &  200.01 &  200.01 &  182.81 &  168.81 &  168.51\\
-                 3 &  199.38 &  194.81 &  164.31 &  127.54 &  130.61\\
-                 4 &  186.14 &  180.61 &  154.23 &  181.58 &   60.28\\
-                 5 &  196.56 &  187.77 &  185.34 &  142.41 &  158.31\\
-                 6 &  196.35 &  174.43 &  179.19 &   23.84 &    9.88\\
-                 7 &  199.26 &  183.07 &   95.40 &    8.01 &    8.28\\
-                 8 &  196.57 &  169.31 &  123.00 &   41.34 &   10.01\\
+                1                      & 200.01              & 199.54              & 170.75              & 155.48              & 41.08               \\
+                2                      & 200.01              & 200.01              & 182.81              & 168.81              & 168.51              \\
+                3                      & 199.38              & 194.81              & 164.31              & 127.54              & 130.61              \\
+                4                      & 186.14              & 180.61              & 154.23              & 181.58              & 60.28               \\
+                5                      & 196.56              & 187.77              & 185.34              & 142.41              & 158.31              \\
+                6                      & 196.35              & 174.43              & 179.19              & 23.84               & 9.88                \\
+                7                      & 199.26              & 183.07              & 95.40               & 8.01                & 8.28                \\
+                8                      & 196.57              & 169.31              & 123.00              & 41.34               & 10.01               \\
                 \hline
@@
                 \hline
-                \textbf{Opponent}&\multicolumn{5}{|c|}{\textbf{Difficulty}} \\
+                \textbf{Opponent} & \multicolumn{5}{|c|}{\textbf{Difficulty}} \\
                 \cline{2-6}
-                & \textbf{\textit{1}}& \textbf{\textit{2}}& \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} \\
+                & \textbf{\textit{1}} & \textbf{\textit{2}} & \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} \\
                 \hline
@@
                 \hline
-                \textbf{Opponent}&\textbf{PPO g.}&\textbf{PPO s.}&\textbf{NEAT s.} \\
+                \textbf{Opponent}      & \textbf{PPO g.} & \textbf{PPO s.} & \textbf{NEAT s.} \\
                 \hline
@@
 
+    \subsection{Specialized PPO at difficulty 5}\label{subsec:specialized-ppo-at-difficulty-5}
+    A PPO agent for each opponent.
+    1000 training epochs per opponent.
+    (10 million frames for each opponent).
+    Harmonic mean of gains is 154.58.
+    \begin{table}[htbp]
+        \caption{Specialized PPO at difficulty 5}
+        \begin{center}
+            \begin{tabular}{|c|c|c|}
+                \hline
+                \textbf{Opponent} & \textbf{Gain} & \textbf{Percent games lost} \\
+                \hline
+                1                 & 166.01        & 13.33                       \\
+                2                 & 197.18        & 0.00                        \\
+                3                 & 72.28         & 100.00                      \\
+                4                 & 191.23        & 0.00                        \\
+                5                 & 196.71        & 0.00                        \\
+                6                 & 198.76        & 0.00                        \\
+                7                 & 176.01        & 0.00                        \\
+                8                 & 172.61        & 0.00                        \\
+                \hline
+            \end{tabular}
+            \label{PPO generalized at difficulty 5}
+        \end{center}
+    \end{table}
+
     \subsection{Time Analysis}\label{subsec:time-analysis}
@@
     The average number of frames per game is 287.
+
+    \subsection{Specialized PPO at multiple difficulties}\label{subsec:specialized-ppo-at-multiple-difficulties}
+
+    \begin{table}[htbp]
+        \caption{Specialized PPO against multiple difficulties (gain)}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|c|}
+                \hline
+                \textbf{Opponent} & \multicolumn{4}{|c|}{\textbf{Difficulty}} \\
+                \cline{2-5}
+                & \textbf{\textit{2}} & \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} \\
+                \hline
+                1                      & 199.68              & 196.51              & 189.21              & 166.01              \\
+                2                      & 199.61              & 199.91              & 199.21              & 197.18              \\
+                3                      & 199.94              & 199.51              & 147.21              & 72.28               \\
+                4                      & 195.95              & 196.14              & 198.81              & 191.23              \\
+                5                      & 198.03              & 193.38              & 191.73              & 196.71              \\
+                6                      & 199.67              & 195.18              & 196.85              & 198.76              \\
+                7                      & 197.73              & 192.45              & 184.85              & 176.01              \\
+                8                      & 195.07              & 185.37              & 178.05              & 172.61              \\
+                \hline
+                \textbf{harmonic mean} & 198.19              & 194.7               & 184.12              & 154.58              \\
+                \hline
+            \end{tabular}
+            \label{Specialized PPO against multiple difficulties (gain)}
+        \end{center}
+    \end{table}
+
+    \begin{figure}[htbp]
+        \centering
+        \includegraphics[width=0.5\textwidth]{images/specialized_ppo_harmonic_gain_per_diff.eps}
+        \caption{Harmonic Gain of specialized PPO at multiple difficulties.}
+        \label{fig:specialized_ppo_harmonic_gain_per_diff}
+    \end{figure}
+
+    \begin{figure}[htbp]
+        \centering
+        \includegraphics[width=0.5\textwidth]{images/specialized_ppo_gain_per_diff_per_opp.eps}
+        \caption{Gain of specialized PPO at multiple difficulties for each opponent.}
+        \label{fig:specialized_ppo_gain_per_diff_per_opp}
+    \end{figure}
+
+    \begin{table}[htbp]
+        \caption{Specialized PPO against multiple difficulties (percentage games lost)}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|c|}
+                \hline
+                \textbf{Opponent} & \multicolumn{4}{|c|}{\textbf{Difficulty}} \\
+                \cline{2-5}
+                & \textbf{\textit{2}} & \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} \\
+                \hline
+                1 & 0.00                & 0.00                & 0.00                & 13.33               \\
+                2 & 0.00                & 0.00                & 0.00                & 0.00                \\
+                3 & 0.00                & 0.00                & 36.67               & 100.00              \\
+                4 & 0.00                & 0.00                & 0.00                & 0.00                \\
+                5 & 0.00                & 0.00                & 0.00                & 0.00                \\
+                6 & 0.00                & 0.00                & 0.00                & 0.00                \\
+                7 & 0.00                & 0.00                & 0.00                & 0.00                \\
+                8 & 0.00                & 0.00                & 0.00                & 0.00                \\
+                \hline
+            \end{tabular}
+            \label{Specialized PPO against multiple difficulties (percentage games lost)}
+        \end{center}
+    \end{table}
 

### 2020-07-20 paper/paper.tex
Add files via upload

--- 2020-06-15 paper/paper.tex
@@
 \usepackage{xcolor}
-\usepackage{hyperref}
+\usepackage[draft]{hyperref}
 
@@
 
-    \title{A Comparison of Algorithms Playing EvoMan}
+    \title{Increasing the Upper Bound for the EvoMan Game Competition}
 
-    \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-        \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-            \textit{name of organization (of Aff.)}\\
-            City, Country \\
-            email address or ORCID}
+    \author{\IEEEauthorblockN{Gabriel-Codrin \textbf{Cojocaru}}
+        \IEEEauthorblockA{\textit{Faculty of Computer Science} \\
+            \textit{``Al. I. Cuza'' University}\\
+            Iasi, Romania \\
+            gabriel.cojocaru@info.uaic.ro}
           \and
-        \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-        \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-            \textit{name of organization (of Aff.)}\\
-            City, Country \\
-            email address or ORCID}
+
+          \IEEEauthorblockN{Sergiu-Andrei \textbf{Dinu}}
+          \IEEEauthorblockA{\textit{Faculty of Computer Science} \\
+            \textit{``Al. I. Cuza'' University}\\
+            Iasi, Romania \\
+            sergiu.dinu@info.uaic.ro}
+
         \and
-        \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-        \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-            \textit{name of organization (of Aff.)}\\
-            City, Country \\
-            email address or ORCID}
+        \IEEEauthorblockN{Eugen \textbf{Croitoru}}
+        \IEEEauthorblockA{\textit{Dept. of Computer Science} \\
+          \textit{``Al. I. Cuza'' University}\\
+          Iasi, Romania \\
+          eugennc@uaic.ro}
       }
@@
     \begin{abstract}
-        This paper describes a comparison between algorithms for evolving agents for the game Evoman.
-        We have tried a cascade ensemble method, starting with an algorithm that is either fast either has
-        an explorative bias and refining its solution with an exploitative algorithm.
-        The algorithms we have tested for the first stage are Q-learning, genetic algorithms and particle swarm optimization.
-        All of these algorithms are searching for the weights of a neural network of a fixed structure.
-        Both the genetic algorithms and the particle swarm optimization algorithms were tested with a sparse and
-        an iterative approach.
-        The best explorative algorithm was the sparse particle swarm optimization.
-        The exploitative algorithm we tested is proximal policy optimization.
-        Using PPO with random initialization lead to better results than an ensemble method of PSO cascaded with PPO\@.
-        We also detected overfitting of the agent trained with PPO\@.
+      This paper describes a comparison between algorithms for evolving agents able to play the game Evoman.
+      Our team took part in the ``Evoman: Game-playing Competition for WCCI 2020'', and won second place;
+      beyond finding a good agent to satisfy the requirements of the competition - which aim at a good ability to generalise -,
+      we have surpassed the existing non-general, best-known upper-bound.
+      We have managed to exceed this upper bound with a Proximal Policy Optimization algorithm, by discarding the competition requirements to generalise.
+      We also present our other exploratory attempts: Q-learning, Genetic Algorithms, Particle Swarm Optimisation, and their PPO hybridizations. 
+      Finally, we map the behaviour of our algorithm in the space of game difficulty, generating plausible extensions to the existing upper-bound.
     \end{abstract}
@@
     \begin{IEEEkeywords}
-        game-playing agent, artificial intelligence, EvoMan, genetic algorithm, reinforcement learning,
-        q-learning, neuroevolution, particle swarm optimization, proximal policy optimization, ppo
+        game-playing agent, Artificial Intelligence, EvoMan, Genetic Algorithm, Reinforcement Learning,
+        Q-learning, Neuroevolution, Particle Swarm Optimization, Proximal Policy Optimization
     \end{IEEEkeywords}
@@
     \section{Introduction}\label{sec:introduction}
-    This paper presents our solution for the "Evoman: Game-playing Competition for WCCI 2020"\cite{evoman_competition}.
-    We are attempting to train an agent playing the 2D shooting game Evoman\cite{evoman}.
-    We have tried an ensemble cascade method with two stages.
-    For the first stage we tested algorithms that either found acceptable solutions fast either had a high
-    explorative bias.
-    The algorithms we considered for this stage are Q-Learning\cite{q_learning}, genetic algorithms\cite{genetic_algorithm}
-    and particle swarm optimization\cite{pso}.
-    Both the GA and the PSO train the weights of a neural network of a fixed structure.
-    Both the GA and the PSO were tested with a sparse approach and an iterative approach.
-    For the exploitative stage we used Proximal Policy Optimization\cite{ppo}.
+    This paper contains our improvement on the upper-bound for playing the 2D shooting / platformer game Evoman\cite{evoman}.
+    Our team took part in the ``Evoman: Game-playing Competition for WCCI 2020''\cite{evoman_competition}.
+While placing second\cite{evoman_competition_results}, we have managed to exceed the best-known upper-bound.
+This difference is due to the competition requirements for generalisation; the upper-bound has been provided by the organisers by dropping such requirements, and only by doing the same could we exceed it.
+
+    Our initial exploratory attempts focused on an ensemble, $2$-stage cascade method.
+The first stage, designed to find good candidate solutions quickly, was either a Q-Learning\cite{q_learning}, a Genetic Algorithm\cite{genetic_algorithm} or Particle Swarm Optimization\cite{pso} algorithm.
+The candidate solutions are weights on a fixed-structure Artificial Neural Network, which plays the Evoman game.
+The second stage, aimed at further refining the solutions found by the first stage, is a slower Proximal Policy Optimization\cite{ppo} algorithm.
+
+    However, a simple random initialisation outperforms any of the first-stage algorithms, and the later experiments use PPO alone.
+
+    The competition required the algorithm to train with half the opponents in the game, and to test with all of them.
+Our method shows significant overfit in this challenge.
+However, its otherwise good performance makes it appropriate for setting a new upper bound, since the upper bound is being computed individually, by training and testing against each opponent in the Evoman game.
 
@@
 
-    EvoMan~\cite{karinemiras,evoman} is a framework for testing competitive game-playing agents.
-    This framework is inspired by Mega Man II~\cite{capcom}, the game created by Capcom.
-    EvoMan is a 2D shooting game where the player controls an agent playing against an opponent.
-    The agent will collect information about the environment through 20 sensors (Fig 1):
+    Evoman\cite{karinemiras,evoman} is a framework for testing competitive game-playing agents.
+    This framework is inspired by Mega Man II\cite{capcom}, a game created by Capcom.
+    Evoman is a 2D shooting game where the player controls an agent playing against an opponent.
+    The agent will collect information about the environment through 20 sensors (Fig. 1):
     \begin{itemize}
@@
     \end{figure}
-    The actions which the agent may take are:
-    \begin{itemize}
-        \item walk left
-        \item walk right
-        \item jump
-        \item shoot
-        \item release of the jump
-    \end{itemize}
+    The 5 actions which the agent may take are: walk left, walk right, shoot, press jump, release jump.
 
-    The lives of the player and the enemy start at 100.
-    Everytime one of them gets hit, their life deplenishes.
-    Whoever's life reaches 0 loses the game.
+    The player and opponent start with 100 life points. On each hit received, their life points decrease.
+    The player wins by reducing the opponent's life points to $0$, and lose if their life points are reduced to $0$ first.
 
-    In the original Capcom game the player would have to beat $8$ opponents and acquire their weapons
-    as they are defeated.
+    In the original Capcom game, the player would have to beat $8$ opponents and acquire their weapons
+    once they are defeated.
     The additional difficulty of EvoMan comes from the fact that the player has to defeat all
     the opponents using only the starting weapon.
-    Each opponent can be fought on a specified difficulty level.
+    Each opponent can be fought at a specific difficulty level.
     The difficulty level is an integer greater or equal than $1$ which is translated into a factor for the damage
-    taken and damage given by the player, the higher the difficulty level the lower the damage given and the higher the damage taken.
-    The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman_framework}}
-    and it is currently compatible with Python 3.6 and 3.7.
+    taken and damage given by the player; the higher the difficulty level the lower the damage given, and the higher the damage taken.
+    The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman_framework}}.
     There is also an extensive documentation
@@
     \subsection{Problem}\label{subsec:problem}
-    Beating an opponent is relatively easy when using specialized models against a specific enemy.
-    The authors of the EvoMan framework trained multiple specialized agents against all opponents.
-    After playing a game, the formula for the gain by which an agent is evaluated is:
-    \boldmath
+    How well an agent performs against an enemy is measured by how much damage they inflict, and how much damage to themselves they avoid - or, conversely, how much life points they have remaining, and how much life points the opponent has at the end of the game:
+
     \begin{gather*}
-        gain = 100.01 + player\_life - enemy\_life
+      gain = 100.01 + player\_life - enemy\_life \hspace{1cm} (1)
     \end{gather*}
-    \unboldmath
-    The highest final gain is 185.67\cite{evoman} and it was obtained with the NEAT\cite{neat} algorithm. \\
 
-    The problem we are trying to solve is a generalization of the one above.
-    We are to train an agent against four enemies and then test its performance against all eight.
-    The final score metric of an agent is the harmonic mean of the gains against all enemies.
-    The combination of four enemies we are to train the agent is not fixed.
-    The difficulty level for solving this problem is $2$ for all enemies in training and testing,
-    but we did some exploratory work with greater level of difficulties outside training and
-    testing for obtaining the final solution for the problem.
+    As both start with $100$ life points, the best possible value for gain is $200.01$, for a player which defeats the opponent without getting hit.
 
-    The specifications above, like the number of enemies chosen for training or the difficulty level,
-    come directly from a competition organized by the creators of the framework\cite{evoman_competition}.
+    The best known gain was $185.67$ (as a harmonic average of $8$ values, each for one of the $8$ opponents)
+    \cite{evoman} and was obtained with a NEAT\cite{neat} algorithm. It is this value we've tried to exceed.\\
+
+    There are also various levels of generalisation available; the least is training a single agent on a single opponent, as the above method does.
+The next is training a single agent on all opponents: instead of having $8$ specialised models for each opponent, a single unified model, trained on all $8$, could be attempted.
+Furthermore, we could train a single agent on a subset of the $8$, and test that agent on the full $8$.
+The competition\cite{evoman_competition} allowed free choice of $4$ training agents, and testing against all $8$.
+
+    Another level of generalisation is learning from one difficulty level and applying to the others.
+While seeing this possibility, we did not attempt such a task, due to the current high difficulty of the problem.
 
     \section{Approach}\label{sec:approach}
-    Our intention was to develop a cascade ensemble method where the resulting model of an algorithm
-    which can achieve a decent gain fast or has a high exploratory bias is the starting point of an algorithm with high exploitative bias.
-    The algorithm we expected to obtain a decent gain fast is a classic Q-learning\cite{q_learning} algorithm.
-The exploratory algorithms we have considered are genetic algorithms\cite{genetic_algorithm} and particle swarm optimization\cite{pso}.
-    The exploitative algorithm we used was Proximal Policy Optimization\cite{ppo}.
+    Our initial intention was to develop a cascade ensemble method where a fast algorithm could provide a better starting point to a slower algorithm.
+    We expected the classic Q-learning\cite{q_learning} algorithm to quickly obtain a decent gain.
+Two more exploratory algorithms we have considered are Genetic Algorithms\cite{genetic_algorithm} and Particle Swarm Optimisation\cite{pso}.
+    The slower algorithm, intended to exploit the starting points provided by the first algorithm in the cascade, is a Proximal Policy Optimization\cite{ppo} algorithm.
 
-    To find the best bootstraping algorithm we have made the following changes, which had no impact in the training and testing for the final solution to the problem:
+    To find the best first-stage algorithm we have made the following changes (which had no impact in the training and testing for the final solution to the problem, other than guiding the choice of algorithm):
     \begin{itemize}
-        \item we have increased the difficulty from the default 2 to 5.
+        \item we have increased the difficulty from the default 2 up to 5.
         \item the evaluation is made only on the second opponent due to the varied environment.
         \item the gain function was modified as in Karine Miras' analysis\cite{evoman_blog} to:
-        \boldmath
         \begin{gather*}
+          fitness = 
             0.9 \cdot (100 - enemy\_life) + 0.1 \cdot player\_life\\
-            -\hspace{0.2cm} ln(nr\_of\_game\_timesteps)\\
+            -\hspace{0.2cm} ln(nr\_of\_game\_timesteps) \hspace{1cm} (2)
         \end{gather*}
-        \unboldmath
     \end{itemize}
 
+    The best possible value for the fitness is $100$.
+
     The best exploratory algorithm would be trained on four enemies and
-    used in cascade with proximal policy optimization without the changes mentioned in this section.
+    used in cascade with PPO, without the changes mentioned in this section.
+
+    \subsection{Solution structure}
+    In all cases, a candidate solution is a fixed-structure feedforward Artificial Neural Network.
+
+    The input and output layers are determined by the game framework.
+Each game frame gives us $20$ sensor values, and we also remember the past $2$ frames, alongside the movements we've made.
+Thus, $3 \cdot 20 + 2 = 62$ inputs.
+
+    We've decided on two hidden layers.
+At first, we've used $32$ perceptrons on each, but later extended to $64$ (only when using PPO).
+We haven't increased the number of perceptrons or hidden layers further, since our main issue proved to be overfit, not the ANN's inability to properly approximate the required function\cite{cybenko}.
+    The activation function used is the Logistic Sigmoid; each layer has l2 regularisation applied to it, and each weight has a decay of $0.01$.
+    The weights are initialised in the range $[-2, 2]$, with a minimum precision of $6$ decimal paces.
 
     \subsection{Q-Learning}\label{subsec:q-learning}
-    The algorithm we expected to obtain a decent gain fast is a classic Q-learning\cite{q_learning} algorithm with neural networks.
+    We used a classic Q-Learning ANN to predict the reward function for each possible move.
+    After each predicted move, we update the neural network using backpropagation. We compute the reward as the difference between the current score and the previous score:
 
-    We used a neural network to predict the reward function for each possible move
-    (left, right, shoot, jump, release of jump) from a given state.
-    We then take the action with the highest predicted reward.
-    The input of this neural network is composed of the current game sensors and the
-    previous 2 game sensors with the moves taken at that specific point.
+    $ (prev\_enemy\_life - curr\_enemy\_life) \cdot 0.8 + (curr\_player\_life - prev\_player\_life) \cdot 0.2 $
 
-    The neural network used for experiments has 2 hidden layers with 32 neurons each.
-    Each layer has l2 regularization applied with a weight decay of 0.01 and sigmoid activation.
-    After each predicted move, we updated the neural network using backpropagation,
-    using as input the game sensors and as output the true reward function.
-    A game ends when either the agent or the enemy lose all life.
-    We have trained the agent on 5000 games.
-    The average number of frames per game for the best model is 287.
+    We have trained the agent on $5000$ games.
+%    The (arithmetic) average number of frames per game for the best model is $287$.
 
-    \subsection{Evolving Neural Network Weights with Genetic Algorithms}\label{subsec:evolving-neural-network-weights-with-genetic-algorithms}
-    We used genetic algorithms to evolve the weights of neural networks with fixed structures.
-
-    \subsubsection{Sparse Reward Genetic Algorithm}
-    The second algorithm is a sparse reward neuroevolution\cite{neuro}.
-    The reward is defined as "sparse" because an agent doesn't find out how well it's doing until
+    \subsection{Sparse Reward Genetic Algorithm}
+    The second algorithm is a sparse reward Neuroevolution\cite{neuro} algorithm.
+    The reward is defined as ``sparse'' because an agent doesn't find out how well it's doing until
     the end of the game, with no feedback during the game.
-    An individual is represented as the weights of a neural network.
-
-    The neural network role is to predict the next move using the current game sensors and
-    the previous 2 game sensors with the moves taken at that specific point.
-    We used 2 hidden layers with 32 neurons for each individual.
-
-    We start with randomly initialized neural network weights and then represent them as a bitstring.
-    The weights used for the neural network were values between 2 and -2 with a precision of 6 digits.
-
-    For evaluation, the bitstring is transformed into the weights of the neural network and a game is played,
-    from which we can obtain the individual fitness.
+    An individual represents the binary encoding of the weights of a Neural Network.
 
     Since we are representing the individuals as bitstrings we were able to apply
-    a simple genetic algorithm\cite{genetic_algorithm}.
+    a Simple Genetic Algorithm\cite{genetic_algorithm}.
 
-    The configurations we have used for the genetic algorithm is:
+    \pagebreak
+
+    The configurations we have used for the GA are:
     \begin{itemize}
-        \item population size: 50
-        \item number of generation: 500
-        \item crossover rate: 0.7
-        \item mutation rate: \{0.008, 0.1\}
-        \item elitism: 1
+        \item population size: $50$
+        \item number of generations: $500$
+        \item crossover rate: $0.7$
+        \item mutation rate: $\{0.008, 0.1\}$
+        \item elitism: $1$ individual
     \end{itemize}
-    We have tried two experiments with different mutation rates in order to observe whether a very high
-    mutation rate can lead to good results for the problem. \\
+    We have tried $2$ experiments with different mutation rates in order to observe if a high
+    mutation rate can lead to good results for the problem, since the role of this GA was
+    chiefly exploratory in our ensemble. \\
 
-    The solution of the genetic algorithm is the best individual from the last generation.
-    Since we have used an elitism of 1, this means that the solution of the genetic algorithm is
-    the best individual ever evaluated.
+    The solution of the GA is the best individual from the last generation.
 
-    \subsubsection{Iterative Genetic Algorithm}
-    The next algorithm is an iterative neuroevolution.
+    \subsection{Iterative Genetic Algorithm}
+    The next algorithm is an iterative neuroevolution one.
     It is "iterative" because the agents are trained on a small number of game steps
-    first and then the number of game steps slowly increases.
+    first, and then the number of game steps slowly increases.
+We've used this approach in order to save time; we believe it is reasonable to expect the initial agents to perform poorly, and thus require less time to decide on their quality.
 
-    After a number of generations we increase the number of game timesteps the agents are allowed to train on.
-    The fitness function was scaled based on the number of game timesteps in a way that if an agent
-    training on x game timesteps will always have a fitness lower than an agent training on y game
-    timesteps if x is lower than y.
+    The fitness function is scaled based on the number of game timesteps in a way that an agent
+    training on fewer timesteps will always have a fitness lower than an agent training on more timesteps.
 
-    \subsection{Searching Neural Network Weights with Particle Swarm Optimization}\label{subsec:searching-neural-network-weights-with-particle-swarm-optimization}
-    We have used particle swarm optimization\cite{pso} to search for the weights of a neural network of a
-    fixed structure which maximize the fitness defined in this section.
-    We used the same neural network configuration as in the case of genetic algorithms. \\
-
-    The configurations used for the particle swarm optimization\cite{pso} algorithm are:
+    \subsection{Particle Swarm Optimization}\label{subsec:searching-neural-network-weights-with-particle-swarm-optimization}
+    The configurations used for the PSO are:
     \begin{itemize}
-        \item population size: 30
-        \item number of iterations: 200
-        \item cognitive weight: \{0.4, 0.8, 1.5\}
-        \item social weight: \{0.8, 0.4, 3\}
-        \item inertia weight: \{constant 1, decreasing from 1 by 0.0035 every epoch\}
+        \item population size: $30$
+        \item number of iterations: $200$
+        \item cognitive weight: $\{0.4, 0.8, 1.5\}$
+        \item social weight: $\{0.8, 0.4, 3\}$
+        \item inertia weight: \{constant $1$, decreasing from $1$ by $0.0035$ every iteration\}
     \end{itemize}
-
-    The same separation between an sparse and an iterative search was made in the case
-    of particle swarm optimization, as in the case of the genetic algorithms.
-
-    We also tried to search with pso using an uniformly decreasing inertia weight with respect
-    to the percent of the iterations passed, having values from 1 to 0.3.
-The inertia weight controls the level of exploration versus exploitation, higher values in the beginning of the algorithm lead to better exploration while lower values towards the end of the algorithm lead to better exploitation.
-    This was tried to make the PSO focus even more on exploitation, rather than exploration.
+    The same separation between sparse and iterative evaluations was made in the case
+    of PSO, as was in the case of GA.
 
     \subsection{Proximal Policy Optimization}\label{subsec:proximal-policy-optimization}
-    PPO\cite{ppo} is the algorithm we have used for exploitation.
-    For this stage we have used the default difficulty of 2, the training was done on 4 opponents
-    and the evaluation was done as described in the Problem Description(\ref{sec:problem-description}).
-
-    The neural network is randomly initialized. \\
+    We have devoted significant computational resources to a PPO\cite{ppo}, expecting it to exploit and refine solutions.
     The configuration we have used for the PPO is:
     \begin{itemize}
-        \item hidden layers: (64, 64)
-        \item steps per epoch: 10000
-        \item epochs: 3000
-        \item gamma: 0.99
-        \item clip\_ratio: 0.2
-        \item pi\_lr: 3e-4
-        \item vf\_lr: 1e-3
-        \item train\_pi\_iterations: 80
-        \item train\_v\_iterations: 80
-        \item lambda: 0.97
-        \item target\_KL: 0.01
+        \item randomly initialized weights
+        \item steps per epoch: $10000$
+        \item epochs: $3000$
+        \item gamma: $0.99$
+        \item clip\_ratio: $0.2$
+        \item pi\_lr: $3e-4$
+        \item vf\_lr: $1e-3$
+        \item train\_pi\_iterations: $80$
+        \item train\_v\_iterations: $80$
+        \item lambda: $0.97$
+        \item target\_KL: $0.01$
     \end{itemize}
@@
     \subsection{Particle Swarm Optimization Cascaded With Proximal Policy Optimization}\label{subsec:particle-swarm-optimization-cascaded-with-proximal-policy-optimization}
-    The output weights of a neural network resulted from a PSO search is the starting agent for PPO\@.
-    In order to solve the generalized problem we have made the following changes from the PSO from the exploratory stage:
-    \begin{itemize}
-        \item The sizes of the hidden layers were increased from (32, 32) to (64, 64).
-        \item It is trained on four opponents, not only one.
-        \item The fitness function is the one from the Problem Description (\ref{sec:problem-description}).
-    \end{itemize}
-    The same configuration was used for PPO as in the case of the PPO with random
-    weights initialization.
+    The output weight configuration of an Artificial Neural Network resulting from a PSO search is the starting agent for PPO\@.
+    While using the same PSO configuration as before, the ANN hidden layers size is increased from $32$ to $64$.
 
@@
 
-    \subsection{Q-Learning}\label{subsec:q-learning2}
-    The Q-learning algorithm was the starting point of our comparison.
-    Even when trained and evaluated against the same opponent, it would lose every game
+    The Q-learning algorithm serves as the starting point of our comparison.
+    Even when trained and evaluated against the same opponent, it loses every game
     while inflicting almost no damage.
 
-    \subsection{Genetic Algorithms}\label{subsec:genetic-algorithms}
-    The iterative genetic algorithm leads to much better results than Q-learning (Fig.~\ref{fig:q_vs_ga_iterative}).
-    \begin{figure}[!h]
+    The iterative GA leads to much better results than Q-learning (Fig.~\ref{fig:q_vs_ga_iterative}).
+    We've also chosen a mutation probability of $0.008$. The results of the sparse GA and the iterative GA
+    are not significantly different (Fig.~\ref{fig:overview}).
+
+    Both PSO algorithms lead to much better results than either GA.
+    The iterative PSO leads to worse results than the sparse PSO\@.
+    In Fig.\ref{fig:overview} we can see that the best first-stage algorithm is the sparse PSO\@.
+
+    We have tested PSO performance (Fig.~\ref{fig:pso_levels}), on multiple difficulties.
+
+    \begin{figure}[!ht]
         \centering
         \includegraphics[width=0.5\textwidth]{old_images/q_vs_ga_iterative.eps}
-        \caption{The fitness comparison between Q-learning and iterative genetic algorithms.}
+        \caption{The fitness comparison between Q-learning and iterative genetic algorithms. Difficulty: $5$.}
         \label{fig:q_vs_ga_iterative}
@@
 
-    We have shown that a mutation rate of 0.008 leads to better results than a mutation rate of 0.1 (Fig.~\ref{fig:ga_mutation_rates}).
-    \begin{figure}[!h]
-        \centering
-        \includegraphics[width=0.5\textwidth]{old_images/ga_mutation_rates.eps}
-        \caption{The fitness comparison between sparse genetic algorithms with mutation rate of 0.1 and 0.008.}
-        \label{fig:ga_mutation_rates}
-    \end{figure}
+    \pagebreak
 
-    The results of the sparse genetic algorithm and the iterative genetic algorithm
-    are not significantly different (Fig.~\ref{fig:overview}).
-
-    \subsection{Particle Swarm Optimization}\label{subsec:sparse-particle-swarm-optimization}
-    Both particle swarm optimization algorithms lead to much better results than either of the
-    sparse and iterative genetic algorithms.
-    The iterative PSO lead to worse results than the sparse PSO\@.
-    In Fig.~\ref{fig:overview} we can see that the best exploratory algorithm is the sparse PSO\@.
-    \begin{figure}[!h]
+    \begin{figure}[!ht]
         \centering
@@
         \caption{The fitness comparison between the iterative and the sparse approached,
-        both for the genetic algorithms and the PSO. }
+        both for the genetic algorithms and the PSO.  Difficulty: $5$.}
         \label{fig:overview}
@@
 
-    The time advantage of iterative PSO vs sparse PSO is not significant (Fig.~\ref{fig:pso_sparse_vs_iterative_time}).
-    \begin{figure}[!h]
-        \centering
-        \includegraphics[width=0.5\textwidth]{old_images/pso_sparse_vs_iterative_time.eps}
-        \caption{The time comparison between sparse and iterative PSO.}
-        \label{fig:pso_sparse_vs_iterative_time}
-    \end{figure}
-
-    We have also searched for good PSO weights for our problem.
-    In the next stages we have used a cognitive weight of 0.4 and a social weight of 0.8 due to
-    its higher variance in results (Fig.~\ref{fig:pso_configs}).
-
-    \begin{figure}[!h]
-        \centering
-        \includegraphics[width=0.5\textwidth]{old_images/pso_configs.eps}
-        \caption{The fitness comparison between sparse PSO with different cognitive and social weights.}
-        \label{fig:pso_configs}
-    \end{figure}
-
-    We have also tried two PSO inertia weight updates (Fig.~\ref{fig:pso_inertia}).
-    We continued with the constant inertia weight approach.
-
-    \begin{figure}[!h]
-        \centering
-        \includegraphics[width=0.5\textwidth]{old_images/pso_inertia.eps}
-        \caption{The fitness comparison between PSO with constant inertia
-        weight and with uniformly decreasing in time inertia weight.}
-        \label{fig:pso_inertia}
-    \end{figure}
-
-    After deciding what the starting model should be, we have tested its performance (Fig.~\ref{fig:pso_levels})
-    and run time (Fig.~\ref{fig:pso_levels_time}) on multiple difficulties.
-
-    \begin{figure}[!h]
+    \begin{figure}[!ht]
         \centering
@@
 
-    \begin{figure}[!h]
-        \centering
-        \includegraphics[width=0.5\textwidth]{old_images/pso_levels_time.eps}
-        \caption{The time comparison of PSO at different game difficulty levels.}
-        \label{fig:pso_levels_time}
-    \end{figure}
+  %    \phantom{ }
 
-    \section{Results}\label{sec:results}
-    For evaluation of an agent, we ran 30 games against each opponent, leading to 8 averages
-    (one per opponent), of which we computed the harmonic mean which is the final result of the agent.
+    \section{Initial Results}\label{sec:results}
+    For evaluating an agent, we ran $30$ games against each opponent,
+    leading to $8$ averages ($1$ per opponent), of which we computed the harmonic mean,
+    which is the final result for the agent.
+    We use the harmonic mean because it is used for reporting other results for this problem\cite{evoman_competition}.
 
     \subsection{Best Combination of Opponents for Training}\label{subsec:best-combination-of-opponents-for-training}
-    We searched for the opponents that lead to the best results.
-    The first combination of train opponents we considered was \{1, 2, 6, 7\} which was chosen empirically after manually playing against every opponent and chosing the ones which behaved in such a way that covered the most general concepts our agents could learn in our opinion.
-    We observed its final gain as 125.37.
-    We realized a set of exploratory experiments where we would search for other combinations of train opponents,
-    but starting with the resulting model of PPO trained for 1000 epochs on the combination of opponents \{1, 2, 6, 7\}.
-    If by starting with a pre-trained model another combination of train enemies does not lead to better results,
-    then we can believe that the respective combination does not lead to better results than \{1, 2, 6, 7\}.
-    \begin{table}[htbp]
-        \caption{Results for various opponents when starting with a pre-trained PPO model on enemies \{1, 2, 6, 7\}}
-        \begin{center}
-            \begin{tabular}{|c|c|}
-                \hline
-                \textbf{Train opponents}&{\textbf{gain (harmonic mean)}} \\
-                \hline
-                $\{1, 2, 6, 7\} \cup \{3, 4, 5, 8\}$ & $44.16$ \\
-                $\{1, 2, 6, 7\} \cup \{3, 4, 6, 7\}$ & $75.3$ \\
-                $\{1, 2, 6, 7\} \cup \{1, 3, 6, 7\}$ & $110.19$ \\
-                $\{1, 2, 6, 7\} \cup \{2, 3, 6, 7\}$ & $136.1$ \\
-                $\{1, 2, 6, 7\} \cup \{2, 4, 6, 7\}$ & $85.22$ \\
-                \hline
-            \end{tabular}
-            \label{table:variations_of_opponents}
-        \end{center}
-    \end{table}
-    \\
-    The only combination that lead to a better score than 125.37 was \{2, 3, 6, 7\}.
-    We ran an experiment with PPO with random initialization with the combination of train enemies \{2, 3, 6, 7\} and
-    we obtained a final gain of 99.97.
-    This means that we did not find a better combination of train opponents than \{1, 2, 6, 7\}, so it is the one
-    we used for the following experiments.
+    We searched for the best $4$ opponents to serve as a basis for generalization.
+    The first subset considered was $\{1, 2, 6, 7\}$, which was chosen empirically after manually playing against every opponent.
+We choose the ones which exhibited the most general behaviors our agents could learn, in our opinion.
+We performed a non-exhaustive search, and found no other $4$-opponent subset better than the human-chosen one.
 
-    \subsection{Random Initialization PPO vs PSO Cascading PPO}\label{subsec:random-initialization-ppo-vs-pso-cascading-ppo}
-    The best PSO configuration was used in cascade before the PPO in 4 runs.
-    PPO with random initialization was ran 3 times.
-    The small number of runs is due to the long run time of the training.
-    The worst result of the PPO with random initialization is higher than the best result
-    of PSO cascaded with PPO\@ (Fig.~\ref{fig:random_vs_pso_initialization}).
+    \subsection{Random Initialization PPO vs Cascading PSO $+$ PPO}\label{subsec:random-initialization-ppo-vs-pso-cascading-ppo}
+    The best PSO configuration was used in cascade before the PPO in $4$ runs.
+    PPO with random initialization was ran $3$ times.
+    The small number of runs is due to the long training time.
+    The worst result of the PPO with random initialization is far better than the best result
+    of PSO cascaded with PPO\@, and we used this large difference to guide our choice, despite the small sample size. (Fig.~\ref{fig:random_vs_pso_initialization}).
     \begin{figure}[htbp]
@@
     \end{figure}
-    Our explanation for getting worse results when using the technique mentioned above is that the function landscape is very big and PSO doesn't explore enough.
-Another possible explanation would be that the landscape is tricky when it comes to searching for generalist versus specialized solutions.
-We could try solving this problem by giving more exploratory power to PSO, but we did not experiment with this due to the lack of time.
 
-    \subsection{PPO on the best configuration}\label{subsec:ppo-on-the-best-configuration}
-    Considering the results above, we decided that PPO with random initialization and enemies \{1, 2, 6, 7\} chosen for training would be the best configuration.
-We ran 3000 epochs of this configuration, saving all the models along the way after every 250 epochs.
-    After looking at the testing results, we noticed that the final gain is greater after 2000 epochs than after 3000 epochs (Fig.~\ref{fig:ppo_2000_vs_3000_epochs}).
-    This means that there is overfitting during the training.
-    We can conclude that by stopping the PPO algorithm after 2000 epochs instead of 3000 epochs
-    we can end up with more generalized agents that perform better on average against all opponents.
+    \subsection{PPO in the best configuration}\label{subsec:ppo-on-the-best-configuration}
+    Considering the results above, we decided that PPO with random initialization and training enemies $\{1, 2, 6, 7\}$ is the best configuration.
+We ran $3000$ epochs (chosen \textit{a priori}, considering only our experiences with PPO and general ANNs in other problems) in this configuration, saving all the models along the way after every $250$ epochs - since we anticipated overfit.
+    \textit{After} looking at the testing results, we noticed that the final \textit{test} gain is greater after $2000$ epochs than after $3000$ epochs (Fig.\ref{fig:ppo_2000_vs_3000_epochs}).
     \begin{figure}[htbp]
@@
         \includegraphics[width=0.5\textwidth]{images/ppo_2000_vs_3000_epochs.eps}
-        \caption{Gain comparison between PPO after 2000 epochs and after 3000 epochs.}
+        \caption{Gain comparison between PPO after $2000$ epochs and after $3000$ epochs.}
         \label{fig:ppo_2000_vs_3000_epochs}
@@
 
-    \begin{figure}[htbp]
-        \centering
-        \includegraphics[width=0.5\textwidth]{images/ppo_gain_evolution_3_runs.eps}
-        \caption{Gain Evolution for PPO in 3 runs, each model was ran 30 times on each opponent for calculating the gains}
-        \label{fig:ppo_gain_evolution_3_runs}
-    \end{figure}
+    \subsection{Best Train Agent vs Best Tested Agent}\label{subsec:best-tested-agent}
+    The highest training gain is $198.4$.
+    When computed against all opponents, this agent has an overall gain of $117.22$.
+    Out of the $3$ PPO with random initialization runs, the highest test gain is $127.01$ (but they had lower training gains, due to overfit).
+    After the experiments were concluded, we also looked at the overall gains of the agents before $3000$ epochs.
+    The highest test gain was observed during the third run after $1750$ epochs, and it has a value of $137.18$.
 
-    \subsection{Best Train Agent vs Best Tested Agent}\label{subsec:best-tested-agent}
-    The best train gain was obtained in the first run of the PPO with random initialization.
-    The train gain was computed as the harmonic mean of the averages per opponent of the 30 games played against
-    the enemies \{1, 2, 6, 7\}.
-    The highest train gain is 198.4.
-    When computed against all opponents, this agent has an overall gain of 117.22.
-    Out of the three PPO with random initialization runs, the highest final gain is 127.01.
-    After the experiments were concluded we also looked at the overall gains of the agents before 3000 epochs (Fig.~\ref{fig:ppo_gain_evolution_3_runs}).
-    For the runs 2 and 3 the snapshots of the models were saved every 250 epochs.
-    In the first run these periodic snapshots start after 2000 epochs.
-    When looking at the results of the intermediary agents we observed that there are some high peaks in test gain.
-    The highest test gain was observed during the third run after 1750 epochs and it has a value of 137.18.
-
-    \subsection{Comparison with the Upper Bound}\label{subsec:comparison-with-the-upper-bound}
-    With the PPO algorithm trained on the opponents \{1, 2, 6, 7\} we have
-    beat the best specialized models from the original paper which were given as upper bounds\cite{evoman}.
+    We were considerably computation-starved in these experiments; requiring a minimum sample size of $30$ meant we could only try out $3$ different agents.
+The competition organisers, upon running our algorithm, reported a different (and higher) value than in our experiments, of $138.14$ \cite{evoman_competition_results}.
 
@@
                 \cline{4-5}
-                & 3000 ep & 3000 ep & 3000 ep & 1750 ep & \\
-                & (best at training) & & & (best tested) & \\
+                & $3000$ ep & $3000$ ep & $3000$ ep & $1750$ ep & \\
+                & best train & & & best test & \\
                 \hline
-                1 & 198.41 & 199.07 & 199.81 & 199.61 & 190.01 \\
-                2 & 199.74 & 199.27 & 190.34 & 189.14 & 194.01 \\
-                3 & 58.94 & 46.67 & 70.27 & 85.01 & 180.01 \\
-                4 & 58.34 & 66.5 & 64.11 & 80.17 & 194.01 \\
-                5 & 172.61 & 165.65 & 174.61 & 158.83 & 194.01\\
-                6 & 195.73 & 196.95 & 195.85 & 194.37 & 173.01 \\
-                7 & 199.79 & 195.05 & 193.27 & 191.17 & 177.01 \\
-                8 & 122.17 & 145.49 & 141.89 & 140.61 & 186.01 \\
+                $1$ & $198.41$ & $199.07$ & $199.81$ & $199.61$ & $190.01$ \\
+                $2$ & $199.74$ & $199.27$ & $190.34$ & $189.14$ & $194.01$ \\
+                $3$ & $58.94$  & $46.67$  & $70.27$  & $85.01$  & $180.01$ \\
+                $4$ & $58.34$  & $66.5$   & $64.11$  & $80.17$  & $194.01$ \\
+                $5$ & $172.61$ & $165.65$ & $174.61$ & $158.83$ & $194.01$ \\
+                $6$ & $195.73$ & $196.95$ & $195.85$ & $194.37$ & $173.01$ \\
+                $7$ & $199.79$ & $195.05$ & $193.27$ & $191.17$ & $177.01$ \\
+                $8$ & $122.17$ & $145.49$ & $141.89$ & $140.61$ & $186.01$ \\
                 \hline
-                \textbf{\{1, 2, 6, 7\}} & & & & & \\
-                \textbf{harmonic mean} & 198.4 & 197.57 & 194.75 & 193.49 & 183.09 \\
+              $\{1, 2, 6, 7\}$ & $198.4$ & $197.57$ & $194.75$ & $193.49$ & $183.09$ \\
               \hline
-                \textbf{final} & & & & & \\
-                \textbf{harmonic mean} & 117.22 & 114.91 & 127.01 & 137.18 & 185.67\\
+              $\{1,\dots,8\}$  & $117.22$ & $114.91$ & $127.01$ & $137.18$ & $185.67$ \\
               \hline
-
             \end{tabular}
@@
     \begin{table}[htbp]
-        \caption{Specialized NEAT vs Generalized PPO (Percentage of games won)}
+        \caption{Specialized NEAT vs Generalized PPO (Percentage of games lost)}
         \begin{center}
@@
                 \cline{4-5}
-                & 3000 ep & 3000 ep & 3000 ep & 1750 ep\\
-                & (best at training) & & & (best tested) \\
+                & $3000$ ep & $3000$ ep & $3000$ ep & $1750$ ep\\
+                & best train & & & best test \\
                 \hline
-                1 & 100 & 100 & 100 & 100 \\
-                2 & 100 & 100 & 100 & 100 \\
-                3 & 3.33 & 0 & 3.33 & 46.66 \\
-                4 & 0 & 6.66 & 0 & 6.66 \\
-                5 & 100 & 96.66 & 100 & 100 \\
-                6 & 100 & 100 & 100 & 100 \\
-                7 & 100 & 93.33 & 100 & 100 \\
-                8 & 80 & 100 & 96.66 & 90 \\
+                $1$ & $0$     & $0$    & $0$     & $0$   \\
+                $2$ & $0$     & $0$    & $0$     & $0$   \\
+                $3$ & $96.67$ & $100$  & $96.67$ & $53.3$ \\
+                $4$ & $100$   & $93.3$ & $100$   & $93.3$  \\
+                $5$ & $0$     & $3.3$  & $0$     & $0$   \\
+                $6$ & $0$     & $0$    & $0$     & $0$   \\
+                $7$ & $0$     & $6.67$ & $0$     & $0$   \\
+                $8$ & $20$    & $0$    & $3.3$   & $10$    \\
                 \hline
@@
 
+%    \begin{table}[!htbp]
+%        \caption{Average player and enemy life. Average game duration (in frames)}
+%        \begin{center}
+%            \begin{tabular}{|c|c|c|c|c|c|c|}
+%                \hline
+%                \textbf{Opp}&\multicolumn{3}{|c|}{\textbf{Best train agent}}&\multicolumn{3}{|c|}{\textbf{Best test agent}} \\
+%                \cline{2-7}
+%                & \textbf{\textit{pl. life}}& \textbf{\textit{opp. life}}& \textbf{\textit{frames}} & \textbf{\textit{pl. life}}& \textbf{\textit{opp. life}}& \textbf{\textit{frames}} \\
+%              \hline
+%              1 & 97.26 &     0 &  159.63 &    100 &     0 &  174.23 \\
+%              2 & 98.33 &     0 &   176.6 &  89.73 &     0 &  186.43 \\
+%              3 &  0.13 & 43.33 &  457.26 &   7.66 & 21.33 &  410.53 \\
+%              4 &     0 & 37.66 &   547.5 &   1.84 & 21.66 &  736.83 \\
+%              5 & 71.32 &     0 &  379.13 &  60.52 &     0 &  367.93 \\
+%              6 & 96.32 &     0 &  165.06 &     95 &     0 &  203.53 \\
+%              7 & 99.74 &     0 &  162.06 &   85.9 &     0 &  135.13 \\
+%              8 &  34.3 &     6 &  520.93 &  38.76 &     2 &   443.6 \\ \hline
+%              \textbf{Avg} & 62.17 & 10.87 & 321.02 & 59.92 & 5.62 & 332.27 \\
+%              \hline
+%            \end{tabular}
+%        \end{center}
+%    \end{table}
+
+    \section{Increasing the Upper Bound}\label{subsec:increasing-the-upper-bound}
+    Our PPO implementation wasn't able to generalise sufficiently, neither in the competition (where our gain of $138.14$ was below the best gain of $146.88$)\cite{evoman_competition_results}, nor in our own tests.
+Overfitting was a significant problem, and we believe it to be a good indicator of our issues.
+
+    However, with the PPO trained on opponents $\{1, 2, 6, 7\}$, we have surpassed the best specialised opponent-specific models from the original paper, which were given as upper bounds\cite{evoman}.
+Thus, we can understand this issue as an exploration versus exploitation trade-off: while we haven't been able to develop the best exploratory, generalising model, we seem to have found a very good exploitative, targeted solution.
+
+    As such, we've performed $3$ further experiments which:
+    \begin{itemize}
+    \item develop a unified model for opponents $\{1, \dots, 8\}$, training on opponents $\{1, \dots, 8\}$, at difficulty $2$.
+    \item develop a unified model for opponents $\{1, \dots, 8\}$, training on opponents $\{1, \dots, 8\}$, at difficulties $\{1, \dots, 5\}$.
+    \item develop individual models for each opponent, training on each respective opponent, at difficulty $2$.
+    \end{itemize}
+
+    We shall call the unified model ``generalised'' and the individual model(s) ``specialised''.
+
+    Beyond these changes in aim, we haven't modified the above-described PPO algorithm employed.
+
+    \subsection{Single difficulty}
+
     \begin{table}[htbp]
-        \caption{Average player and enemy life. Average game duration}
+        \caption{PPO generalised vs PPO specialised vs NEAT specialised, difficulty $= 2$}
         \begin{center}
-            \begin{tabular}{|c|c|c|c|c|c|c|}
+            \begin{tabular}{|c|c|c|c|}
                 \hline
-                \textbf{Opponent}&\multicolumn{3}{|c|}{\textbf{Best train agent}}&\multicolumn{3}{|c|}{\textbf{Best test agent}} \\
-                \cline{2-7}
-                & \textbf{\textit{Avg player life}}& \textbf{\textit{Avg enemy life}}& \textbf{\textit{Avg duration}} & \textbf{\textit{Avg player life}}& \textbf{\textit{Avg enemy life}}& \textbf{\textit{Avg duration}} \\
+                \textbf{Opponent}&\textbf{PPO gen.}&\textbf{PPO spec.}&\textbf{NEAT spec.} \\
                 \hline
-                1 & 97.26 &   100 &     0 &     0 & 159.63 & 174.23 \\
-                2 & 98.33 & 89.73 &     0 &     0 &  176.6 & 186.43 \\
-                3 &  0.13 &  7.66 & 43.33 & 21.33 & 457.26 & 410.53 \\
-                4 &     0 &  1.84 & 37.66 & 21.66 &  547.5 & 736.83 \\
-                5 & 71.32 & 60.52 &     0 &     0 & 379.13 & 367.93 \\
-                6 & 96.32 &    95 &     0 &     0 & 165.06 & 203.53 \\
-                7 & 99.74 &  85.9 &     0 &     0 & 162.06 & 135.13 \\
-                8 &  34.3 & 38.76 &     6 &     2 & 520.93 & 443.6 \\
+                 1 &  199.54 &  199.67 &  190.01 \\
+                 2 &  200.01 &  199.61 &  194.01 \\
+                 3 &  194.81 &  199.94 &  180.01 \\
+                 4 &  180.61 &  195.95 &  194.01 \\
+                 5 &  187.77 &  198.03 &  194.01 \\
+                 6 &  174.43 &  199.67 &  173.01 \\
+                 7 &  183.07 &  197.73 &  177.01 \\
+                 8 &  169.31 &  195.07 &  186.01 \\
+                \hline
+                \textbf{harmonic mean} & 185.57 & 198.19 & 185.67 \\
                 \hline
             \end{tabular}
+            \label{PPO generalized vs specialized vs NEAT}
         \end{center}
@@
 
-    \subsection{Time Analysis}\label{subsec:time-analysis}
-    Since the match time is relevant, the machine on which the games were run is also relevant.
-    The tests were run on an i7 4750HQ processor.
-    The training was done on 3 threads.
-    All computations were done on the CPU\@.
-    The median game time during training is 28 seconds.
-    The average number of frames per game is 287.
+    The specialised PPO was able to surpass the specialised NEAT\cite{evoman} algorithm, both in harmonic mean and in each individual case (Table \ref{PPO generalized vs specialized vs NEAT}).
+The specialised PPO agents were trained for $1000$ epochs, as they achieved excellent results.
+
+    Additionally, PPO was able to create a general model which is extremely close (a $0.1$ gain difference) in harmonic mean to the results of the $8$ specialised models evolved through NEAT.
+
+    \subsection{Multiple difficulties}
+
+    \begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/general_harmonic_gain.eps}
+    \caption{Gain Evolution for PPO when trained against all 8 opponents.
+(In case of grayscale viewing: the plots are fully sorted, with difficulty $1$ being the highest, and difficulty $5$ the lowest.)}
+    \label{fig:ppo_train_8}
+    \end{figure}
+
+    \begin{table}[htbp]
+        \caption{PPO (generalised) trained against all opponents (Gain)}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|c|c|}
+                \hline
+                \textbf{Opponent}&\multicolumn{5}{|c|}{\textbf{Difficulty}} \\
+                \cline{2-6}
+                & \textbf{\textit{1}}& \textbf{\textit{2}}& \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} \\
+                \hline
+                 1 &  200.01 &  199.54 &  170.75 &  155.48 &   41.08\\
+                 2 &  200.01 &  200.01 &  182.81 &  168.81 &  168.51\\
+                 3 &  199.38 &  194.81 &  164.31 &  127.54 &  130.61\\
+                 4 &  186.14 &  180.61 &  154.23 &  181.58 &   60.28\\
+                 5 &  196.56 &  187.77 &  185.34 &  142.41 &  158.31\\
+                 6 &  196.35 &  174.43 &  179.19 &   23.84 &    9.88\\
+                 7 &  199.26 &  183.07 &   95.40 &    8.01 &    8.28\\
+                 8 &  196.57 &  169.31 &  123.00 &   41.34 &   10.01\\
+                \hline
+                \textbf{harmonic mean} & 196.68 & 185.57 & 149.57 & 35.76 & 20.89 \\
+                \hline
+            \end{tabular}
+            \label{PPO against all opponents gain}
+        \end{center}
+    \end{table}
+
+    \begin{table}[htbp]
+      \caption{PPO (specialised), trained $1000$ epochs (gain)}
+              \begin{center}
+            \begin{tabular}{|c|c|c|c|c|}
+                \hline
+                \textbf{Opponent} & \multicolumn{4}{|c|}{\textbf{Difficulty}} \\
+                \cline{2-5}
+                & \textbf{\textit{2}} & \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} \\
+                \hline
+                1                      & 199.68              & 196.51              & 189.21              & 166.01              \\
+                2                      & 199.61              & 199.91              & 199.21              & 197.18              \\
+                3                      & 199.94              & 199.51              & 147.21              & 72.28               \\
+                4                      & 195.95              & 196.14              & 198.81              & 191.23              \\
+                5                      & 198.03              & 193.38              & 191.73              & 196.71              \\
+                6                      & 199.67              & 195.18              & 196.85              & 198.76              \\
+                7                      & 197.73              & 192.45              & 184.85              & 176.01              \\
+                8                      & 195.07              & 185.37              & 178.05              & 172.61              \\
+                \hline
+                \textbf{harmonic mean} & 198.19              & 194.7               & 184.12              & 154.58              \\
+                \hline
+            \end{tabular}
+            \label{Specialized PPO against multiple difficulties (gain)}
+        \end{center}
+        \end{table}
+
+    \begin{table}[htbp]
+        \caption{PPO - percentage of games lost}
+        \begin{center}
+            \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
+                \hline
+              \textbf{Opp.}&\multicolumn{9}{|c|}{\textbf{Difficulty}} \\
+              \cline{2-10}
+                           & \multicolumn{5}{|c|}{Generalised} & \multicolumn{4}{|c|}{Specialised} \\
+              \cline{2-10}
+                & \textbf{\textit{1}}& \textbf{\textit{2}}& \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}} & \textbf{\textit{2}}& \textbf{\textit{3}} & \textbf{\textit{4}} & \textbf{\textit{5}}\\
+                \hline
+                 1 &  0 &  0 &     20 &  23.33 &   100 & 0 & 0 & 0     & 13.33 \\
+                 2 &  0 &  0 &      0 &      0 &     0 & 0 & 0 & 0     & 0 \\
+                 3 &  0 &  0 &      0 &  13.33 &  3.33 & 0 & 0 & 36.67 & 100 \\
+                 4 &  0 &  0 &      0 &      0 &    60 & 0 & 0 & 0     & 0 \\
+                 5 &  0 &  0 &      0 &      0 &     0 & 0 & 0 & 0     & 0 \\
+                 6 &  0 &  0 &      0 &    100 &   100 & 0 & 0 & 0     & 0 \\
+                 7 &  0 &  0 &  56.66 &    100 &   100 & 0 & 0 & 0     & 0 \\
+                 8 &  0 &  0 &     10 &    100 &   100 & 0 & 0 & 0     & 0 \\
+                \hline
+            \end{tabular}
+            \label{PPO against all opponents games lost}
+        \end{center}
+    \end{table}
+
+    Since our generalised PPO model is very close to the previous upper bound, we believe our single-model, all-opponent, all-difficulty exploration should provide a good estimation of what constitutes a good result on all difficulty levels.
+
+    At difficulty $5$, with opponent-specific models (Table \ref{Specialized PPO against multiple difficulties (gain)}) we also establish a good upper bound.
+$4$ opponents are easily defeated, and a single opponents proves difficult.
 
     \section{Conclusions}\label{sec:conclusions}
-    After observing that the sparse methods were better than the iterative ones,
-    PPO with random initialization is better than PSO cascaded with PPO and
-    PPO with random initialization trained on four opponents leads to better train results
-    than specialized agents trained with NEAT we can conclude that the problem space is in such a
-    way that the methods with a greedy bias are moving away from the global optimum.
+    \subsection{Upper Bound}
+    As stated previously, interpreting the initial issue as an exploration versus exploitation trade-off allowed us to repurpose the PPO algorithm.
+While it wasn't neither exploratory nor flexible enough to provide the best generalisation from $4$ opponents to $8$, it was exploitative enough to significantly increase the opponent-specific upper bound.
+
+    Even with a more modest $1000$-epoch runtime, the specialised PPO is able to easily exceed the previous upper bound\cite{evoman}, and approach the maximum gain of $200.01$.
+In fact, it does so for opponent $2$, apparently the easiest for our algorithm to defeat, according to Table \ref{PPO against all opponents gain}.
+
+    This previous upper-bound was established with a specialised (one model per opponent) NEAT algorithm, and we feel the comparison is fair: matching specialised algorithm to specialised algorithm.
+
+    While the previous upper bound does not have a time estimate attached to it, given our experience with both methods, we believe our proposed algorithm to be slower than NEAT, due to NEAT's ability to change, and minimse, the structure of the evolved ANN, and its ability to make larger exploratory jumps.
+For instance, establishing specialised upper bounds for each difficulty (Table \ref{Specialized PPO against multiple difficulties (gain)}) required $10$ million Evoman frames (accumulated during $1000$ training epochs) per opponent, per difficulty level.
+
+    The single-model, all-opponent PPO seems to match the previous upper bound; given the gain range of $[0, 200.01]$, and the range the results take values in, we believe that a difference of $0.1$ is not significant.
+Thus, we show that a single model can perform very well against all $8$ opponents (how an algorithm might generalise such a model from fewer opponents to all opponents remains an open, difficult problem).
+
+    We feel this is an important result: while the initial upper bound\cite{evoman} provided $8$ separate models to deal with $8$ opponents, we show that there exists a single model which is, in harmonic mean, just as good.
+We also believe that there is room for improvement for opponent-generalising algorithms.
+For the first statement, we have proof-by-existence: if our PPO found the model, then it most certainly exists, and is possible to achieve through other algorithms.
+The second statement - belief - we cannot prove yet; it would require showing that the behaviour patterns learned on one opponent can be applied to others, to a very high degree.
+
+    \subsection{Other Difficulty Levels}
+    In both our PSO (Fig.
+\ref{fig:pso_levels}) and PPO (Table \ref{PPO against all opponents gain}) explorations, we have found the game difficulty to matter a great deal.
+The generalised PPO, while inferior to the opponent-specialised PPO models in most cases, does establish the upper bound for opponent $3$ (Tables \ref{PPO against all opponents gain} and \ref{Specialized PPO against multiple difficulties (gain)}).
+
+    We'd like to propose difficulty $3$ as an interesting problem in itself.
+As shown in Fig.
+\ref{fig:ppo_train_8}, PPO shows the greatest improvement there, from the first to the last epoch.
+Difficulties $1$ and $2$ are  easier to solve, while $4$ and $5$ are currently very difficult - but all $4$ show relatively flat improvement curves.
+Algorithms playing Evoman at difficulty $3$ could show the greatest variance in results and further guide our algorithm-crafting decisions.
+
+    \subsection{Future Work}
+
+    We believe that, given the difficulty of the generalisation problem, pushing for greater exploration, while retaining our current level of exploitation are both required.
+We've attempted this with the $2$-stage cascade ensemble method, but we believe now that separating the exploratory and exploitative stages was a mistake (guided by our limited computational resources).
+The game is too difficult for the best strategies to be \textit{local} refinements of simple strategies.
+In other words, our results show that the function landscape contains multiple wider, shallower strictly-local optima basins, while the attraction basins for global optima are deep and narrow.
+There is also a degree of optimum dynamics: opponents don't perform the same recipe of actions, but have a limited ability to react to our agent / player.
+
+    Retaining the current level of exploration, while increasing exploitation can be done by greatly increasing the computational resources allocated to solving the problem\cite{nfl}.
+Based on our results (Fig.
+\ref{fig:overview}), we recommend a PSO$+$PPO Memetic Algorithm\cite{memetic}.
+
+    A way to sidestep the increase in computational time could be creating an on-line learning algorithm / agent.
+Thinking about the problem as a prediction problem - ``if half the opponents behaved in their certain ways, how will the rest behave?
+'' - , we can have $2$ basic ways of dealing with the prediction.
+
+%\vfill\null
+
+%\pagebreak
+
+    We can try to predict the full behaviour of the remaining test opponents ahead of time, or we can try to predict each action-by-action, as we confront them.
+Despite the increase in algorithmic difficulty for on-line learning, the second problem seems far easier than the first.
+Building on the previous proposal, we'd advise a sparse PSO$+$PPO Memetic Algorithm.
+While sparse PPO didn't achieve better results in our exploratory tests, they were evaluated only on basic (simpler, and requiring fewer computational resources) strategies.
+
+    \section{Acknowledgments}
+    We'd like to thank the creators and developers of Evoman for their implementation and documentation efforts. We'd also like to thank Ioana-Teodora Norocea for her editing help.
 
@@
     \bibitem{evoman_competition} Evoman: Game-playing Competition for WCCI 2020, \url{http://pesquisa.ufabc.edu.br/hal/Evoman.html}
+    \bibitem{evoman_competition_results} Evoman: Game-playing Competition for WCCI 2020, results \url{http://pesquisa.ufabc.edu.br/hal/Evoman.html#results}
     \bibitem{evoman} Fabricio Olivetti de Franca, Denis Fantinato, Karine Miras, A.E. Eiben and Patricia A. Vargas.
@@
     \bibitem{ppo} John Schulman, Filip Wolski, Prafulla Dhariwal, Alex Radford, Oleg Klimov (2017) "Proximal Policy Optimization Algorithms", arXiv:1707.06347v2
+    \bibitem{memetic} Moscato, P. (1989). "On Evolution, Search, Optimization, Genetic Algorithms and Martial Arts: Towards Memetic Algorithms". Caltech Concurrent Computation Program (report 826).
+    \bibitem{cybenko} Cybenko, G. (1989). "Approximations by superpositions of sigmoidal functions", Mathematics of Control, Signals, and Systems, 2(4), 303–314. doi:10.1007/BF02551274
+    \bibitem{nfl} Wolpert, D.H., Macready, W.G. (1997), "No Free Lunch Theorems for Optimization", IEEE Transactions on Evolutionary Computation 1, 67.
     \bibitem{evoman_blog} Karine Miras, Evoman, \url{https://karinemirasblog.wordpress.com/portfolio/evoman/}
+    \bibitem{spinningUp} Joshua Achiam, 2018, ``Spinning Up in Deep Reinforcement Learning'' \url{https://github.com/openai/spinningup}
     \end{thebibliography}

### 2020-08-28 paper/paper.tex
Add files via upload

@@
       we have surpassed the existing non-general, best-known upper-bound.
-      We have managed to exceed this upper bound with a Proximal Policy Optimization algorithm, by discarding the competition requirements to generalise.
+      We have managed to exceed this upper bound with a Proximal Policy Optimization(PPO) algorithm, by discarding the competition requirements to generalise.
       We also present our other exploratory attempts: Q-learning, Genetic Algorithms, Particle Swarm Optimisation, and their PPO hybridizations. 
@@
 The candidate solutions are weights on a fixed-structure Artificial Neural Network, which plays the Evoman game.
-The second stage, aimed at further refining the solutions found by the first stage, is a slower Proximal Policy Optimization\cite{ppo} algorithm.
+The second stage, aimed at further refining the solutions found by the first stage, is a slower Proximal Policy Optimization (PPO)\cite{ppo} algorithm.
 
@@
 However, its otherwise good performance makes it appropriate for setting a new upper bound, since the upper bound is being computed individually, by training and testing against each opponent in the Evoman game.
+
+    All datapoints from our research are arithmetic averages of $30$ repeats.
 
@@
     How well an agent performs against an enemy is measured by how much damage they inflict, and how much damage to themselves they avoid - or, conversely, how much life points they have remaining, and how much life points the opponent has at the end of the game:
-
+    \vspace{-0.1cm}
     \begin{gather*}
@@
 Two more exploratory algorithms we have considered are Genetic Algorithms\cite{genetic_algorithm} and Particle Swarm Optimisation\cite{pso}.
+For our algorithmic parameters, we chose widely-used starting values, performing small-scale experiments to fine-tune them.
+
     The slower algorithm, intended to exploit the starting points provided by the first algorithm in the cascade, is a Proximal Policy Optimization\cite{ppo} algorithm.
@@
 Thus, $3 \cdot 20 + 2 = 62$ inputs.
+The ANN has $5$ outputs, corresponding to the $5$ actions accessible to the agent.
 
@@
 
-    \pagebreak
+    %\pagebreak
 

### 2020-08-28 paper/slides.tex
Add files via upload

--- 2020-07-20 paper/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
+\documentclass[boxes]{beamer}
+%\usepackage[bookmarks=false,hidelinks,pdfpagelabels=false,hyperfootnotes=false,hyperindex=false,pageanchor=false,draft]{hyperref}
 \usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage[draft]{hyperref}
-
-\selectcolormodel{cmyk}
-
-\usepackage{tikz, pgfplots}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\interdisplaylinepenalty=2500
+\usepackage{array}
+\usepackage{mdwmath}
+\usepackage{mdwtab}
+\usepackage{stfloats}
+\usepackage{url}
+\usepackage{multirow}
+\usepackage{graphicx, tikz, pgfplots}
+\pgfplotsset{compat=1.16}
 \usetikzlibrary{calc}
@@
 
+\usepackage{attachfile}
+
+\usepackage{xargs}
+\usepackage{filecontents}
+%%\usepackage{datatool-pgfmath}
+\usepackage{datatool}
+\DTLsetseparator{ }
+\usepackage{xstring}
 \usepackage{ifthen}
-\usepackage{xargs}
 
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\makeatletter
+
+\makeatother
+
+%for printing, though this current paper should be black and white
+\selectcolormodel{rgb}
+
+%fit the plots with the page margins
+\def\plotWidth{8.5cm}
+
+% correct bad hyphenation here
+%\hyphenation{op-tical net-works semi-conduc-tor}
+
+\definecolor{SynascPurple}{RGB}{130,30,250}
+\setbeamercolor{title}{fg=SynascPurple}
+\setbeamercolor{frametitle}{fg=SynascPurple}
+\setbeamercolor{structure}{fg=SynascPurple}
+
+\setbeamertemplate{navigation symbols}{}%remove navigation symbols
+
 \begin{document}
@@
 
-    \author{\IEEEauthorblockN{Gabriel-Codrin \textbf{Cojocaru}}
-        \IEEEauthorblockA{\textit{Faculty of Computer Science} \\
-            \textit{``Al. I. Cuza'' University}\\
-            Iasi, Romania \\
-            gabriel.cojocaru@info.uaic.ro}
-          \and
+\author{Gabriel-Codrin \textbf{Cojocaru} \and Sergiu-Andrei \textbf{Dinu} (presenter) \and Eugen \textbf{Croitoru} }
 
-          \IEEEauthorblockN{Sergiu-Andrei \textbf{Dinu}}
-          \IEEEauthorblockA{\textit{Faculty of Computer Science} \\
-            \textit{``Al. I. Cuza'' University}\\
-            Iasi, Romania \\
-            sergiu.dinu@info.uaic.ro}
+\institute[shortinst]{\textit{Faculty of Computer Science} 
+            \textit{``Al. I. Cuza'' University}
+            Iasi, Romania 
+          }
 
-        \and
-        \IEEEauthorblockN{Eugen \textbf{Croitoru}}
-        \IEEEauthorblockA{\textit{Dept. of Computer Science} \\
-          \textit{``Al. I. Cuza'' University}\\
-          Iasi, Romania \\
-          eugennc@uaic.ro}
-      }
+\date{SYNASC 2020}
 
@@
 
-    \begin{abstract}
-      This paper describes a comparison between algorithms for evolving agents able to play the game Evoman.
-      Our team took part in the ``Evoman: Game-playing Competition for WCCI 2020'', and won second place;
-      beyond finding a good agent to satisfy the requirements of the competition - which aim at a good ability to generalise -,
-      we have surpassed the existing non-general, best-known upper-bound.
-      We have managed to exceed this upper bound with a Proximal Policy Optimization algorithm, by discarding the competition requirements to generalise.
-      We also present our other exploratory attempts: Q-learning, Genetic Algorithms, Particle Swarm Optimisation, and their PPO hybridizations. 
-      Finally, we map the behaviour of our algorithm in the space of game difficulty, generating plausible extensions to the existing upper-bound.
-    \end{abstract}
+\begin{frame}{TOC}
+  \tableofcontents
+  \end{frame}
 
-    \begin{IEEEkeywords}
-        game-playing agent, Artificial Intelligence, EvoMan, Genetic Algorithm, Reinforcement Learning,
-        Q-learning, Neuroevolution, Particle Swarm Optimization, Proximal Policy Optimization
-    \end{IEEEkeywords}
+\section{Introduction}
+\begin{frame}{Introduction: Game}
+  \begin{itemize}
+  \item Aim: play the Megaman/EvoMan game
+    \includegraphics[width=0.9\textwidth]{images/Evoman3.png}
+  \end{itemize}
+\end{frame}
 
-    \section{Introduction}\label{sec:introduction}
-    This paper contains our improvement on the upper-bound for playing the 2D shooting / platformer game Evoman\cite{evoman}.
-    Our team took part in the ``Evoman: Game-playing Competition for WCCI 2020''\cite{evoman_competition}.
-While placing second\cite{evoman_competition_results}, we have managed to exceed the best-known upper-bound.
-This difference is due to the competition requirements for generalisation; the upper-bound has been provided by the organisers by dropping such requirements, and only by doing the same could we exceed it.
+\begin{frame}{Introduction: Motivation}
+  \begin{itemize}
+  \item Games are an opportunity to improve our algorithms.
+  \item They can be low-cost simulators for interesting problems or topics.
+  \item EvoMan was used by a competition at WCCI 2020 (where we placed second). 
+  \item We found out that we can greatly improve on the upper bound provided.
+  \end{itemize}
+\end{frame}
 
-    Our initial exploratory attempts focused on an ensemble, $2$-stage cascade method.
-The first stage, designed to find good candidate solutions quickly, was either a Q-Learning\cite{q_learning}, a Genetic Algorithm\cite{genetic_algorithm} or Particle Swarm Optimization\cite{pso} algorithm.
-The candidate solutions are weights on a fixed-structure Artificial Neural Network, which plays the Evoman game.
-The second stage, aimed at further refining the solutions found by the first stage, is a slower Proximal Policy Optimization\cite{ppo} algorithm.
+\begin{frame}{Introduction: The difference}
+  \begin{itemize}
+  \item The initial competition required training with $4$ opponents, and testing on all $8$.
+  \item Our approach works better on simply fighting each opponent.
+  \item It significantly improves the state of the art results.
+  \end{itemize}
+\end{frame}
 
-    However, a simple random initialisation outperforms any of the first-stage algorithms, and the later experiments use PPO alone.
+\section{Problem description}
+\begin{frame}{Problem description: Evaluation}
+  \begin{gather*}
+    gain = 100.01 + player\_life - enemy\_life
+  \end{gather*}
+  \begin{itemize}
+  \item Both start with $100$ life points.
+  \item the absolute maximum score is $200.01$.
+  \item There are $8$ opponents, each with a different strategy.
+  \item There are $5$ difficulty levels, each lowering the damage the player does, and increasing the damage the player takes.
+  \end{itemize}
+\end{frame}
 
-    The competition required the algorithm to train with half the opponents in the game, and to test with all of them.
-Our method shows significant overfit in this challenge.
-However, its otherwise good performance makes it appropriate for setting a new upper bound, since the upper bound is being computed individually, by training and testing against each opponent in the Evoman game.
+\begin{frame}{Problem description: Agent}
+  \begin{itemize}
+  \item The game outputs $20$ sensor values, updated each simulation frame.
+  \item Each frame, we can perform $5$ actions (e.g. move left, shoot).
+    \pause
+  \item We chose to place an ANN in this loop.
+    \pause
+    \item We also remember the past $2$ frames' sensors values, and our past $2$ actions, for a total of $62$ inputs.
+  \end{itemize}
+\end{frame}
 
-    \section{Problem Description}\label{sec:problem-description}
+\section{Search Algorithms}
 
-    \subsection{Environment}\label{subsec:environment}
+\begin{frame}{Search Algorithms: $2$-stage approach}
+  \begin{itemize}
+  \item Idea: use an exploratory algorithm to find a good starting point. \pause Tested:
+    \begin{itemize}
+    \item Random initialisation.
+    \item Q-Learning.
+    \item Genetic Algorithm (GA).
+    \item Particle Swarm Optimisation (PSO).
+    \end{itemize}
+    \pause
+  \item then use an exploitative algorithm to refine that game-playing strategy. We've only used a Proximal Policy Optimisation (PPO) algorithm for this stage.
+  \end{itemize}
+\end{frame}
 
-    Evoman\cite{karinemiras,evoman} is a framework for testing competitive game-playing agents.
-    This framework is inspired by Mega Man II\cite{capcom}, a game created by Capcom.
-    Evoman is a 2D shooting game where the player controls an agent playing against an opponent.
-    The agent will collect information about the environment through 20 sensors (Fig. 1):
+\begin{frame}{Search Algorithms: initial comparison}
+  \begin{tabular}{|c|c|c|c|c|} \hline
+    Method                  & Q-Learning & GA & PSO & Random init.\\ \hline
+    Result (approx. median) & -5         & 46 &  59 & 117    \\ \hline
+  \end{tabular}
+  \pause
   \begin{itemize}
-        \item 16 correspond to horizontal and vertical distances to a maximum of 8 different opponent projectiles.
-        \item 2 correspond to the horizontal and vertical distance to the enemy.
-        \item 2 describe the directions the player and the enemy is facing
+  \item No pre-exploration algorithm performed better than random initialisation.
+  \item The game is too difficult for basic strategies to constitute the basis for optimal strategies (a misleading / trap function landscape).
   \end{itemize}
-    \begin{figure}
-        \centering
-        \includegraphics[width=0.5\textwidth]{images/Evoman3.png}
-        \caption{Sensors available for the player\cite{evoman}.}
-        \label{fig:sensors}
-    \end{figure}
-    The 5 actions which the agent may take are: walk left, walk right, shoot, press jump, release jump.
+\end{frame}
 
-    The player and opponent start with 100 life points. On each hit received, their life points decrease.
-    The player wins by reducing the opponent's life points to $0$, and lose if their life points are reduced to $0$ first.
+\section{Results}
+\begin{frame}{Results: Experiment}
+  \begin{itemize}
+  \item PPO ran for between $1000$ to $3000$ epochs, per opponent, $30$ repeats for each datapoint.
+  \item Time: $1000$ epochs require $\approx 10000000$ ($10M$) Evoman frames.
+  \item We ran specialised and generalised PPO models.
+  \item Specialised means $8$ models, $1$ for each of the $8$ opponents, $1000$ epochs.
+  \item Generalised means $1$ model for all $8$ opponents, $3000$ epochs.
+  \end{itemize}
+\end{frame}
 
-    In the original Capcom game, the player would have to beat $8$ opponents and acquire their weapons
-    once they are defeated.
-    The additional difficulty of EvoMan comes from the fact that the player has to defeat all
-    the opponents using only the starting weapon.
-    Each opponent can be fought at a specific difficulty level.
-    The difficulty level is an integer greater or equal than $1$ which is translated into a factor for the damage
-    taken and damage given by the player; the higher the difficulty level the lower the damage given, and the higher the damage taken.
-    The framework is freely available\footnote{\url{https://github.com/karinemiras/evoman_framework}}.
-    There is also an extensive documentation
-    available\footnote{\url{https://github.com/karinemiras/evoman_framework/blob/master/evoman1.0-doc.pdf}}.
-
-    \subsection{Problem}\label{subsec:problem}
-    How well an agent performs against an enemy is measured by how much damage they inflict, and how much damage to themselves they avoid - or, conversely, how much life points they have remaining, and how much life points the opponent has at the end of the game:
-
-    \begin{gather*}
-      gain = 100.01 + player\_life - enemy\_life \hspace{1cm} (1)
-    \end{gather*}
-
-    As both start with $100$ life points, the best possible value for gain is $200.01$, for a player which defeats the opponent without getting hit.
-
-    The best known gain was $185.67$ (as a harmonic average of $8$ values, each for one of the $8$ opponents)
-    \cite{evoman} and was obtained with a NEAT\cite{neat} algorithm. It is this value we've tried to exceed.\\
-
-    There are also various levels of generalisation available; the least is training a single agent on a single opponent, as the above method does.
-The next is training a single agent on all opponents: instead of having $8$ specialised models for each opponent, a single unified model, trained on all $8$, could be attempted.
-Furthermore, we could train a single agent on a subset of the $8$, and test that agent on the full $8$.
-The competition\cite{evoman_competition} allowed free choice of $4$ training agents, and testing against all $8$.
-
-    Another level of generalisation is learning from one difficulty level and applying to the others.
-While seeing this possibility, we did not attempt such a task, due to the current high difficulty of the problem.
-
-    \section{Approach}\label{sec:approach}
-    Our initial intention was to develop a cascade ensemble method where a fast algorithm could provide a better starting point to a slower algorithm.
-    We expected the classic Q-learning\cite{q_learning} algorithm to quickly obtain a decent gain.
-Two more exploratory algorithms we have considered are Genetic Algorithms\cite{genetic_algorithm} and Particle Swarm Optimisation\cite{pso}.
-    The slower algorithm, intended to exploit the starting points provided by the first algorithm in the cascade, is a Proximal Policy Optimization\cite{ppo} algorithm.
-
-    To find the best first-stage algorithm we have made the following changes (which had no impact in the training and testing for the final solution to the problem, other than guiding the choice of algorithm):
-    \begin{itemize}
-        \item we have increased the difficulty from the default 2 up to 5.
-        \item the evaluation is made only on the second opponent due to the varied environment.
-        \item the gain function was modified as in Karine Miras' analysis\cite{evoman_blog} to:
-        \begin{gather*}
-          fitness = 
-            0.9 \cdot (100 - enemy\_life) + 0.1 \cdot player\_life\\
-            -\hspace{0.2cm} ln(nr\_of\_game\_timesteps) \hspace{1cm} (2)
-        \end{gather*}
-    \end{itemize}
-
-    The best possible value for the fitness is $100$.
-
-    The best exploratory algorithm would be trained on four enemies and
-    used in cascade with PPO, without the changes mentioned in this section.
-
-    \subsection{Solution structure}
-    In all cases, a candidate solution is a fixed-structure feedforward Artificial Neural Network.
-
-    The input and output layers are determined by the game framework.
-Each game frame gives us $20$ sensor values, and we also remember the past $2$ frames, alongside the movements we've made.
-Thus, $3 \cdot 20 + 2 = 62$ inputs.
-
-    We've decided on two hidden layers.
-At first, we've used $32$ perceptrons on each, but later extended to $64$ (only when using PPO).
-We haven't increased the number of perceptrons or hidden layers further, since our main issue proved to be overfit, not the ANN's inability to properly approximate the required function\cite{cybenko}.
-    The activation function used is the Logistic Sigmoid; each layer has l2 regularisation applied to it, and each weight has a decay of $0.01$.
-    The weights are initialised in the range $[-2, 2]$, with a minimum precision of $6$ decimal paces.
-
-    \subsection{Q-Learning}\label{subsec:q-learning}
-    We used a classic Q-Learning ANN to predict the reward function for each possible move.
-    After each predicted move, we update the neural network using backpropagation. We compute the reward as the difference between the current score and the previous score:
-
-    $ (prev\_enemy\_life - curr\_enemy\_life) \cdot 0.8 + (curr\_player\_life - prev\_player\_life) \cdot 0.2 $
-
-    We have trained the agent on $5000$ games.
-%    The (arithmetic) average number of frames per game for the best model is $287$.
-
-    \subsection{Sparse Reward Genetic Algorithm}
-    The second algorithm is a sparse reward Neuroevolution\cite{neuro} algorithm.
-    The reward is defined as ``sparse'' because an agent doesn't find out how well it's doing until
-    the end of the game, with no feedback during the game.
-    An individual represents the binary encoding of the weights of a Neural Network.
-
-    Since we are representing the individuals as bitstrings we were able to apply
-    a Simple Genetic Algorithm\cite{genetic_algorithm}.
-
-    \pagebreak
-
-    The configurations we have used for the GA are:
-    \begin{itemize}
-        \item population size: $50$
-        \item number of generations: $500$
-        \item crossover rate: $0.7$
-        \item mutation rate: $\{0.008, 0.1\}$
-        \item elitism: $1$ individual
-    \end{itemize}
-    We have tried $2$ experiments with different mutation rates in order to observe if a high
-    mutation rate can lead to good results for the problem, since the role of this GA was
-    chiefly exploratory in our ensemble. \\
-
-    The solution of the GA is the best individual from the last generation.
-
-    \subsection{Iterative Genetic Algorithm}
-    The next algorithm is an iterative neuroevolution one.
-    It is "iterative" because the agents are trained on a small number of game steps
-    first, and then the number of game steps slowly increases.
-We've used this approach in order to save time; we believe it is reasonable to expect the initial agents to perform poorly, and thus require less time to decide on their quality.
-
-    The fitness function is scaled based on the number of game timesteps in a way that an agent
-    training on fewer timesteps will always have a fitness lower than an agent training on more timesteps.
-
-    \subsection{Particle Swarm Optimization}\label{subsec:searching-neural-network-weights-with-particle-swarm-optimization}
-    The configurations used for the PSO are:
-    \begin{itemize}
-        \item population size: $30$
-        \item number of iterations: $200$
-        \item cognitive weight: $\{0.4, 0.8, 1.5\}$
-        \item social weight: $\{0.8, 0.4, 3\}$
-        \item inertia weight: \{constant $1$, decreasing from $1$ by $0.0035$ every iteration\}
-    \end{itemize}
-    The same separation between sparse and iterative evaluations was made in the case
-    of PSO, as was in the case of GA.
-
-    \subsection{Proximal Policy Optimization}\label{subsec:proximal-policy-optimization}
-    We have devoted significant computational resources to a PPO\cite{ppo}, expecting it to exploit and refine solutions.
-    The configuration we have used for the PPO is:
-    \begin{itemize}
-        \item randomly initialized weights
-        \item steps per epoch: $10000$
-        \item epochs: $3000$
-        \item gamma: $0.99$
-        \item clip\_ratio: $0.2$
-        \item pi\_lr: $3e-4$
-        \item vf\_lr: $1e-3$
-        \item train\_pi\_iterations: $80$
-        \item train\_v\_iterations: $80$
-        \item lambda: $0.97$
-        \item target\_KL: $0.01$
-    \end{itemize}
-
-    \subsection{Particle Swarm Optimization Cascaded With Proximal Policy Optimization}\label{subsec:particle-swarm-optimization-cascaded-with-proximal-policy-optimization}
-    The output weight configuration of an Artificial Neural Network resulting from a PSO search is the starting agent for PPO\@.
-    While using the same PSO configuration as before, the ANN hidden layers size is increased from $32$ to $64$.
-
-    \section{Experimental Investigation}\label{sec:experimental-investigation}
-
-    The Q-learning algorithm serves as the starting point of our comparison.
-    Even when trained and evaluated against the same opponent, it loses every game
-    while inflicting almost no damage.
-
-    The iterative GA leads to much better results than Q-learning (Fig.~\ref{fig:q_vs_ga_iterative}).
-    We've also chosen a mutation probability of $0.008$. The results of the sparse GA and the iterative GA
-    are not significantly different (Fig.~\ref{fig:overview}).
-
-    Both PSO algorithms lead to much better results than either GA.
-    The iterative PSO leads to worse results than the sparse PSO\@.
-    In Fig.\ref{fig:overview} we can see that the best first-stage algorithm is the sparse PSO\@.
-
-    We have tested PSO performance (Fig.~\ref{fig:pso_levels}), on multiple difficulties.
-
-    \begin{figure}[!ht]
-        \centering
-        \includegraphics[width=0.5\textwidth]{old_images/q_vs_ga_iterative.eps}
-        \caption{The fitness comparison between Q-learning and iterative genetic algorithms. Difficulty: $5$.}
-        \label{fig:q_vs_ga_iterative}
-    \end{figure}
-
-    \pagebreak
-
-    \begin{figure}[!ht]
-        \centering
-        \includegraphics[width=0.5\textwidth]{old_images/overview.eps}
-        \caption{The fitness comparison between the iterative and the sparse approached,
-        both for the genetic algorithms and the PSO.  Difficulty: $5$.}
-        \label{fig:overview}
-    \end{figure}
-
-    \begin{figure}[!ht]
-        \centering
-        \includegraphics[width=0.5\textwidth]{old_images/pso_levels.eps}
-        \caption{The fitness comparison of PSO against different game difficulty levels.}
-        \label{fig:pso_levels}
-    \end{figure}
-
-  %    \phantom{ }
-
-    \section{Initial Results}\label{sec:results}
-    For evaluating an agent, we ran $30$ games against each opponent,
-    leading to $8$ averages ($1$ per opponent), of which we computed the harmonic mean,
-    which is the final result for the agent.
-    We use the harmonic mean because it is used for reporting other results for this problem\cite{evoman_competition}.
-
-    \subsection{Best Combination of Opponents for Training}\label{subsec:best-combination-of-opponents-for-training}
-    We searched for the best $4$ opponents to serve as a basis for generalization.
-    The first subset considered was $\{1, 2, 6, 7\}$, which was chosen empirically after manually playing against every opponent.
-We choose the ones which exhibited the most general behaviors our agents could learn, in our opinion.
-We performed a non-exhaustive search, and found no other $4$-opponent subset better than the human-chosen one.
-
-    \subsection{Random Initialization PPO vs Cascading PSO $+$ PPO}\label{subsec:random-initialization-ppo-vs-pso-cascading-ppo}
-    The best PSO configuration was used in cascade before the PPO in $4$ runs.
-    PPO with random initialization was ran $3$ times.
-    The small number of runs is due to the long training time.
-    The worst result of the PPO with random initialization is far better than the best result
-    of PSO cascaded with PPO\@, and we used this large difference to guide our choice, despite the small sample size. (Fig.~\ref{fig:random_vs_pso_initialization}).
-    \begin{figure}[htbp]
-        \centering
-        \includegraphics[width=0.5\textwidth]{images/random_vs_pso_initialization.eps}
-        \caption{Gain comparison between PPO with random initialization and with PSO cascading.}
-        \label{fig:random_vs_pso_initialization}
-    \end{figure}
-
-    \subsection{PPO in the best configuration}\label{subsec:ppo-on-the-best-configuration}
-    Considering the results above, we decided that PPO with random initialization and training enemies $\{1, 2, 6, 7\}$ is the best configuration.
-We ran $3000$ epochs (chosen \textit{a priori}, considering only our experiences with PPO and general ANNs in other problems) in this configuration, saving all the models along the way after every $250$ epochs - since we anticipated overfit.
-    \textit{After} looking at the testing results, we noticed that the final \textit{test} gain is greater after $2000$ epochs than after $3000$ epochs (Fig.\ref{fig:ppo_2000_vs_3000_epochs}).
-    \begin{figure}[htbp]
-        \centering
-        \includegraphics[width=0.5\textwidth]{images/ppo_2000_vs_3000_epochs.eps}
-        \caption{Gain comparison between PPO after $2000$ epochs and after $3000$ epochs.}
-        \label{fig:ppo_2000_vs_3000_epochs}
-    \end{figure}
-
-    \subsection{Best Train Agent vs Best Tested Agent}\label{subsec:best-tested-agent}
-    The highest training gain is $198.4$.
-    When computed against all opponents, this agent has an overall gain of $117.22$.
-    Out of the $3$ PPO with random initialization runs, the highest test gain is $127.01$ (but they had lower training gains, due to overfit).
-    After the experiments were concluded, we also looked at the overall gains of the agents before $3000$ epochs.
-    The highest test gain was observed during the third run after $1750$ epochs, and it has a value of $137.18$.
-
-    We were considerably computation-starved in these experiments; requiring a minimum sample size of $30$ meant we could only try out $3$ different agents.
-The competition organisers, upon running our algorithm, reported a different (and higher) value than in our experiments, of $138.14$ \cite{evoman_competition_results}.
-
+\begin{frame}{Results: vs the upper bound}
       \begin{table}[htbp]
-        \caption{Specialized NEAT vs Generalized PPO (Gain)}
-        \begin{center}
-            \begin{tabular}{|c|c|c|c|c|c|}
-                \hline
-                \textbf{Opponent}&\multicolumn{4}{|c|}{\textbf{PPO}}&\textbf{Specialized} \\
-                \cline{2-5}
-                & \textbf{\textit{Run 1}}& \textbf{\textit{Run 2}}& \multicolumn{2}{|c|}{\textbf{\textit{Run 3}}} & NEAT \\
-                \cline{4-5}
-                & $3000$ ep & $3000$ ep & $3000$ ep & $1750$ ep & \\
-                & best train & & & best test & \\
-                \hline
-                $1$ & $198.41$ & $199.07$ & $199.81$ & $199.61$ & $190.01$ \\
-                $2$ & $199.74$ & $199.27$ & $190.34$ & $189.14$ & $194.01$ \\
-                $3$ & $58.94$  & $46.67$  & $70.27$  & $85.01$  & $180.01$ \\
-                $4$ & $58.34$  & $66.5$   & $64.11$  & $80.17$  & $194.01$ \\
-                $5$ & $172.61$ & $165.65$ & $174.61$ & $158.83$ & $194.01$ \\
-                $6$ & $195.73$ & $196.95$ & $195.85$ & $194.37$ & $173.01$ \\
-                $7$ & $199.79$ & $195.05$ & $193.27$ & $191.17$ & $177.01$ \\
-                $8$ & $122.17$ & $145.49$ & $141.89$ & $140.61$ & $186.01$ \\
-                \hline
-              $\{1, 2, 6, 7\}$ & $198.4$ & $197.57$ & $194.75$ & $193.49$ & $183.09$ \\
-              \hline
-              $\{1,\dots,8\}$  & $117.22$ & $114.91$ & $127.01$ & $137.18$ & $185.67$ \\
-              \hline
-            \end{tabular}
-            \label{tab1}
-        \end{center}
-    \end{table}
-
-    \begin{table}[htbp]
-        \caption{Specialized NEAT vs Generalized PPO (Percentage of games lost)}
-        \begin{center}
-            \begin{tabular}{|c|c|c|c|c|}
-                \hline
-                \textbf{Opponent}&\multicolumn{4}{|c|}{\textbf{PPO}} \\
-                \cline{2-5}
-                & \textbf{\textit{Run 1}}& \textbf{\textit{Run 2}}& \multicolumn{2}{|c|}{\textbf{\textit{Run 3}}} \\
-                \cline{4-5}
-                & $3000$ ep & $3000$ ep & $3000$ ep & $1750$ ep\\
-                & best train & & & best test \\
-                \hline
-                $1$ & $0$     & $0$    & $0$     & $0$   \\
-                $2$ & $0$     & $0$    & $0$     & $0$   \\
-                $3$ & $96.67$ & $100$  & $96.67$ & $53.3$ \\
-                $4$ & $100$   & $93.3$ & $100$   & $93.3$  \\
-                $5$ & $0$     & $3.3$  & $0$     & $0$   \\
-                $6$ & $0$     & $0$    & $0$     & $0$   \\
-                $7$ & $0$     & $6.67$ & $0$     & $0$   \\
-                $8$ & $20$    & $0$    & $3.3$   & $10$    \\
-                \hline
-
-            \end{tabular}
-        \end{center}
-    \end{table}
-
-%    \begin{table}[!htbp]
-%        \caption{Average player and enemy life. Average game duration (in frames)}
-%        \begin{center}
-%            \begin{tabular}{|c|c|c|c|c|c|c|}
-%                \hline
-%                \textbf{Opp}&\multicolumn{3}{|c|}{\textbf{Best train agent}}&\multicolumn{3}{|c|}{\textbf{Best test agent}} \\
-%                \cline{2-7}
-%                & \textbf{\textit{pl. life}}& \textbf{\textit{opp. life}}& \textbf{\textit{frames}} & \textbf{\textit{pl. life}}& \textbf{\textit{opp. life}}& \textbf{\textit{frames}} \\
-%              \hline
-%              1 & 97.26 &     0 &  159.63 &    100 &     0 &  174.23 \\
-%              2 & 98.33 &     0 &   176.6 &  89.73 &     0 &  186.43 \\
-%              3 &  0.13 & 43.33 &  457.26 &   7.66 & 21.33 &  410.53 \\
-%              4 &     0 & 37.66 &   547.5 &   1.84 & 21.66 &  736.83 \\
-%              5 & 71.32 &     0 &  379.13 &  60.52 &     0 &  367.93 \\
-%              6 & 96.32 &     0 &  165.06 &     95 &     0 &  203.53 \\
-%              7 & 99.74 &     0 &  162.06 &   85.9 &     0 &  135.13 \\
-%              8 &  34.3 &     6 &  520.93 &  38.76 &     2 &   443.6 \\ \hline
-%              \textbf{Avg} & 62.17 & 10.87 & 321.02 & 59.92 & 5.62 & 332.27 \\
-%              \hline
-%            \end{tabular}
-%        \end{center}
-%    \end{table}
-
-    \section{Increasing the Upper Bound}\label{subsec:increasing-the-upper-bound}
-    Our PPO implementation wasn't able to generalise sufficiently, neither in the competition (where our gain of $138.14$ was below the best gain of $146.88$)\cite{evoman_competition_results}, nor in our own tests.
-Overfitting was a significant problem, and we believe it to be a good indicator of our issues.
-
-    However, with the PPO trained on opponents $\{1, 2, 6, 7\}$, we have surpassed the best specialised opponent-specific models from the original paper, which were given as upper bounds\cite{evoman}.
-Thus, we can understand this issue as an exploration versus exploitation trade-off: while we haven't been able to develop the best exploratory, generalising model, we seem to have found a very good exploitative, targeted solution.
-
-    As such, we've performed $3$ further experiments which:
-    \begin{itemize}
-    \item develop a unified model for opponents $\{1, \dots, 8\}$, training on opponents $\{1, \dots, 8\}$, at difficulty $2$.
-    \item develop a unified model for opponents $\{1, \dots, 8\}$, training on opponents $\{1, \dots, 8\}$, at difficulties $\{1, \dots, 5\}$.
-    \item develop individual models for each opponent, training on each respective opponent, at difficulty $2$.
-    \end{itemize}
-
-    We shall call the unified model ``generalised'' and the individual model(s) ``specialised''.
-
-    Beyond these changes in aim, we haven't modified the above-described PPO algorithm employed.
-
-    \subsection{Single difficulty}
-
-    \begin{table}[htbp]
-        \caption{PPO generalised vs PPO specialised vs NEAT specialised, difficulty $= 2$}
+        \caption{PPO generalised vs PPO specialised vs NEAT specialised, difficulty $= 2$ }
         \begin{center}
@@
       \end{table}
+\end{frame}
 
-    The specialised PPO was able to surpass the specialised NEAT\cite{evoman} algorithm, both in harmonic mean and in each individual case (Table \ref{PPO generalized vs specialized vs NEAT}).
-The specialised PPO agents were trained for $1000$ epochs, as they achieved excellent results.
-
-    Additionally, PPO was able to create a general model which is extremely close (a $0.1$ gain difference) in harmonic mean to the results of the $8$ specialised models evolved through NEAT.
-
-    \subsection{Multiple difficulties}
-
-    \begin{figure}[htbp]
-    \centering
-    \includegraphics[width=0.5\textwidth]{images/general_harmonic_gain.eps}
-    \caption{Gain Evolution for PPO when trained against all 8 opponents.
-(In case of grayscale viewing: the plots are fully sorted, with difficulty $1$ being the highest, and difficulty $5$ the lowest.)}
-    \label{fig:ppo_train_8}
-    \end{figure}
-
+\begin{frame}{Results: all difficulties, generalised}
   \begin{table}[htbp]
@@
     \end{table}
+\end{frame}
 
+\begin{frame}{\phantom{Results: all difficulties, generalised}}
+  \vspace{-1.1cm}
+  \begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.9\textwidth]{images/general_harmonic_gain.eps}
+    \caption{Gain Evolution for PPO when trained against all 8 opponents.}
+    \label{fig:ppo_train_8}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Results: all difficulties, specialised}
       \begin{table}[htbp]
@@
         \end{table}
+\end{frame}
 
+\begin{frame}{Results: percentage of games lost}
       \begin{table}[htbp]
@@
     \end{table}
+\end{frame}
 
-    Since our generalised PPO model is very close to the previous upper bound, we believe our single-model, all-opponent, all-difficulty exploration should provide a good estimation of what constitutes a good result on all difficulty levels.
+\begin{frame}{Conclusions}
+  \begin{itemize}
+  \item We have surpassed the state-of-the-art through an exploitative and time-consuming algorithm.
+  \item It is a difficult, misleading problem, since optimal game-playing strategies cannot be easily inferred from suboptimal strategies.
+  \item Thus, the cascade method cannot work well. A memetic algorithm (e.g. GSO + PPO) seems feasible.
+  \end{itemize}
+\end{frame}
 
-    At difficulty $5$, with opponent-specific models (Table \ref{Specialized PPO against multiple difficulties (gain)}) we also establish a good upper bound.
-$4$ opponents are easily defeated, and a single opponents proves difficult.
+%\section{Bibliography}
 
-    \section{Conclusions}\label{sec:conclusions}
-    \subsection{Upper Bound}
-    As stated previously, interpreting the initial issue as an exploration versus exploitation trade-off allowed us to repurpose the PPO algorithm.
-While it wasn't neither exploratory nor flexible enough to provide the best generalisation from $4$ opponents to $8$, it was exploitative enough to significantly increase the opponent-specific upper bound.
+\begin{thebibliography}{17}
 
-    Even with a more modest $1000$-epoch runtime, the specialised PPO is able to easily exceed the previous upper bound\cite{evoman}, and approach the maximum gain of $200.01$.
-In fact, it does so for opponent $2$, apparently the easiest for our algorithm to defeat, according to Table \ref{PPO against all opponents gain}.
-
-    This previous upper-bound was established with a specialised (one model per opponent) NEAT algorithm, and we feel the comparison is fair: matching specialised algorithm to specialised algorithm.
-
-    While the previous upper bound does not have a time estimate attached to it, given our experience with both methods, we believe our proposed algorithm to be slower than NEAT, due to NEAT's ability to change, and minimse, the structure of the evolved ANN, and its ability to make larger exploratory jumps.
-For instance, establishing specialised upper bounds for each difficulty (Table \ref{Specialized PPO against multiple difficulties (gain)}) required $10$ million Evoman frames (accumulated during $1000$ training epochs) per opponent, per difficulty level.
-
-    The single-model, all-opponent PPO seems to match the previous upper bound; given the gain range of $[0, 200.01]$, and the range the results take values in, we believe that a difference of $0.1$ is not significant.
-Thus, we show that a single model can perform very well against all $8$ opponents (how an algorithm might generalise such a model from fewer opponents to all opponents remains an open, difficult problem).
-
-    We feel this is an important result: while the initial upper bound\cite{evoman} provided $8$ separate models to deal with $8$ opponents, we show that there exists a single model which is, in harmonic mean, just as good.
-We also believe that there is room for improvement for opponent-generalising algorithms.
-For the first statement, we have proof-by-existence: if our PPO found the model, then it most certainly exists, and is possible to achieve through other algorithms.
-The second statement - belief - we cannot prove yet; it would require showing that the behaviour patterns learned on one opponent can be applied to others, to a very high degree.
-
-    \subsection{Other Difficulty Levels}
-    In both our PSO (Fig.
-\ref{fig:pso_levels}) and PPO (Table \ref{PPO against all opponents gain}) explorations, we have found the game difficulty to matter a great deal.
-The generalised PPO, while inferior to the opponent-specialised PPO models in most cases, does establish the upper bound for opponent $3$ (Tables \ref{PPO against all opponents gain} and \ref{Specialized PPO against multiple difficulties (gain)}).
-
-    We'd like to propose difficulty $3$ as an interesting problem in itself.
-As shown in Fig.
-\ref{fig:ppo_train_8}, PPO shows the greatest improvement there, from the first to the last epoch.
-Difficulties $1$ and $2$ are  easier to solve, while $4$ and $5$ are currently very difficult - but all $4$ show relatively flat improvement curves.
-Algorithms playing Evoman at difficulty $3$ could show the greatest variance in results and further guide our algorithm-crafting decisions.
-
-    \subsection{Future Work}
-
-    We believe that, given the difficulty of the generalisation problem, pushing for greater exploration, while retaining our current level of exploitation are both required.
-We've attempted this with the $2$-stage cascade ensemble method, but we believe now that separating the exploratory and exploitative stages was a mistake (guided by our limited computational resources).
-The game is too difficult for the best strategies to be \textit{local} refinements of simple strategies.
-In other words, our results show that the function landscape contains multiple wider, shallower strictly-local optima basins, while the attraction basins for global optima are deep and narrow.
-There is also a degree of optimum dynamics: opponents don't perform the same recipe of actions, but have a limited ability to react to our agent / player.
-
-    Retaining the current level of exploration, while increasing exploitation can be done by greatly increasing the computational resources allocated to solving the problem\cite{nfl}.
-Based on our results (Fig.
-\ref{fig:overview}), we recommend a PSO$+$PPO Memetic Algorithm\cite{memetic}.
-
-    A way to sidestep the increase in computational time could be creating an on-line learning algorithm / agent.
-Thinking about the problem as a prediction problem - ``if half the opponents behaved in their certain ways, how will the rest behave?
-'' - , we can have $2$ basic ways of dealing with the prediction.
-
-%\vfill\null
-
-%\pagebreak
-
-    We can try to predict the full behaviour of the remaining test opponents ahead of time, or we can try to predict each action-by-action, as we confront them.
-Despite the increase in algorithmic difficulty for on-line learning, the second problem seems far easier than the first.
-Building on the previous proposal, we'd advise a sparse PSO$+$PPO Memetic Algorithm.
-While sparse PPO didn't achieve better results in our exploratory tests, they were evaluated only on basic (simpler, and requiring fewer computational resources) strategies.
-
-    \section{Acknowledgments}
-    We'd like to thank the creators and developers of Evoman for their implementation and documentation efforts. We'd also like to thank Ioana-Teodora Norocea for her editing help.
-
-    \begin{thebibliography}{00}
     \bibitem{evoman_competition} Evoman: Game-playing Competition for WCCI 2020, \url{http://pesquisa.ufabc.edu.br/hal/Evoman.html}
@@
     \bibitem{capcom} M. MEGA, "Produced by capcom, distributed by capcom, 1987," System: NES.
+
+      \pagebreak
+
     \bibitem{q_learning} Watkins, C.J.C.H., Dayan, P. Q-learning. Machine Learning 8, 279–292 (1992)
@@
     \bibitem{ppo} John Schulman, Filip Wolski, Prafulla Dhariwal, Alex Radford, Oleg Klimov (2017) "Proximal Policy Optimization Algorithms", arXiv:1707.06347v2
+
+      \pagebreak
+
     \bibitem{memetic} Moscato, P. (1989). "On Evolution, Search, Optimization, Genetic Algorithms and Martial Arts: Towards Memetic Algorithms". Caltech Concurrent Computation Program (report 826).
@@
     \bibitem{spinningUp} Joshua Achiam, 2018, ``Spinning Up in Deep Reinforcement Learning'' \url{https://github.com/openai/spinningup}
+
 \end{thebibliography}
 
+% that's all folks
 \end{document}
